### 2017-05-11 Manuscript.tex
Initial files

% ****** Start of file aipsamp.tex ******
%
%   This file is part of the AIP files in the AIP distribution for REVTeX 4.
%   Version 4.1 of REVTeX, October 2009
%
%   Copyright (c) 2009 American Institute of Physics.
%
%   See the AIP README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex  aipsamp
%  2)  bibtex aipsamp
%  3)  latex  aipsamp
%  4)  latex  aipsamp
%
% Use this file as a source of example code for your aip document.
% Use the file aiptemplate.tex as a template for your document.
\documentclass[%
 aip,
%jmp,%
%bmf,%
%sd,%
rsi,%
 amsmath,amssymb,
%preprint,%
 reprint,%
%author-year,%
%author-numerical,%
]{revtex4-1}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

\newcommand	 {\sbar}	{{s}}
\newcommand	 {\rbar}	{{r}}
\newcommand	 {\hi}		{{h_\mathrm{i}}}
\newcommand	 {\pii}  	{{p_\mathrm{i}}}
\newcommand	 {\kb}		{{k_\mathrm{B}}}
\newcommand	 {\Tlow}	{{T_\mathrm{low}}}
\newcommand 	 {\Pnat} 	{{P_\mathrm{nat}}}
\newcommand {\Qa}		{{Q_\mathrm{A}}}
\newcommand {\Qb}		{{Q_\mathrm{B}}}
\newcommand {\Pcut}     	{{P_\mathrm{cut}}}
\newcommand {\TlowI}     {{T^\mathrm{I}_\mathrm{low}}}
\newcommand {\TlowII}    {{T^\mathrm{II}_\mathrm{low}}}
\newcommand {\Ptot}	{{P_\mathrm{tot}}}
\newcommand {\PaI}    	{{P_\mathrm{IA}}}
\newcommand {\PbI}    	{{P_\mathrm{IB}}}
\newcommand {\PaII}    	{{P_\mathrm{IIA}}}
\newcommand {\PbII}    	{{P_\mathrm{IIB}}}
\newcommand {\SR}		{{S144}}
\newcommand {\SI}		{{SBI}}
\newcommand {\SII}		{{SBII}}

\begin{document}

\preprint{AIP/123-QED}

\title[Multisequence Monte Carlo simulations]{Exploring mutation-induced protein fold switching using multisequence Monte Carlo simulations}

\author{A. Aina}
\author{S. Wallin}
\email{swallin@mun.ca}
\affiliation{ 
Memorial University of Newfoundland, Department of Physics and Physical Oceanography, A1B 3X7 St John's, NL, Canada}

\date{\today}

\begin{abstract}
We test an algorithm that can calculate the thermodynamic behavior of multiple amino acid  sequences in a single enhanced Monte Carlo run, in a coarse-grained protein model. 
\end{abstract}

\pacs{Add  here}

\keywords{Generalized-ensemble algorithm, simulated tempering, protein folding, protein fold switching, coarse-grained model}

\maketitle

\section{Introduction}
\noindent
Modern investigations into biophysical properties of biomolecular systems are routinely carried out not only on a single species but on multiple molecular variants.~\cite{Vukmirovic2000,Nickson2010} For example, protein engineering methods are used to provide mechanistic insight into the specificity of protein-protein and protein-DNA interactions, to compare the  conformational propensities of disease-related mutations of aggregation-prone proteins, and optimize biomolecular properties through directed evolution. Moreover, large-scale screening techniques such as phage display and microarray chips are used to provide biophysical properties across entire genomes. In short, through the study of a biomolecular system and related variants molecular variants are studied in order to gain a more complete picture of the biophysical and biochemical properties of the system and provide evolutionary context.

While protein or DNA sequences are easily changed on the computer, the high computational cost of simulating even a single moderately-sized biomolecule makes systematic explorations of sequence space challenging. Some notable efforts have nonetheless been attempted, including simulations of all 136 unique tetranucleotide sequences in B-DNA oligonucleotides using all-atom explicit water molecular dynamics.~\cite{Beveridge2004}

Here we explore a Monte Carlo (MC)-based algorithm for calculating the folding thermodynamics of multiple protein sequences in a single enhanced run. The algorithm relies on sampling at fixed temperature a generalized ensemble in which the sequence is treated as a dynamic parameter. Hence, this multisequence MC method includes two main types of updates: conformational updates $\rbar\rightarrow\rbar'$ and sequences updates $\sbar\rightarrow\sbar'$. As illustrated in Fig.~1, this approach requires that $\rbar$ and $\sbar$ are ``perpendicular" coordinates such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$. While this may be hard to achieve for detailed protein models, it is likely possible for many coarse-grained models.

We first apply the MS algorithm to a set of 144 model sequences with 16 amino acids and compare with the simulated tempering (ST) algorithm. We find that the computational efficiency of the MS algorithm is never less efficient than ST even at the lowest studied temperature. The model folds spontaneously designed sequences into stable native states with diverse protein-like sequences, and captures features such as mutational robustness. It therefore provide a biophysical basis for the sequence-structure and sequence-stability relationships of proteins at a conceptual level. This model has been used, e.g., to study the sequence-structure relationship of proteins in the "marshlands" between protein folds. Experimental studies have indicated that single point mutations are capable to take change one protein fold into another. The multisequence algorithm allow us to characterize the stability of and structure of all 1,024 sequence in the binary space between the two folds. The implication of these results for the fold switching and evolution is discussed. \\

We then demonstrate that the multisequence algorithm can be used to calculate the thermodynamic behavior for $>1,000$ sequences in a 3-letter model for protein folding~\cite{Bhattacherjee2012} at temperatures well below the typical folding temperatures $T_\mathrm{f}$ of the included sequences.

While the 16-amino acid A1 and N1 sequences fold into $\alpha$-helical and $\beta$-hairpin folds, respectively, the two 35-amino acid sequences A2 and TA adopt the folds denoted IIA and IIB, as shown in Fig.~X.

The large number of sequences considered allows us to carry out a much more systematic analysis than previously~\cite{Holzgrafe2014,Holzgrafe2015} of the sequence-structure-stability relationship in the sequence region that connect these two folds pairs. We find that...

\begin{figure}
\includegraphics[width=4.2cm]{Fig1}
\caption{The two types of Monte Carlo updates in the multisequence Monte Carlo algorithm.}
\end{figure}

\section{Theory}

\subsection{Generalized-ensemble algorithms and simulated tempering}
\noindent
Conventional Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy energy minima and hamper representative sampling of configurational space. The basic idea of generalized-ensemble algorithms~\cite{Mitsutake2001} is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters, such that a more efficient random walk in potential energy can be achieved.

A well-known generalized-ensemble algorithm is simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} in which it is the temperature $T$ that becomes a dynamic parameter. In this scheme, frequent visits to high-$T$ allow simulations to readily escape from local energy traps. The ST algorithm thus simulates the joint probability distribution 
\begin{equation}
P(\rbar,m) =\dfrac{1}{\hat{Z}} e^{-\beta_m E(\rbar) + g_m}\,,
\label{ST}
\end{equation}
where  $\beta_m=1/k_\mathrm{B} T_m$, $\{T_m\}_{m=1}^\mathrm{M}$ a set of temperatures and $k_\mathrm{B}$ is Boltzmann's constant. The normalization constant in Eq.~\ref{ST} is  
\begin{equation}
\hat{Z} = \sum_r \sum_{m=1}^{\mathrm{M}}e^{-\beta_m E(\rbar) + g_m}\,,
\end{equation}
where the first sum is over all conformations $\rbar$. The simulation parameters $g_m$ control the marginal probability distribution
\begin{equation}
P(k) = \frac{1}{\hat{Z}}\sum_r e^{-\beta_m E(\rbar) + g_m} \,,
\end{equation}
and must therefore be carefully chosen. A common and convenient choice is $g_m\approx \beta_m F_m$, where $F_m$ is the free energy at temperature $T_m$. With this choice, $P(k)$ becomes approximately flat ensuring all temperatures are frequently visited.

\subsection{Multisequence algorithm}
\noindent 
The basic idea of the MS algorithm for biomolecular simulation is to let the sequence $\sbar$ become a dynamic parameter rather than the temperature as in ST. A dynamic $\sbar$ is technically feasible as long as the potential energy can be written as $E(\sbar,\rbar)$, where $\sbar$ and $\rbar$ are independent variables. This is the case in our coarse-grained protein model with only backbone degrees of freedom and can also be achieved for more detailed  models~\cite{Bhattacherjee2013,Wallin2017}.

In analogy with ST, the MS algorithm simulates the joint probability distribution
\begin{equation}
P(\sbar,\rbar) =\dfrac{1}{\tilde{Z}}e^{-\beta E(\sbar,\rbar) + h(\sbar)}\,, 
\label{MS}
\end{equation}
where  
\begin{equation}
\tilde{Z} = \sum_{\sbar}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)}\,
\end{equation}
and the first sum goes over a set of allowed sequences $\sbar$. The simulation parameters $h(\sbar)$, similarly to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=\tilde{Z}^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.

Two types of MC updates are required to sample from the distrubution in Eq.~\ref{MS}, ordinary conformational update $\rbar\rightarrow\rbar'$ and mutational updates $\sbar\rightarrow\sbar'$. The acceptance probabilities for the latter  becomes
\begin{equation}
P_\mathrm{acc} (\sbar\rightarrow\sbar') = \min [1, \exp\{-\beta\Delta E+\Delta h\}]\,,
\label{accrej}
\end{equation}
where $\Delta E = E(\sbar',\rbar)-E(\sbar,\rbar)$ and $\Delta h = h_{\sbar'}-h(\sbar)$.

\section{Methods}
\subsection{Coarse-grained 3-letter model for protein folding}
\noindent
All calculations were carried out on a coarse-grained model for protein folding~\citenum{Bhattacherjee2012}. In this model, there are 3 different amino acid types: hydrophobic (h), polar (p) and turn-type (t). The backbone chain is represented atomistically by the N, H, $\mathrm{C}_\alpha$, $\mathrm{H}_{\alpha 1}$, C$'$ and O atoms. By contrast, the sidechain represention is simplified to a single enlarged $\mathrm{C}_\beta$ atom, which is geometrically identical for h and p types. The sidechain is absent for the t type which instead has an $\mathrm{H}_{\alpha 2}$ atom. The t type is therefore closely related to glycine. All bond lengths, bond angles, and peptide plane angles (180$^\circ$) are held fixed. Hence, an $N$-amino acid chain conformation $\rbar$ can, for any sequence $\sbar$, is therefore described by the set of 2$N$ backbone torsional angles $\{\phi_i$, $\psi_i\}_{i=1}^{N}$.

This geometrical description is paired with a simplified but finely tuned energy function with 4 terms: $E= E_\mathrm{ev} + E_\mathrm{loc} + E_\mathrm{hb} + E_\mathrm{hp}$. The first two, $E_\mathrm{ev}$ and $E_\mathrm{loc}$, represent excluded-volume effects of all atoms and local electrostatic effects, respectively. The hydrogen-bond energy, $E_\mathrm{hb}$, represents directionally dependent interactions between NH and CO groups which are necessary for secondary structure formation. Finally, the ``hydrophobicity" term, $E_\mathrm{hp}$, implements pairwise Lennard-Jones-like interactions between the $C_\beta$ atoms of h amino acids which are necessary for driving chain collapse during folding. Various model parameters, e.g., the strengths of hydrophobic attractions and hydrogen bonding, were determined based on the ability of the model to spontaneously fold a set of model sequences with 18-54 amino acids into  structurally diverse and thermodynamically stable native states with both $\beta$ and $\alpha$-structure. As it turned out, this strategy made the model robust enough to fold sequences designed to have mised $\alpha$ and $\beta$ structure.

\subsection{Sequence sets}
\noindent

\begin{table}
\caption{\label{tab:table1} Model sequences of different sequence lengths $N$ studied in this work. In addition to the 6 sequences listed, we studied sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs, respectively, (see text) denoted S16 and S35, respectively, and a restricted S16-set of 144 sequences denoted $\mathrm{S16}_{144}$ (also studied in Ref.~\protect\citep{Holzgrafe2014}).  }
\begin{ruledtabular}
\begin{tabular}{lcr}
Name & $N$ & Sequence\\
\hline
N1 & 16 & phphphpttphphphp \\
A1 & 16 & pphpphhpphpphhpp \\
Ran1 & 16 & \textbf{(add)}\\
Ran2 & 16 &  \textbf{(add)}\\
A2 & 35 & A1-ttt-A1\\
TN & 35 & A1-ttt-N1\\
\end{tabular}
\end{ruledtabular}
\end{table}

\subsection{Monte Carlo simulation parameters}
\noindent
Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assigned a new value between $-\pi$ and $\pi$; and (2) an approximately local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2003} In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are attempted every $1000$ step and carried out by randomly picking a new sequence $\sbar'\ne\sbar$ and applying the accept-reject criterion in Eq.~\ref{accrej}. In ST, temperature updates $m\rightarrow m'$ are attempted every 100 steps. Observables are sampled every 10 or 100 MC step. All simulations carried out in this work are summarized in Table~1.

\begin{table}
\caption{\label{tab:table1} List of simulations carried out in this work. }
\begin{ruledtabular}
\begin{tabular}{lccccr}
Runs & Algorithm & $T$ & MC steps &  Sequence set\\
\hline
32$\times$4\footnote{32 runs carried out for each of the 4 sequences. } & ST & $8\times 10^6$ &A1,N1,Ran1,Ran3\\
32$\times$8\footnote{32 independent runs carried out at each of the temperatures $T_1$ to $T_8$} & MS & $144\times 10^6$ & 16 &$\mathrm{S16}_{144}$\\
16 & MS & $5\times 10^9$ & 16 & S16 \\
16 & MS & $4\times 10^9$ & 35 & S35 \\
\end{tabular}
\end{ruledtabular}
\end{table}

\subsection{Stability measures}
\noindent
To quantify the stabilities of various folds, we use the measures used in Ref.~\cite{Holzgrafe2015}. We classify conformations $r$ as part of a fold if $Q\ge0.8$ where $Q\in [0,1]$ quantifies the similarity of $r$ to that fold. For folds A and B, $Q$ is determined by defining a set of backbone-backbone hydrogen bonds that are present in those folds (ignoring N and C terminal amino acids). For fold A, the fold-specific set of contacts (i,j) between the CO group of amino acid i and the NH group of amino acid j is $\{$(2,6), (3,7), (4,8), (5,9), (6,10), (7,11), (8,12), (9,13), (10,14), (11,15)$\}$, and for fold B the contact set is $\{$(3,14), (5,12), (7,10), (10,7), (12,5), (14,3)$\}$. A hydrogen bond is considered formed if (i) the NHO and COH angles both are $>90^\circ$ and (ii) the HO distance is $<2.7$~{\AA}.

%\begin{equation}
%Q_\mathrm{2\alpha} = ( Q_\alpha^\mathrm{1-16} + Q_\alpha^\mathrm{20-35} + Q_\mathrm{tert}) ) / 3
%\end{equation}
%
%\begin{equation}
%Q_\mathrm{\alpha\beta} = ( Q_\alpha^\mathrm{1-16} + Q_\beta^\mathrm{20-35} + Q_\mathrm{tert} ) / 3
%\end{equation}

\begin{figure}
\includegraphics[width=6.0cm]{MCtraj}
\caption{Example of a multisequence simulation.}
\end{figure}

%
%\begin{equation}
%Q_\mathrm{\alpha\beta\beta} = ( Q_\alpha^\mathrm{1-12} + Q_\beta^\mathrm{XX} +Q_\beta^\mathrm{20-35} + Q_\mathrm{tert} ) / 4
%\end{equation}
%
%
%\begin{equation}
%Q^*_\mathrm{\alpha\beta} = [ \max(Q_\alpha^\mathrm{1-16},Q_\beta^\mathrm{XX}) +Q_\beta^\mathrm{20-35} + Q_\mathrm{tert}) ] / 3
%\end{equation}
%

\section{Results}

\subsection{Computational efficiency of the multisequence algorithm}
\noindent
To assess the computational efficiency and properties of the multisequence algorithm, we used as a test case the set of 144 3-letter model sequences with 16 amino acids studied in Refs.\cite{Holzgrafe2014} Two of the sequences are A1 and N1 which were designed to fold into stable $\alpha$-helix and $\beta$-hairpin structures, respectively. A1 and N1 differ at 10 positions such that 10 consecutive point mutations can transform A1 into N1, and vice versa. The binary sequence space between A1 and N1 in which any combination of these mutations have been carried out, therefore contains $2^{10}=1,024$ sequences. The 144 sequences were selected from this binary space with the constraint that the total number of h-amino acids are not too high or too unevenly distributed.~\cite{Holzgrafe2014}

\begin{figure}
\includegraphics[width=8.0cm]{Stderr}
\caption{Computational efficiency of the multisequence and simulated tempering algorithm.}
\end{figure}

We apply the multisequence algorithm to this set of 144 sequences across a range of temperatures. Figure~2 illustrates a typical simulation carried out at the lowest temperature, $\Tlow$, which is well below the folding temperature of both A1 and N1~\cite{Holzgrafe2014,Holzgrafe2015}. From the MC evolution of the total energy $E$, sequence index j, and similarity measures $Q_A$ and $Q_B$, it is evident that visits to various sequences drive transitions into a range of structural states. For example, visits to A1 ($\mathrm{j}=1$) and N1 ($\mathrm{j}=144$) tend to coincide with visits to high-$Q_A$ and high-$Q_B$ conformations, respectively, as is necessary to generate the correct conformational ensembles for these sequences.

Sequence updates $\sbar\rightarrow\sbar'$ are carried out by picking a new sequence $\sbar'$ randomly from the entire set of allowed sequences. Therefore, a sequence update typically changes more than one amino acid position. In other words, if the step size in sequence space is $\Delta H$, then typically $\Delta H>1$. Interestingly, despite the large $\Delta H$ and low $T$, the overall (average) acceptance rate $P_\mathrm{acc}=0.xx$ for $\sbar\rightarrow\sbar'$ updates which is not unreasonably low. Higher $P_\mathrm{acc}$ can be obtained by restricting proposed moves such that $\Delta H\le\Delta H_\mathrm{max}$, as shown in Figure~3. For example, an acceptance rate close to the oft-quoted rule-of-thumb value $0.25~\cite{Gilks1996} $ is obtained with $\Delta H_\mathrm{max}=4$, and with $\Delta H_\mathrm{max} = 1$ (only point mutations), $P_\mathrm{acc}=0.xx$. For simplicity, in this work, all simulations are carried without restrictions on how new sequences $\sbar'$ are drawn.

\begin{figure}
\includegraphics[width=5.8cm]{Pacc}
\caption{Averages acceptance rates $P_\mathrm{acc}$ of mutational updates $\sbar\rightarrow\sbar'$ as a function of the number of changed amino acid positions, $\Delta H$.}
\end{figure}

To assess computational efficiency, we compare with results from simulated tempering (ST) simulations carried out on 4 of the 144 sequences, namely A1 and N1 and the two sequences R1 and R2 chosen randomly at $H=4$ and $H=6$ from A1, respectively (see Table~1). While ST provides the thermodynamics of a given sequence across the 8 temperatures $T_1, ..., T_8$ in a single run, a single MS simulation provides the thermodynamics of all 144 sequences for one $T_\mathrm{i}$. We adjust the simulation lengths for ST and MS runs such that an identical number of sampled conformations are obtained for each $\sbar$ and $T_\mathrm{i}$ combination, ensuring also that similar computational resources are used for the two algorithms.

Figure 4 shows the statistical error, $\sigma_{\left <E\right>}$, estimated for the average energy, $\left <E\right >$, for the 4 different sequences.

\begin{figure*}
\rotatebox{0}{\includegraphics[width=16.6cm]{ntwk16}}\\
\rotatebox{0}{\includegraphics[width=12.6cm]{ntwk35}}
\caption{Networks of the A1-N1 (top) and A2-TN (bottom) binary sequence space. }
\end{figure*}

Hence, we conclude that in the comparison made above the computational efficiency of the MS algorithm is comparable to ST. It must be stressed, however, that while the ST simulation requires that the thermodynamic behavior is calculated at a range of temperatures, MS works well on its own directly at low temperatures. ST was introduced as an extension to the standard, fixed-$T$ Metropolis-Hastings algorithm in order to overcome the poor mixing of states at low temperatures, i.e., the tendency for the simulations at low $T$ to become trapped in local free energy minima. The MS algorithm, however, efficiently calculates the thermodynamics at low $T$'s without visits to other $T$'s. In the MS algorithm, it is visits to sequences with poor native state stability rather than high $T$'s that drive the decorrelation of generated states and escape from local traps. This way, calculation of low-$T$ thermodynamics is enhanced in MS simulations while at the same time enhancing coverage in sequence space. This makes the MS algorithm ideal for situations in which the finite-$T$ thermodynamic behavior of multiple sequences are of interest.

\vspace{12pt}
\subsection{A1-N1 and A2-TN binary sequence spaces}
\noindent
We now apply the MS algorithm to the sets {\SI} and {\SII} which represent the full binary sequence spaces for the A1-N1 and A2-TN systems, respectively. Both sets include $2^{10}=1,024$ sequences. Using MS simulations, we determined the thermodynamic behavior at the temperatures $\TlowI$ and $\TlowII$ for the A1-N1 and A2-TN systems, respectively (see Table~1). In particular, this allow us to find the stabilities of folds IA and IB ($\PaI$ and $\PbI$) for all sequences in {\SI} and the stabilities of folds IIA and IIB ($\PaII$ and $\PbII$) for all sequences in {\SII}.

Having calculated these fold stabilities, we are in a position to determine if there are pathways in sequence space that lead to IA-IB or IIA-IIB fold changes without passing through an unstable intermediate sequence. To this end, we construct graphs in which each stable sequence is represented by a node and any two nodes are connected if their sequences differ at only one amino acid position. To determine if a sequence is stable we used the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PaI + \PbI$ and $\PaII + \PbII$ for A1-N1 and A2-AN1, respectively. The precise network obtained depends, of course, on the cut-off value of $\Pcut$. In general, a higher $\Pcut$ means a selection of more stable pathways. Fig.~5 shows the networks obtained with $\Pcut=0.50$. The number of viable mutational pathways that connect the start and end points are 516,972 between A1 and N1 and  57,912 between A2 and AN1. Because there are $10! =3,628,800$ possible mutational pathways between start and end points in both cases, these pathways represent 14.2 \% and 1.6 \% of all possible pathways in A1-N1 and A2-AN1, respectively. Hence, A1 and N1 are rather highly connected at this stability threshold. For $\Pcut=0.60$, the numbers are 104,640 pathways (2.9\%) for A1-N1 and 22,512 (0.6\%) pathways for A2-AN1. We find that there are no possible A1-to-N1 or A2-to-AN1 pathways when $\Pcut\ge0.74$ and $\Pcut\ge 0.66$, respectively. \\

\subsection{Stability properties of fold-to-fold mutational pathways}
\noindent 
An apparently general characteristic of designed and natural protein folding switching is that the stability of sequences near the switch point have reduced stability.~\cite{Bryan2010} The results from our model exhibit a similar trend. Fig.~6A and B shows the total stability $\Ptot$ at different mutational distances between start and end points. Intermediate sequences are less stable than at distances $h=0$ and $h=10$ although there are large variations as indicated by the spread between upper and lower bounds. Nonetheless, there is a clear statistical trend that sequences become less stability as more and more mutations are applied to the any of the 4 starting points. The minimum in stability occur closer to N1 or TN indicating that the all-$\alpha$ folds IA and IIA are more mutationally robust than the $\beta$-structure containing folds IB and IIB.

\begin{figure}
\includegraphics[width=7.8cm]{Paths}
\caption{Averages stability as a function of Hamming distance $h$. }
\end{figure}

%The smooth average stability trends in Fig.~6 might have been underpinned by individual mutational pathways that gradually shift the population between the two folds, however, this is not the case.

The smooth average stability trends in Fig.~6A and B belie the real character of the individual mutational pathways which abruptly switch from one fold to another. The switch in fold along individual pathways are typically completed in only 1-2 mutations. This can be seen from the distribution of switch lengths, $L_\mathrm{s}$, in Fig.~5C and D, taken over all pathways with $\Pcut=0.50$, where  $L_\mathrm{s} = 2$ for a switch that goes directly from a node $P_A>P_B$ to a node $P_B>P_A$ and $L_\mathrm{s} > 2$ if it also passes through one or more bistable sequences. Hence, for $\Pcut = 0.5$, it is fairly common that viable pathways pass through one or more sequences are that bistable, i.e., that have substantial populations of both folds.

Interestingly, the passage through bistable sequences become less common for more stable pathways in the A1-TN system, as can be seen from the dependence of $\left <L_\mathrm{s}\right >$ on $\Pcut$ (see Fig.~5C(inset)). For example, for $\Pcut=0.70$, there are no viable fold switching pathways between the $\alpha$-helix and $\beta$-hairpin folds that passes through a bistable point because $\left <L_\mathrm{s}\right > =  2$. The situation for the A2-TN system is different, where $\left <L_\mathrm{s}\right >$ is relatively stable around 3. An underlying reason for this behavior is apparent by comparing the total stability and bistability of all sequences. As shown in Fig.~5E and F, there is a clear trend that sequences with high total stability do not exhibit high bistability but instead they tend to have one clearly dominant fold.

%\subsection{Super-highways in sequence space}
%
%While there are a large number possible pathways (for $\Pcut<0.60$) in both systems, there are relatively few ways to pass between the two neutral nets. For $\Pcut=0.50$, we find that there are 109 nodes in the IA neutral net and 34 nodes in the IIB neutral net through which all viable pathways pass. For $\Pcut=0.60$, these numbers reduce to 14 and 10 nodes in the IA and IIA neutral nets, respectively.
%

\section{Discussion}

\section{Conclusion}

\bibliography{/Users/stefan/Documents/Manuscripts/Jshort,/Users/stefan/Documents/Manuscripts/MyRefs}

\end{document}

### 2017-05-11 Manuscript.tex
minor

@@
 A1 & 16 & pphpphhpphpphhpp \\
-Ran1 & 16 & \textbf{(add)}\\
-Ran2 & 16 &  \textbf{(add)}\\
-A2 & 35 & A1-ttt-A1\\
-TN & 35 & A1-ttt-N1\\
+R1 & 16 & \textbf{(add)}\\
+R2 & 16 &  \textbf{(add)}\\
+A2 & 35 & (A1)ttt(A1)\\
+TN & 35 & (A1)ttt(N1)\\
 \end{tabular}
@@
 \begin{tabular}{lccccr}
-Runs & Algorithm & $T$ & MC steps &  Sequence set\\
+Runs & Algorithm & $\kb T$  & MC steps &  Sequences\\
 \hline
-32$\times$4\footnote{32 runs carried out for each of the 4 sequences. } & ST & $8\times 10^6$ &A1,N1,Ran1,Ran3\\
-32$\times$8\footnote{32 independent runs carried out at each of the temperatures $T_1$ to $T_8$} & MS & $144\times 10^6$ & 16 &$\mathrm{S16}_{144}$\\
-16 & MS & $5\times 10^9$ & 16 & S16 \\
-16 & MS & $4\times 10^9$ & 35 & S35 \\
+32 & ST & 0.43--0.65 & $8\times 10^6$ &A1\\ 
+32 & ST & 0.43--0.65 & $8\times 10^6$ &N1\\ 
+32 & ST & 0.43--0.65 & $8\times 10^6$ &R1\\ 
+32 & ST & 0.43--0.65 & $8\times 10^6$ &R2\\ 
+32$\times$8\footnote{32  runs per temperature for 8 temperatures in the range given.} & MS &0.43--0.65& $144\times 10^6$ & $\mathrm{S16}_{144}$\\
+16 & MS & 0.43  & $5\times 10^9$ &  S16 \\
+16 & MS & 0.46 & $4\times 10^9$ &  S35 \\
 \end{tabular}

### 2017-05-12 Manuscript.tex
Minor changes

@@
 
-\section{Methods}
+\section{Model and Methods}
 \subsection{Coarse-grained 3-letter model for protein folding}
 \noindent
-All calculations were carried out on a coarse-grained model for protein folding~\citenum{Bhattacherjee2012}.
+All calculations were carried out using the coarse-grained model for protein folding developed in Ref.~\citenum{Bhattacherjee2012}.
 In this model, there are 3 different amino acid types: hydrophobic (h), polar (p) and turn-type (t).
@@
 
-\subsection{Sequence sets}
-\noindent
+\subsection{Model sequences}
+
+The sequences studied in this work are summarized in Table~I.
 
 \begin{table}
-\caption{\label{tab:table1} Model sequences of different sequence lengths $N$ studied in this work.
-In addition to the 6 sequences listed, we studied sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs, respectively, (see text) denoted S16 and S35, respectively, and a restricted S16-set of 144 sequences denoted $\mathrm{S16}_{144}$ (also studied in Ref.~\protect\citep{Holzgrafe2014}).
-}
+\caption{\label{tab1} List of model sequences of different lengths $N$ studied in this work.}
 \begin{ruledtabular}
 \begin{tabular}{lcr}
-Name & $N$ & Sequence\\
+Name & $N$ & Sequence\footnote{In addition, we study two sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs (see text) denoted $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$, respectively, and a restricted set $\mathrm{S16}_{144}$ with 144 sequences taken from S16 ($\mathrm{S16}_{144}$ was also studied in Ref.~\protect\citenum{Holzgrafe2014}).
+} \\
 \hline
+A1 & 16 & pphpphhpphpphhpp \\
 N1 & 16 & phphphpttphphphp \\
-A1 & 16 & pphpphhpphpphhpp \\
 R1 & 16 & \textbf{(add)}\\
@@
 \noindent
-Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assigned a new value between $-\pi$ and $\pi$; and (2) an approximately local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2003} In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are attempted every $1000$ step and carried out by randomly picking a new sequence $\sbar'\ne\sbar$ and applying the accept-reject criterion in Eq.~\ref{accrej}.
-In ST, temperature updates $m\rightarrow m'$ are attempted every 100 steps.
-Observables are sampled every 10 or 100 MC step.
-All simulations carried out in this work are summarized in Table~1.
+Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assigned a new value between $-\pi$ and $\pi$; and (2) a semi-local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2001} In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out by randomly picking a new sequence $\sbar'\ne\sbar$ and applying the accept-reject criterion in Eq.~\ref{accrej}.
+A sequence move is attempted every $1,000$ MC step while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
+The simulations carried out in this work are summarized in Tables~II.
 
 \begin{table}
-\caption{\label{tab:table1} List of simulations carried out in this work. }
+\caption{\label{tab2} List of simulations carried out in this work. }
 \begin{ruledtabular}
@@
 32$\times$8\footnote{32  runs per temperature for 8 temperatures in the range given.} & MS &0.43--0.65& $144\times 10^6$ & $\mathrm{S16}_{144}$\\
-16 & MS & 0.43  & $5\times 10^9$ &  S16 \\
-16 & MS & 0.46 & $4\times 10^9$ &  S35 \\
+16 & MS & 0.43  & $5\times 10^9$ &  $\mathrm{S16}_{1024}$ \\
+16 & MS & 0.46 & $4\times 10^9$ &  $\mathrm{S35}_{1024}$ \\
 \end{tabular}
@@
 \noindent
-To quantify the stabilities of various folds, we use the measures used in Ref.~\cite{Holzgrafe2015}.
+To quantify the stabilities of various folds, we use the measures introduced in Ref.~\cite{Holzgrafe2015}.
 We classify conformations $r$ as part of a fold if $Q\ge0.8$ where $Q\in [0,1]$ quantifies the similarity of $r$ to that fold.
+
 For folds A and B, $Q$ is determined by defining a set of backbone-backbone hydrogen bonds that are present in those folds (ignoring N and C terminal amino acids).
@@
 \includegraphics[width=6.0cm]{MCtraj}
-\caption{Example of a multisequence simulation.}
+\caption{Example of a multisequence simulation trajectory.}
 \end{figure}
@@
 Intermediate sequences are less stable than at distances $h=0$ and $h=10$ although there are large variations as indicated by the spread between upper and lower bounds.
-Nonetheless, there is a clear statistical trend that sequences become less stability as more and more mutations are applied to the any of the 4 starting points.
-The minimum in stability occur closer to N1 or TN indicating that the all-$\alpha$ folds IA and IIA are more mutationally robust than the $\beta$-structure containing folds IB and IIB.
+Nonetheless, there is a clear statistical trend that sequences become less stable as successive mutations are applied to any of the 4 starting points until a minimum is reached.
+These minima in stability occur closer to N1 and TN in the two systems, respectively, indicating that the all-$\alpha$ folds IA and IIA are more mutationally robust than the all-$\beta$ fold IB and mixed $\alpha$/$\beta$-fold IIB.
 
@@
 
-The smooth average stability trends in Fig.~6A and B belie the real character of the individual mutational pathways which abruptly switch from one fold to another.
+The smooth average stability trends in Fig.~6A and B, however, belie the real character of the \textit{individual} mutational pathways which tend to exhibit abrupt switches between the two folds. 
+%To examine how the fold transitions take place in our model, we classify all stable sequences (i.e.
+those with $\Ptot>\Pcut$) as either having one dominant fold, i.e., one of the two folds have little stability, or as bistable meaning that both folds have roughly similar stability $\PaI$ and $\PaII$.
+This can be seen from the distributions of switch lengths, $L_\mathrm{s}$, in Fig.~5C and D, which are taken over all pathways with $\Pcut=0.50$,
+%In determining $L_\mathrm{s}$, we define for each pathway how many nodes are included in the  ...
+where  $L_\mathrm{s} = 2$ for a switch that goes directly from a node $P_A>P_B$ to a node $P_B>P_A$ and $L_\mathrm{s} > 2$ if it also passes through one or more bistable sequences.
 The switch in fold along individual pathways are typically completed in only 1-2 mutations.
-This can be seen from the distribution of switch lengths, $L_\mathrm{s}$, in Fig.~5C and D, taken over all pathways with $\Pcut=0.50$, where  $L_\mathrm{s} = 2$ for a switch that goes directly from a node $P_A>P_B$ to a node $P_B>P_A$ and $L_\mathrm{s} > 2$ if it also passes through one or more bistable sequences.
 Hence, for $\Pcut = 0.5$, it is fairly common that viable pathways pass through one or more sequences are that bistable, i.e., that have substantial populations of both folds.

### 2017-05-18 Manuscript.tex
Small changes, added MCtraj.eps

@@
 \newcommand {\PbII}    	{{P_\mathrm{IIB}}}
-\newcommand {\SR}		{{S144}}
-\newcommand {\SI}		{{SBI}}
-\newcommand {\SII}		{{SBII}}
+\newcommand {\Eave}	{{E_\mathrm{ave}}}
+\newcommand {\sigE}	{{\sigma_{\left < E \right >}}}
+\newcommand {\SR}		{${\mathrm{S16}_{144}}$}
+\newcommand {\SI}		{${\mathrm{S16}_{1024}}$}	
+\newcommand {\SII}		{${\mathrm{S16}_{1024}}$}
 
@@
 
-\title[Multisequence Monte Carlo simulations]{Exploring mutation-induced protein fold switching using multisequence Monte Carlo simulations}
+\title[Multisequence Monte Carlo simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: application to protein folding switching}
 
@@
 \begin{abstract}
-We test an algorithm that can calculate the thermodynamic behavior of multiple amino acid  sequences in a single enhanced Monte Carlo run, in a coarse-grained protein model. 
+We evaluate a generalized-ensemble algorithm for coarse-grained simulations of biomolecules that allows the thermodynamics of multiple sequences to be calculated in a single ``multisequence" run.
+By carrying out a random walk in sequence space, conformational sampling is also enhanced.
+Escape from local energy minima is achieved by visiting sequences for which the minima is shallower or absent.
+We apply the algorithm to a coarse-grained model for protein folding with 7 atoms per amino acid and apply it to sets of sequences with up to 35 amino acids spanning sequence space  With the multisequence algorithm, we show that the folding thermodynamics of $>$1,000 sequences at relatively low temperatures can be readily calculated.
+We use the resulting data to explore the statistics of pathways in sequence space between the two pairs of folds.
+
 \end{abstract}
@@
 \end{equation}
-where $\Delta E = E(\sbar',\rbar)-E(\sbar,\rbar)$ and $\Delta h = h_{\sbar'}-h(\sbar)$.
+where $\Delta E = E(\sbar',\rbar)-E(\sbar,\rbar)$ and $\Delta h = h({\sbar'})-h(\sbar)$.
 
@@
 \includegraphics[width=6.0cm]{MCtraj}
-\caption{Example of a multisequence simulation trajectory.}
+\caption{(A) Example simulation trajectory of the MS algorithm applied to the sequence set $\mathrm{S16}_{144}$.
+The plot shows the MC evolution of the sequence $\sbar$ (numbered 1--144), the total potential energy $E$ and the root-mean-square deviation (RMSD) calculated against the fold IA (light blue) and fold IB (dark red) structures shown in (B).
+The simulation is carried out at $\kb T = 0.43$ corresponding to the lowest studied temperature.
+(B) Representative structures of fold IA and fold IB, selected to be the minimum-energy conformations found for the sequences A1 and N1, respectively.
+(C) Representative fold IIA and fold IIB structures selected to be the minimum-energy conformations of A2 and TN, respectively.
+}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=5.8cm]{Pacc}
-\caption{Averages acceptance rates $P_\mathrm{acc}$ of mutational updates $\sbar\rightarrow\sbar'$ as a function of the number of changed amino acid positions, $\Delta H$.}
+\includegraphics[width=7.8cm]{Pacc}
+\caption{Acceptance rates for $\sbar\rightarrow\sbar'$ updates in MS simulations of the {\SR} sequence set as a function of (A) the number of changed amino acid positions  and (B) temperature. }
 \end{figure}
@@
 
-Figure 4 shows the statistical error, $\sigma_{\left <E\right>}$, estimated for the average energy, $\left <E\right >$, for the 4 different sequences.
+Figure 4 shows the statistical error, $\sigE$, estimated for the average energy, $\left <E\right >$, for the 4 different sequences.
 
@@
 Fig.~6A and B shows the total stability $\Ptot$ at different mutational distances between start and end points.
-Intermediate sequences are less stable than at distances $h=0$ and $h=10$ although there are large variations as indicated by the spread between upper and lower bounds.
-Nonetheless, there is a clear statistical trend that sequences become less stable as successive mutations are applied to any of the 4 starting points until a minimum is reached.
-These minima in stability occur closer to N1 and TN in the two systems, respectively, indicating that the all-$\alpha$ folds IA and IIA are more mutationally robust than the all-$\beta$ fold IB and mixed $\alpha$/$\beta$-fold IIB.
+Intermediate sequences are less stable than at distances $h=0$ and $h=10$, although there are large variations as indicated by the spread between upper and lower bounds.
+Nonetheless, there is a clear statistical trend that sequences become gradually less stable as successive mutations are applied to any of the 4 starting points until a minimum is reached.
+%These minima in stability occur closer to N1 and TN in the two systems, respectively, indicating that the IA/IIA folds are more mutationally robust than the IB/IIB folds.
 
@@
 
-The smooth average stability trends in Fig.~6A and B, however, belie the real character of the \textit{individual} mutational pathways which tend to exhibit abrupt switches between the two folds. 
-%To examine how the fold transitions take place in our model, we classify all stable sequences (i.e.
-those with $\Ptot>\Pcut$) as either having one dominant fold, i.e., one of the two folds have little stability, or as bistable meaning that both folds have roughly similar stability $\PaI$ and $\PaII$.
-This can be seen from the distributions of switch lengths, $L_\mathrm{s}$, in Fig.~5C and D, which are taken over all pathways with $\Pcut=0.50$,
-%In determining $L_\mathrm{s}$, we define for each pathway how many nodes are included in the  ...
-where  $L_\mathrm{s} = 2$ for a switch that goes directly from a node $P_A>P_B$ to a node $P_B>P_A$ and $L_\mathrm{s} > 2$ if it also passes through one or more bistable sequences.
-The switch in fold along individual pathways are typically completed in only 1-2 mutations.
-Hence, for $\Pcut = 0.5$, it is fairly common that viable pathways pass through one or more sequences are that bistable, i.e., that have substantial populations of both folds.
+However, the smooth stability trends in Fig.~6A and B belie the real character of the \textit{individual} mutational pathways which tend to exhibit an abrupt switch between the two folds.
+To examine the statistics of fold transitions in our model, we first make a distinction between two types of stable sequences: those that fold into a single unique fold, thus behaving as classical proteins, and those that exhibit significant stabilities of both folds.
+Such ``bistable" sequences are interesting from a biophysical perspective in that they alternate between populating two different folds and hence are highly dynamic proteins.
+Indeed, they have been proposed to play an important role in the evolution of new protein folds.~\cite{Sikosek2016}
 
-Interestingly, the passage through bistable sequences become less common for more stable pathways in the A1-TN system, as can be seen from the dependence of $\left <L_\mathrm{s}\right >$ on $\Pcut$ (see Fig.~5C(inset)).
-For example, for $\Pcut=0.70$, there are no viable fold switching pathways between the $\alpha$-helix and $\beta$-hairpin folds that passes through a bistable point because $\left <L_\mathrm{s}\right > =  2$.
+We consider a stable sequence to be bistable if $B>0.5$, where $B$ is a bistability measure (see Fig.~6 legend); otherwise, it is considered a classical protein.
+In principle, a fold transition can occur directly between two classical proteins with unique folds, or it can proceed via one or more intermediate bistable sequences which populate both folds to some extent.
+We  define the switch length of a mutational pathway $L_\mathrm{s}=2+n_\mathrm{B}$, where $n_\mathrm{B}$ is the number of bistable sequences in between the two classical sequences that define the switch point.
+Hence, a path with $L_\mathrm{s}=2$ accomplishes a fold switch in a single mutational step without going through a bistable point.
+From the distributions of  $L_\mathrm{s}$ in Fig.~5C and D taken over all pathways with $\Pcut=0.50$, it can be seen that fold switching along individual pathways are typically completed in 1-2 mutations and often in a single step in the A1-N1 system.
+Hence, fold switching is typically abrupt in our model and, for $\Pcut = 0.5$, it is fairly common that viable pathways pass through one or two bistable sequences.
+
+Interestingly, fold switching through bistable sequences become less common as selection for more stable pathways is made in the A1-N1 system, as can be from the decrease in $\left <L_\mathrm{s}\right >$ as a function of $\Pcut$ (see Fig.~5C(inset)).
+For $\Pcut=0.70$, there are no longer any viable pathways between the $\alpha$-helix and $\beta$-hairpin folds that passes through a bistable point because $\left <L_\mathrm{s}\right > =  2$.
 The situation for the A2-TN system is different, where $\left <L_\mathrm{s}\right >$ is relatively stable around 3.
-An underlying reason for this behavior is apparent by comparing the total stability and bistability of all sequences.
-As shown in Fig.~5E and F, there is a clear trend that sequences with high total stability do not exhibit high bistability but instead they tend to have one clearly dominant fold.
+For this system, there are no direct $L_\mathrm{s}=2$ paths available even at $\Pcut=0.50$.
+Therefore, bistable sequence play a crucial role in bridging the folds IIA and IIB even through their stabilities are significantly reduced.
+
+An underlying reason for the preferential behavior is apparent by comparing the total stability and bistability of all sequences.
+As shown in Fig.~5E and F, there is a clear trend that sequences with high total stability do not exhibit high bistability but instead they tend to have one clearly dominant fold.\\
 
@@
 \section{Discussion}
+\noindent
+We have tested an algorithm for coarse-grained biomolecular simulations that relies on making the sequence of the molecule a dynamical parameter.
+
+For comparison, we used the well-established generalized-ensemble ST algorithm and focused on 144 model sequences with 16 amino acids ({\SR}).
+As a measure of computational efficiency, we used on the statistical error $\sigE$ in the estimation of the average total energy $E$ for 4 different sequences which are part of {\SR}.
+
+Our results show that at high $T$, well above the folding temperatures of both A1 and N1, we find that the statistical errors $\sigE$ for ST and MS are small and very similar.
+This is not surprising given that conformational sampling of relatively short polymer chains at high $T$ does not involve crossing major energy barriers.
+Therefore, successive sampled conformational states, for a given combination of sequence $\sbar$ and temperature $T$, are likely uncorrelated in both ST and MS simulations which, in turn, leads to similar estimates of $\Eave$ (and $\sigE$).
+At low $T$,...
+
+Hence, we conclude that the computational efficiency of the MS algorithm is comparable to ST.
+However, that while the ST simulation requires that the thermodynamic behavior is calculated at a range of temperatures, MS works well on its own directly at low temperatures.
+ST was introduced as an extension to the standard, fixed-$T$ Metropolis-Hastings algorithm in order to overcome the poor mixing of states at low temperatures, i.e., the tendency for the simulations at low $T$ to become trapped in local free energy minima.
+The MS algorithm, however, efficiently calculates the thermodynamics at low $T$'s without visits to other $T$'s.
+In the MS algorithm, it is visits to sequences with poor native state stability rather than high $T$'s that drive the decorrelation of generated states and escape from local traps.
+This way, calculation of low-$T$ thermodynamics is enhanced in MS simulations while at the same time enhancing coverage in sequence space.
+This makes the MS algorithm ideal for situations in which the finite-$T$ thermodynamic behavior of multiple sequences are of interest.
+
+A notable feature of the MS algorithm is that provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
+This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in lipid membranes, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Sampling in membrane simulations at high $T$, such as in replica-exchange simulations, are usually avoided in membrane simulations because it can lead to unwanted perturbation of the lipid bilayer structure although it can be a viable approach in particular cases.~\cite{Ulmschneider2010}
 
 \section{Conclusion}
+\noindent
+We have evaluated an algorithm for coarse-grained biomolecular simulations that allow the thermodynamics of multiple different sequences to be calculated in a single run.
+Here we showed that with this MS algorithm, the folding thermodynamics of over 1,000 amino acid sequences with each 35 amino acid could be determined in a coarse-grained model with 3 amino acid types.
+In addition to being a convenient for calculations of multiple sequences, the method enhances conformational sampling with an efficiency comparable to simulated tempering.
+Random walks across sequence space allow both sampling of many sequences and efficient escape from local free energy minima in the free energy landscape.
+The method might be well suited for coarse-grained simulations of other biomolecular systems, such as phospholipid bilayers.
 

### 2017-05-19 Manuscript.tex
Modified abstract and discussion

@@
 
-\title[Multisequence Monte Carlo simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: application to protein folding switching}
+\title[Multisequence Monte Carlo simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: application to protein fold switching}
 
@@
 \begin{abstract}
-We evaluate a generalized-ensemble algorithm for coarse-grained simulations of biomolecules that allows the thermodynamics of multiple sequences to be calculated in a single ``multisequence" run.
-By carrying out a random walk in sequence space, conformational sampling is also enhanced.
-Escape from local energy minima is achieved by visiting sequences for which the minima is shallower or absent.
-We apply the algorithm to a coarse-grained model for protein folding with 7 atoms per amino acid and apply it to sets of sequences with up to 35 amino acids spanning sequence space  With the multisequence algorithm, we show that the folding thermodynamics of $>$1,000 sequences at relatively low temperatures can be readily calculated.
-We use the resulting data to explore the statistics of pathways in sequence space between the two pairs of folds.
+We consider a generalized-ensemble algorithm for coarse-grained simulations of biomolecules, which allows the thermodynamic behaviors of two or more sequences to be determined in a single multisequence run.
+By carrying out a random walk in sequence space, the method also enhances conformational sampling.
+Escape from local energy minima is accelerated by visiting sequences for which the minima are more shallow or absent.
+We apply the method to an intermediate-resolution coarse-grained model for protein folding with 7 atoms per amino acids.
+The potential for large-scale coverage of sequence space is explored by applying the method to two sets with $>$1,000 sequences each.
+The resulting thermodynamic data is used to explore the sequence-structure and sequence-stability relationships between pairs of protein folds with different secondary structures.
+The potential implications for the evolution of new protein folds are discussed.
 
@@
 
-\keywords{Generalized-ensemble algorithm, simulated tempering, protein folding, protein fold switching, coarse-grained model}
+\keywords{Generalized ensembles, simulated tempering, protein folding, coarse-graining, evolution}
 
@@
 \noindent
-We have tested an algorithm for coarse-grained biomolecular simulations that relies on making the sequence of the molecule a dynamical parameter.
+We have evaluated a generalized-ensemble algorithm for coarse-grained simulations of biomolecules that works by making the sequence a dynamic parameter and thereby leading to a random walk in sequence space.
+We find that this approach has two main benefits.
+On the one hand, it provides a convenient way to sample the respective canonical distributions for a set of sequences in a single run and, on the other, it also enhances the sampling of conformational space.
 
-For comparison, we used the well-established generalized-ensemble ST algorithm and focused on 144 model sequences with 16 amino acids ({\SR}).
-As a measure of computational efficiency, we used on the statistical error $\sigE$ in the estimation of the average total energy $E$ for 4 different sequences which are part of {\SR}.
+The sampling efficiency of the MS algorithm can be assessed by the comparison with ST in Figure~3.
+At the highest studied temperatures, which are well above the folding temperatures of both A1 and N1, the statistical errors on the average total energy, $\sigE$, are very similar for the two methods.
+This is not unexpected because conformational sampling of short polymers at high $T$ does not involve crossing any major energy barriers.
+As a result, successively sampled conformations, for a given combination of sequence $\sbar$ and temperature $T$, are therefore likely mostly uncorrelated in both methods which in turn leads to similar $\sigE$ values.
 
-Our results show that at high $T$, well above the folding temperatures of both A1 and N1, we find that the statistical errors $\sigE$ for ST and MS are small and very similar.
-This is not surprising given that conformational sampling of relatively short polymer chains at high $T$ does not involve crossing major energy barriers.
-Therefore, successive sampled conformational states, for a given combination of sequence $\sbar$ and temperature $T$, are likely uncorrelated in both ST and MS simulations which, in turn, leads to similar estimates of $\Eave$ (and $\sigE$).
-At low $T$,...
+At lower temperatures, we find that the MS simulations often yields significantly smaller $\sigE$ than ST simulations.
+It is notable that this acceleration in conformational sampling is achieved despite that simulations are carried out at constant temperature.
+Hence, rather than promoting escape from local minima in the free energy landscape by visits to higher temperature, as in ST~\cite{Marinari1992,Lyubartsev1992} or replica-exchange,~\cite{Swendsen1986} the MS algorithm promotes local minima escape through visits to other sequences for which the minima may not be as severe or are absent altogether.
+For the same reason, the efficiency of the MS algorithm likely depends on the size and character of sequences included in the simulation.
+In particular, the MS simulation is likely dependent on the inclusion of at least some less stable sequences to promote at least partial unfolding of the chain and the transition to different conformational states.
 
-Hence, we conclude that the computational efficiency of the MS algorithm is comparable to ST.
+Clearly, the MS method should not be seen as a general simulation method to speed up conformational sampling due to the complexity of performing sequence updates in most biomolecular models.
+However, our results indicate that for coarse-grained models that permit the sequence updates as a Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be highly efficient.
+A notable feature of the algorithm is that it provides escape from local minima on the free energy landscape of proteins without resorting to a change in $T$.
+This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in the context or ordered dense phases such as lipid bilayers, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Simulations of lipid bilayers at elevated $T$ must typically be avoided because it can lead to unwanted perturbations of the membrane structure, although it can be a viable approach to speed up conformational sampling in particular cases.~\cite{Ulmschneider2010}
+
+%Overall our results show that, when technically feasible, a multisequence approach can be highly efficient for calculating the thermodynamics of a large number of amino acid sequences even at low temperatures.
+To demonstrate that the method to large sequence sets we applied it to the full binary sequences spaces between the model sequence pairs A1-N1 and A2-TN, which both include 1,024 sequences.
+These calculations would have been difficult to carry out ``in series" using some other standard method such as simulated tempering.
+
+%Hence, we conclude that the computational efficiency of the MS algorithm is comparable to ST.
 However, that while the ST simulation requires that the thermodynamic behavior is calculated at a range of temperatures, MS works well on its own directly at low temperatures.
@@
 
-A notable feature of the MS algorithm is that provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
+%A notable feature of the MS algorithm is that provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
 This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in lipid membranes, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Sampling in membrane simulations at high $T$, such as in replica-exchange simulations, are usually avoided in membrane simulations because it can lead to unwanted perturbation of the lipid bilayer structure although it can be a viable approach in particular cases.~\cite{Ulmschneider2010}

### 2017-05-22 Manuscript.tex
Methods

@@
 \newcommand 	 {\Pnat} 	{{P_\mathrm{nat}}}
-\newcommand {\Qa}		{{Q_\mathrm{A}}}
-\newcommand {\Qb}		{{Q_\mathrm{B}}}
+\newcommand {\QIA}	{{Q_\mathrm{IA}}}
+\newcommand {\QIB}	{{Q_\mathrm{IB}}}
+\newcommand {\QIIA}	{{Q_\mathrm{IIA}}}
+\newcommand {\QIIB}	{{Q_\mathrm{IIB}}}
 \newcommand {\Pcut}     	{{P_\mathrm{cut}}}
@@
 \newcommand {\Ptot}	{{P_\mathrm{tot}}}
-\newcommand {\PaI}    	{{P_\mathrm{IA}}}
-\newcommand {\PbI}    	{{P_\mathrm{IB}}}
-\newcommand {\PaII}    	{{P_\mathrm{IIA}}}
-\newcommand {\PbII}    	{{P_\mathrm{IIB}}}
+\newcommand {\PIA}    	{{P_\mathrm{IA}}}
+\newcommand {\PIB}    	{{P_\mathrm{IB}}}
+\newcommand {\PIIA}    	{{P_\mathrm{IIA}}}
+\newcommand {\PIIB}    	{{P_\mathrm{IIB}}}
 \newcommand {\Eave}	{{E_\mathrm{ave}}}
@@
 
-\subsection{Stability measures}
+\subsection{Observables}
 \noindent
-To quantify the stabilities of various folds, we use the measures introduced in Ref.~\cite{Holzgrafe2015}.
-We classify conformations $r$ as part of a fold if $Q\ge0.8$ where $Q\in [0,1]$ quantifies the similarity of $r$ to that fold.
-
-For folds A and B, $Q$ is determined by defining a set of backbone-backbone hydrogen bonds that are present in those folds (ignoring N and C terminal amino acids).
-For fold A, the fold-specific set of contacts (i,j) between the CO group of amino acid i and the NH group of amino acid j is $\{$(2,6), (3,7), (4,8), (5,9), (6,10), (7,11), (8,12), (9,13), (10,14), (11,15)$\}$, and for fold B the contact set is $\{$(3,14), (5,12), (7,10), (10,7), (12,5), (14,3)$\}$.
-A hydrogen bond is considered formed if (i) the NHO and COH angles both are $>90^\circ$ and (ii) the HO distance is $<2.7$~{\AA}.
-
-%\begin{equation}
-%Q_\mathrm{2\alpha} = ( Q_\alpha^\mathrm{1-16} + Q_\alpha^\mathrm{20-35} + Q_\mathrm{tert}) ) / 3
-%\end{equation}
-%
-%\begin{equation}
-%Q_\mathrm{\alpha\beta} = ( Q_\alpha^\mathrm{1-16} + Q_\beta^\mathrm{20-35} + Q_\mathrm{tert} ) / 3
-%\end{equation}
+Fold stabilities are calculated as in Ref.~\citenum{Holzgrafe2015} and also described briefly below.
+We define first two structural similarity measures $\QIA$ and $\QIB$ for folds IA and IB, respectively, indicating what fraction of the fold-specific backbone-backbone hydrogen bonds have been formed.
+The fold IA-hydrogen bonds are (2,6), (3,7), (4,8), (5,9), (6,10), (7,11), (8,12), (9,13), (10,14), (11,15) and the fold IB-bonds are (3,14), (5,12), (7,10), (10,7), (12,5), (14,3), where (i,j) indicates a hydrogen bond between the CO group of amino acid i and the NH group of amino acid j.
+The  stabilities of fold IA and IB are then defined as the probabilities $\PIA = P(\QIA\ge0.8)$ and $\PIB = P(\QIB\ge0.80)$, respectively, i.e., stability is the probability that at least 80\% of the fold's hydrogen bonds are formed.
+$\PIA$ and $\PIB$ depend on both sequence $\sbar$ and temperature $T$.
+For example, $\PIA=0.xx$ for A1 and $\PIB=0.xx$ for N1 at $\kb T = 0.43$.
+Structural similarity measures for 35-amino acid folds IIA and IIB are defined as $\QIIA = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IA}^\mathrm{20-35} + Q_\mathrm{tert}) ) / 3$ and $\QIIB = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IB}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$, respectively, where superscripts on $\QIA$ and $\QIB$ indicate over which amino acid positions those measures are applied to within the longer 35 amino acid sequences and $Q_\mathrm{tert}$ is a measure that counts the number of $C_\beta$-$C_\beta$ contacts between the two secondary structure elements of these folds.~\cite{Holzgrafe2015} As for IA and IB, the stabilities of folds IIA and IIB are defined as $\PIIA = P(\QIIA\ge0.8)$ and $\PIIB = P(\QIIB\ge0.80)$, respectively.
+%In calculating $Q$ values, a hydrogen bonßd is considered formed if (i) the NHO and COH angles are both $>90^\circ$ and (ii) the HO distance is $<2.7$~{\AA}.
 
@@
 Using MS simulations, we determined the thermodynamic behavior at the temperatures $\TlowI$ and $\TlowII$ for the A1-N1 and A2-TN systems, respectively (see Table~1).
-In particular, this allow us to find the stabilities of folds IA and IB ($\PaI$ and $\PbI$) for all sequences in {\SI} and the stabilities of folds IIA and IIB ($\PaII$ and $\PbII$) for all sequences in {\SII}.
+In particular, this allow us to find the stabilities of folds IA and IB ($\PIA$ and $\PIA$) for all sequences in {\SI} and the stabilities of folds IIA and IIB ($\PIIA$ and $\PIIB$) for all sequences in {\SII}.
 
@@
 To this end, we construct graphs in which each stable sequence is represented by a node and any two nodes are connected if their sequences differ at only one amino acid position.
-To determine if a sequence is stable we used the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PaI + \PbI$ and $\PaII + \PbII$ for A1-N1 and A2-AN1, respectively.
+To determine if a sequence is stable we used the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PIA + \PIB$ and $\PIIA + \PIIB$ for A1-N1 and A2-AN1, respectively.
 The precise network obtained depends, of course, on the cut-off value of $\Pcut$.
@@
 \noindent
-We have evaluated a generalized-ensemble algorithm for coarse-grained simulations of biomolecules that works by making the sequence a dynamic parameter and thereby leading to a random walk in sequence space.
-We find that this approach has two main benefits.
-On the one hand, it provides a convenient way to sample the respective canonical distributions for a set of sequences in a single run and, on the other, it also enhances the sampling of conformational space.
+We have tested a generalized-ensemble algorithm for coarse-grained simulations of biomolecules that works by making the sequence a dynamic parameter, thereby leading to a random walk in sequence space.
+This approach was evaluated for protein folding simulations by comparing with ST.
+The results indicate two main benefits of the MS approach.
+On the one hand, it provides a convenient way to sample the respective canonical distributions for a large set of sequences in a single run and, on the other, it also enhances the sampling of conformational space.
 
-The sampling efficiency of the MS algorithm can be assessed by the comparison with ST in Figure~3.
-At the highest studied temperatures, which are well above the folding temperatures of both A1 and N1, the statistical errors on the average total energy, $\sigE$, are very similar for the two methods.
+The conformational sampling efficiency of the MS algorithm can be assessed from the comparison with ST in Figure~3.
+Although there is no single ``fair" way to compare MS and ST, we chose as a measure of efficiency the estimated statistical error on the total energy, $\sigE$, obtained with roughly the same computational cost per temperature and sequence.
+At the highest studied temperatures, which are well above the folding temperatures of both A1 and N1, the statistical errors $\sigE$ are basically the same for the two methods.
 This is not unexpected because conformational sampling of short polymers at high $T$ does not involve crossing any major energy barriers.
-As a result, successively sampled conformations, for a given combination of sequence $\sbar$ and temperature $T$, are therefore likely mostly uncorrelated in both methods which in turn leads to similar $\sigE$ values.
+As a result, successively sampled conformations for a given combination of sequence $\sbar$ and temperature $T$, are therefore likely mostly uncorrelated in both methods leading to similar $\sigE$ values.
 
 At lower temperatures, we find that the MS simulations often yields significantly smaller $\sigE$ than ST simulations.
-It is notable that this acceleration in conformational sampling is achieved despite that simulations are carried out at constant temperature.
-Hence, rather than promoting escape from local minima in the free energy landscape by visits to higher temperature, as in ST~\cite{Marinari1992,Lyubartsev1992} or replica-exchange,~\cite{Swendsen1986} the MS algorithm promotes local minima escape through visits to other sequences for which the minima may not be as severe or are absent altogether.
-For the same reason, the efficiency of the MS algorithm likely depends on the size and character of sequences included in the simulation.
-In particular, the MS simulation is likely dependent on the inclusion of at least some less stable sequences to promote at least partial unfolding of the chain and the transition to different conformational states.
+It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant temperature.
+Hence, rather than promoting escape from local minima in the free energy landscape by visits to higher temperature, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences for which the minima may not be as deep or are altogether absent.
+For the same reason, the performance of the MS algorithm will depend on the size and character of sequences included.
+In particular, the MS simulation likely hinge on the inclusion of at least some sequences with poor stabilities such that at least partial unfolding of the chain is regularly triggered and promoting transitions to new conformational states.
 
-Clearly, the MS method should not be seen as a general simulation method to speed up conformational sampling due to the complexity of performing sequence updates in most biomolecular models.
+The MS method should not be seen as a general method to speed up conformational sampling due to the complexity of performing sequence updates in most biomolecular models.
 However, our results indicate that for coarse-grained models that permit the sequence updates as a Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be highly efficient.

### 2017-05-23 Manuscript.tex
changes220517

--- 2017-05-18 Manuscript.tex
@@
 Escape from local energy minima is achieved by visiting sequences for which the minima is shallower or absent.
-We apply the algorithm to a coarse-grained model for protein folding with 7 atoms per amino acid and apply it to sets of sequences with up to 35 amino acids spanning sequence space  With the multisequence algorithm, we show that the folding thermodynamics of $>$1,000 sequences at relatively low temperatures can be readily calculated.
-We use the resulting data to explore the statistics of pathways in sequence space between the two pairs of folds.
+We apply the algorithm alongside a coarse-grained model for protein folding with 7 atoms per amino acid to sets of sequences with up to 35 amino acids spanning a large sequence space.
+With the multisequence algorithm, we show that the folding thermodynamics of $>$ 1,000 sequences at relatively low temperatures can be readily calculated.
+We use the resulting data to explore the statistics of pathways in sequence space between two pairs of folds.
 
@@
 Moreover, large-scale screening techniques such as phage display and microarray chips are used to provide biophysical properties across entire genomes.
-In short, through the study of a biomolecular system and related variants molecular variants are studied in order to gain a more complete picture of the biophysical and biochemical properties of the system and provide evolutionary context.
+In short, through the study of a biomolecular system and related variants, molecular variants are studied in order to gain a more complete picture of the biophysical and biochemical properties of the system and provide evolutionary context.
 
@@
 We first apply the MS algorithm to a set of 144 model sequences with 16 amino acids and compare with the simulated tempering (ST) algorithm.
-We find that the computational efficiency of the MS algorithm is never less efficient than ST even at the lowest studied temperature.
-The model folds spontaneously designed sequences into stable native states with diverse protein-like sequences, and captures features such as mutational robustness.
-It therefore provide a biophysical basis for the sequence-structure and sequence-stability relationships of proteins at a conceptual level.
+We find that the computational efficiency of the MS algorithm is never less than ST except at the lowest studied temperature for one sequence.
+The model folds spontaneously designed sequences into stable native states with diverse protein-like structures, and captures features such as mutational robustness.
+It therefore provides a biophysical basis for the sequence-structure and sequence-stability relationships of proteins at a conceptual level.
 This model has been used, e.g., to study the sequence-structure relationship of proteins in the "marshlands" between protein folds.
-Experimental studies have indicated that single point mutations are capable to take change one protein fold into another.
-The multisequence algorithm allow us to characterize the stability of and structure of all 1,024 sequence in the binary space between the two folds.
+Experimental studies have indicated that single point mutations are capable of changing one protein fold into another.
+The multisequence algorithm allows us to characterize the stability and structure of all 1,024 sequences in the binary space between the two folds.
 The implication of these results for the fold switching and evolution is discussed.
@@
 
-While the 16-amino acid A1 and N1 sequences fold into $\alpha$-helical and $\beta$-hairpin folds, respectively, the two 35-amino acid sequences A2 and TA adopt the folds denoted IIA and IIB, as shown in Fig.~X.
+While the 16-amino acid A1 and N1 sequences fold into $\alpha$-helical and $\beta$-hairpin folds, respectively, the two 35-amino acid sequences A2 and TN adopt the folds denoted IIA and IIB, as shown in Fig.~X.
 
 The large number of sequences considered allows us to carry out a much more systematic analysis than previously~\cite{Holzgrafe2014,Holzgrafe2015} of the sequence-structure-stability relationship in the sequence region that connect these two folds pairs.
-We find that...
+We find that sequences become gradually less stable as they approach switch point.
+However, fold switching is observed to be typically abrupt at the switch point.
+Folding pathways usually pass through bistable sequences which populate both folds up to a significant level.
+We show that while bistable sequences become less when selecting more stable pathways for A1-N1 system, they generally remain the same for the A2-TN system.
 
@@
 \noindent
-Conventional Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy energy minima and hamper representative sampling of configurational space.
+Conventional Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
 The basic idea of generalized-ensemble algorithms~\cite{Mitsutake2001} is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters, such that a more efficient random walk in potential energy can be achieved.
@@
 and the first sum goes over a set of allowed sequences $\sbar$.
-The simulation parameters $h(\sbar)$, similarly to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=\tilde{Z}^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
+The simulation parameters $h(\sbar)$, similar to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=\tilde{Z}^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
 
@@
 All bond lengths, bond angles, and peptide plane angles (180$^\circ$) are held fixed.
-Hence, an $N$-amino acid chain conformation $\rbar$ can, for any sequence $\sbar$, is therefore described by the set of 2$N$ backbone torsional angles $\{\phi_i$, $\psi_i\}_{i=1}^{N}$.
+Hence, an $N$-amino acid chain conformation $\rbar$ can, for any sequence $\sbar$, therefore be described by the set of 2$N$ backbone torsional angles $\{\phi_i$, $\psi_i\}_{i=1}^{N}$.
 
@@
 Various model parameters, e.g., the strengths of hydrophobic attractions and hydrogen bonding, were determined based on the ability of the model to spontaneously fold a set of model sequences with 18-54 amino acids into  structurally diverse and thermodynamically stable native states with both $\beta$ and $\alpha$-structure.
-As it turned out, this strategy made the model robust enough to fold sequences designed to have mised $\alpha$ and $\beta$ structure.
+As it turned out, this strategy made the model robust enough to fold sequences designed to have mixed $\alpha$ and $\beta$ structures.
 
@@
 \begin{tabular}{lcr}
-Name & $N$ & Sequence\footnote{In addition, we study two sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs (see text) denoted $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$, respectively, and a restricted set $\mathrm{S16}_{144}$ with 144 sequences taken from S16 ($\mathrm{S16}_{144}$ was also studied in Ref.~\protect\citenum{Holzgrafe2014}).
+Name & $N$ & Sequence\footnote{In addition, we study two sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs (see text) denoted $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$, respectively, and a restricted set $\mathrm{S16}_{144}$ with 144 sequences taken from $\mathrm{S16}_{1024}$ ($\mathrm{S16}_{144}$ was also studied in Ref.~\protect\citenum{Holzgrafe2014}).
 } \\
@@
 N1 & 16 & phphphpttphphphp \\
-R1 & 16 & \textbf{(add)}\\
-R2 & 16 &  \textbf{(add)}\\
+R1 & 16 & pphhphptthpphhpp\\
+R2 & 16 & ppphphhtthhphppp\\
 A2 & 35 & (A1)ttt(A1)\\
@@
 \noindent
-Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assigned a new value between $-\pi$ and $\pi$; and (2) a semi-local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2001} In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out by randomly picking a new sequence $\sbar'\ne\sbar$ and applying the accept-reject criterion in Eq.~\ref{accrej}.
-A sequence move is attempted every $1,000$ MC step while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
+Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assign a new value between $-\pi$ and $\pi$; and (2) a semi-local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2001} In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out by randomly picking a new sequence $\sbar'\ne\sbar$ and applying the accept-reject criterion in Eq.~\ref{accrej}.
+A sequence move is attempted every $1,000$ MC steps while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
 The simulations carried out in this work are summarized in Tables~II.
@@
 \begin{tabular}{lccccr}
-Runs & Algorithm & $\kb T$  & MC steps &  Sequences\\
+Runs & Algorithm & $\kb T$  & MC steps/run &  Sequences\\
 \hline
@@
 For fold A, the fold-specific set of contacts (i,j) between the CO group of amino acid i and the NH group of amino acid j is $\{$(2,6), (3,7), (4,8), (5,9), (6,10), (7,11), (8,12), (9,13), (10,14), (11,15)$\}$, and for fold B the contact set is $\{$(3,14), (5,12), (7,10), (10,7), (12,5), (14,3)$\}$.
-A hydrogen bond is considered formed if (i) the NHO and COH angles both are $>90^\circ$ and (ii) the HO distance is $<2.7$~{\AA}.
+A hydrogen bond is considered formed if (i) the NHO and COH angles are both $>90^\circ$ and (ii) the HO distance is $<2.7$~{\AA}.
 
@@
 \begin{figure}
-\includegraphics[width=6.0cm]{MCtraj}
+\includegraphics[width=10.0cm]{MCTrajFolds}
 \caption{(A) Example simulation trajectory of the MS algorithm applied to the sequence set $\mathrm{S16}_{144}$.
-The plot shows the MC evolution of the sequence $\sbar$ (numbered 1--144), the total potential energy $E$ and the root-mean-square deviation (RMSD) calculated against the fold IA (light blue) and fold IB (dark red) structures shown in (B).
+The plot shows the MC evolution of the sequence $\sbar$ (numbered 1--144), the total potential energy $E$ and the root-mean-square deviation (RMSD) calculated against fold IA (light blue) and fold IB (dark red) structures shown in (B).
 The simulation is carried out at $\kb T = 0.43$ corresponding to the lowest studied temperature.
@@
 \includegraphics[width=8.0cm]{Stderr}
-\caption{Computational efficiency of the multisequence and simulated tempering algorithm.}
+\caption{Computational efficiency of the multisequence and simulated tempering algorithms for 4 sequences (A) A1, (B) N1, (C) R1 and (D) R2.}
 \end{figure}
 
-We apply the multisequence algorithm to this set of 144 sequences across a range of temperatures.
-Figure~2 illustrates a typical simulation carried out at the lowest temperature, $\Tlow$, which is well below the folding temperature of both A1 and N1~\cite{Holzgrafe2014,Holzgrafe2015}.
-From the MC evolution of the total energy $E$, sequence index j, and similarity measures $Q_A$ and $Q_B$, it is evident that visits to various sequences drive transitions into a range of structural states.
-For example, visits to A1 ($\mathrm{j}=1$) and N1 ($\mathrm{j}=144$) tend to coincide with visits to high-$Q_A$ and high-$Q_B$ conformations, respectively, as is necessary to generate the correct conformational ensembles for these sequences.
+We apply the multisequence algorithm to this set of 144 sequences across a range of temperatures (see Table~II).
+Figure~3 illustrates a typical simulation carried out at the lowest temperature, $\Tlow$, which is well below the folding temperature of both A1 and N1~\cite{Holzgrafe2014,Holzgrafe2015}.
+From the MC evolution of the total energy $E$, sequence index $\sbar$, and similarity measures $Q_A$ and $Q_B$, it is evident that visits to various sequences drive transitions into a range of structural states.
+For example, visits to A1 ($\sbar=1$) and N1 ($\sbar=144$) tend to coincide with visits to high-$Q_A$ and high-$Q_B$ conformations, respectively, as is necessary to generate the correct conformational ensembles for these sequences.
 
@@
 For example, an acceptance rate close to the oft-quoted rule-of-thumb value $0.25~\cite{Gilks1996} $ is obtained with $\Delta H_\mathrm{max}=4$, and with $\Delta H_\mathrm{max} = 1$ (only point mutations), $P_\mathrm{acc}=0.xx$.
-For simplicity, in this work, all simulations are carried without restrictions on how new sequences $\sbar'$ are drawn.
+For simplicity, in this work, all simulations are carried out without restrictions on how new sequences $\sbar'$ are drawn.
 
@@
 
-To assess computational efficiency, we compare with results from simulated tempering (ST) simulations carried out on 4 of the 144 sequences, namely A1 and N1 and the two sequences R1 and R2 chosen randomly at $H=4$ and $H=6$ from A1, respectively (see Table~1).
+In order to assess computational efficiency, we compare with results from simulated tempering (ST) simulations carried out on 4 of the 144 sequences, namely A1 and N1 and the two sequences R1 and R2 chosen randomly at $H=4$ and $H=6$ from A1, respectively (see Table~I).
 While ST provides the thermodynamics of a given sequence across the 8 temperatures $T_1, ..., T_8$ in a single run, a single MS simulation provides the thermodynamics of all 144 sequences for one $T_\mathrm{i}$.
@@
 
-Figure 4 shows the statistical error, $\sigE$, estimated for the average energy, $\left <E\right >$, for the 4 different sequences.
+Figure 4 shows the statistical error, $\sigE$, estimated for the average energy, $\left <E\right >$, for the 4 different sequences using ST and MS algorithms.
+Since an identical number of sampled conformations were obatined for each s and $T_i$ combination, we can therefore compare the statistical errors directly.
+Both MS and ST algorithms give the same small error at the highest T studied which implies consistent result between them.
+It should be noted that the highest T studied is well above the folding temperature for all the 4 sequences.
+Therefore, this consistent result is not particularly surprising since at high-T the free-energy landscape is relatively smooth.
+Hence, conformational space requires little or no difficulty to sample.
+
+There is a sequence-dependent statistical trend observed in Figure 6.
+As temperature decreases, the statistical error from the MS algorithm becomes smaller than those from ST.
+Even at the lowest T studied, below the folding temperature of all 4 sequences, we notice that the MS algorithm is at least consistent and most times better than ST in its estimates except for N1.
+This departure from the dominant trend is most likely due to statistical flunctuations, in which case, does not have any considerable significance.
+
+Insight from the statistical measurements suggests that the MS algorithm has a similar computational efficiency at $high_T$ and even better at lower T's.
+The MS algorithm is able to achieve  better efficency by sampling a wider region of both conformational and sequence space.
+Sampling the sequence space in addition to conformational space ensures a decorrelated ensemble of states is generated.
 
@@
 \rotatebox{0}{\includegraphics[width=12.6cm]{ntwk35}}
-\caption{Networks of the A1-N1 (top) and A2-TN (bottom) binary sequence space. }
+\caption{Networks of the A1-N1 (top) and A2-TN (bottom) binary sequence space showing the possible stable pathways from one fold to another. }
 \end{figure*}
@@
 \vspace{12pt}
-\subsection{A1-N1 and A2-TN binary sequence spaces}
+\subsection{A1-N1 and A2-TN binary sequence space}
 \noindent
-We now apply the MS algorithm to the sets {\SI} and {\SII} which represent the full binary sequence spaces for the A1-N1 and A2-TN systems, respectively.
-Both sets include $2^{10}=1,024$ sequences.
+We now apply the MS algorithm to the sets {\SI} and {\SII} which represent the full binary sequence space for the A1-N1 and A2-TN systems, respectively.
+Each set includes $2^{10}=1,024$ sequences.
 Using MS simulations, we determined the thermodynamic behavior at the temperatures $\TlowI$ and $\TlowII$ for the A1-N1 and A2-TN systems, respectively (see Table~1).
-In particular, this allow us to find the stabilities of folds IA and IB ($\PaI$ and $\PbI$) for all sequences in {\SI} and the stabilities of folds IIA and IIB ($\PaII$ and $\PbII$) for all sequences in {\SII}.
+In particular, this allows us to find the stabilities relative to folds IA and IB ($\PaI$ and $\PbI$) for all sequences in {\SI} and the stabilities relative to folds IIA and IIB ($\PaII$ and $\PbII$) for all sequences in {\SII}.
 
@@
 To this end, we construct graphs in which each stable sequence is represented by a node and any two nodes are connected if their sequences differ at only one amino acid position.
-To determine if a sequence is stable we used the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PaI + \PbI$ and $\PaII + \PbII$ for A1-N1 and A2-AN1, respectively.
+To determine if a sequence is stable we used the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PaI + \PbI$ and $\PaII + \PbII$ for A1-N1 and A2-TN, respectively.
 The precise network obtained depends, of course, on the cut-off value of $\Pcut$.
@@
 Fig.~5 shows the networks obtained with $\Pcut=0.50$.
-The number of viable mutational pathways that connect the start and end points are 516,972 between A1 and N1 and  57,912 between A2 and AN1.
+The number of viable mutational pathways that connect the start and end points are 516,972 between A1 and N1 and  57,912 between A2 and TN.
 Because there are $10!
-=3,628,800$ possible mutational pathways between start and end points in both cases, these pathways represent 14.2 \% and 1.6 \% of all possible pathways in A1-N1 and A2-AN1, respectively.
+=3,628,800$ possible mutational pathways between start and end points in both cases, these pathways represent 14.2 \% and 1.6 \% of all possible pathways in A1-N1 and A2-TN, respectively.
 Hence, A1 and N1 are rather highly connected at this stability threshold.
-For $\Pcut=0.60$, the numbers are 104,640 pathways (2.9\%) for A1-N1 and 22,512 (0.6\%) pathways for A2-AN1.
-We find that there are no possible A1-to-N1 or A2-to-AN1 pathways when $\Pcut\ge0.74$ and $\Pcut\ge 0.66$, respectively.
+For $\Pcut=0.60$, the numbers are 104,640 pathways (2.9\%) for A1-N1 and 22,512 (0.6\%) pathways for A2-TN.
+We find that there are no possible A1-to-N1 or A2-to-TN pathways when $\Pcut\ge0.74$ and $\Pcut\ge 0.66$, respectively.
 \\
@@
 \noindent 
-An apparently general characteristic of designed and natural protein folding switching is that the stability of sequences near the switch point have reduced stability.~\cite{Bryan2010} The results from our model exhibit a similar trend.
-Fig.~6A and B shows the total stability $\Ptot$ at different mutational distances between start and end points.
-Intermediate sequences are less stable than at distances $h=0$ and $h=10$, although there are large variations as indicated by the spread between upper and lower bounds.
-Nonetheless, there is a clear statistical trend that sequences become gradually less stable as successive mutations are applied to any of the 4 starting points until a minimum is reached.
+An apparently general characteristic of designed and natural protein folding switching is that sequences near the switch point have reduced stability.~\cite{Bryan2010} The results from our model exhibit a similar trend.
+Fig.~6A and B show the total stability $\Ptot$ at different mutational distances between start and end points.
+Intermediate sequences are less stable than sequences at distances $h=0$ and $h=10$, although there are large variations as indicated by the spread between upper and lower bounds.
+Nonetheless, there is a clear statistical trend that sequences become gradually less stable as successive mutations are applied to any of the 4 start and end points until a minimum is reached.
 %These minima in stability occur closer to N1 and TN in the two systems, respectively, indicating that the IA/IIA folds are more mutationally robust than the IB/IIB folds.
@@
 \includegraphics[width=7.8cm]{Paths}
-\caption{Averages stability as a function of Hamming distance $h$. }
+\caption{(Left) A1-N1 System.
+(Right) A2-TN System.
+(Top) Average stability as a function of Hamming distance $h$.
+(Middle) Fraction of pathways as a function of Swicth length $L_s$.
+(Bottom) Average stability as a function of Bistability $B$.}
 \end{figure}
@@
 
-Interestingly, fold switching through bistable sequences become less common as selection for more stable pathways is made in the A1-N1 system, as can be from the decrease in $\left <L_\mathrm{s}\right >$ as a function of $\Pcut$ (see Fig.~5C(inset)).
+Interestingly, fold switching through bistable sequences become less common as selection for more stable pathways is made in the A1-N1 system, as can be seen from the decrease in $\left <L_\mathrm{s}\right >$ as a function of $\Pcut$ (see Fig.~5C(inset)).
 For $\Pcut=0.70$, there are no longer any viable pathways between the $\alpha$-helix and $\beta$-hairpin folds that passes through a bistable point because $\left <L_\mathrm{s}\right > =  2$.
@@
 For this system, there are no direct $L_\mathrm{s}=2$ paths available even at $\Pcut=0.50$.
-Therefore, bistable sequence play a crucial role in bridging the folds IIA and IIB even through their stabilities are significantly reduced.
+Therefore, bistable sequences play a crucial role in bridging the folds IIA and IIB even though their stabilities are significantly reduced.
 
@@
 For comparison, we used the well-established generalized-ensemble ST algorithm and focused on 144 model sequences with 16 amino acids ({\SR}).
-As a measure of computational efficiency, we used on the statistical error $\sigE$ in the estimation of the average total energy $E$ for 4 different sequences which are part of {\SR}.
+As a measure of computational efficiency, we used the statistical error $\sigE$ in the estimation of the average total energy $E$ for 4 different sequences which are part of {\SR}.
 
@@
 
-A notable feature of the MS algorithm is that provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
-This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in lipid membranes, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Sampling in membrane simulations at high $T$, such as in replica-exchange simulations, are usually avoided in membrane simulations because it can lead to unwanted perturbation of the lipid bilayer structure although it can be a viable approach in particular cases.~\cite{Ulmschneider2010}
+A notable feature of the MS algorithm is that it provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
+This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in lipid membranes, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Sampling in membrane simulations at high $T$, such as in replica-exchange simulations, are usually avoided because it can lead to unwanted perturbation of the lipid bilayer structure although it can be a viable approach in particular cases.~\cite{Ulmschneider2010}
 
@@
 We have evaluated an algorithm for coarse-grained biomolecular simulations that allow the thermodynamics of multiple different sequences to be calculated in a single run.
-Here we showed that with this MS algorithm, the folding thermodynamics of over 1,000 amino acid sequences with each 35 amino acid could be determined in a coarse-grained model with 3 amino acid types.
-In addition to being a convenient for calculations of multiple sequences, the method enhances conformational sampling with an efficiency comparable to simulated tempering.
+Here we showed that with this MS algorithm, the folding thermodynamics of over 1,000 amino acid sequences with each 35 amino acids could be determined in a coarse-grained model with 3 amino acid types.
+In addition to being convenient for calculations of multiple sequences, the method enhances conformational sampling with an efficiency comparable to simulated tempering.
 Random walks across sequence space allow both sampling of many sequences and efficient escape from local free energy minima in the free energy landscape.
@@
 \end{document}
+\grid
+\grid

### 2017-05-23 Manuscript.tex

@@
 \begin{abstract}
-We evaluate a generalized-ensemble algorithm for coarse-grained simulations of biomolecules that allows the thermodynamics of multiple sequences to be calculated in a single ``multisequence" run.
-By carrying out a random walk in sequence space, conformational sampling is also enhanced.
-Escape from local energy minima is achieved by visiting sequences for which the minima is shallower or absent.
-We apply the algorithm alongside a coarse-grained model for protein folding with 7 atoms per amino acid to sets of sequences with up to 35 amino acids spanning a large sequence space.
-With the multisequence algorithm, we show that the folding thermodynamics of $>$ 1,000 sequences at relatively low temperatures can be readily calculated.
-We use the resulting data to explore the statistics of pathways in sequence space between two pairs of folds.
+
+We consider a generalized-ensemble algorithm for coarse-grained simulations of biomolecules, which allows the thermodynamic behaviors of two or more sequences to be determined in a single multisequence run.
+By carrying out a random walk in sequence space, the method also enhances conformational sampling.
+Escape from local energy minima is accelerated by visiting sequences for which the minima are more shallow or absent.
+We apply the method alongside an intermediate-resolution coarse-grained model for protein folding with 7 atoms per amino acids.
+The potential for large-scale coverage of sequence space is explored by applying the method to two sets with $>$1,000 sequences each.
+The resulting thermodynamic data is used to explore the sequence-structure and sequence-stability relationships between pairs of protein folds with different secondary structures.
+The potential implications for the evolution of new protein folds are discussed.
 
@@
 
-\subsection{Stability measures}
+\subsection{Observables}
 \noindent
-To quantify the stabilities of various folds, we use the measures introduced in Ref.~\cite{Holzgrafe2015}.
-We classify conformations $r$ as part of a fold if $Q\ge0.8$ where $Q\in [0,1]$ quantifies the similarity of $r$ to that fold.
-
-For folds A and B, $Q$ is determined by defining a set of backbone-backbone hydrogen bonds that are present in those folds (ignoring N and C terminal amino acids).
-For fold A, the fold-specific set of contacts (i,j) between the CO group of amino acid i and the NH group of amino acid j is $\{$(2,6), (3,7), (4,8), (5,9), (6,10), (7,11), (8,12), (9,13), (10,14), (11,15)$\}$, and for fold B the contact set is $\{$(3,14), (5,12), (7,10), (10,7), (12,5), (14,3)$\}$.
-A hydrogen bond is considered formed if (i) the NHO and COH angles are both $>90^\circ$ and (ii) the HO distance is $<2.7$~{\AA}.
-
-%\begin{equation}
-%Q_\mathrm{2\alpha} = ( Q_\alpha^\mathrm{1-16} + Q_\alpha^\mathrm{20-35} + Q_\mathrm{tert}) ) / 3
-%\end{equation}
-%
-%\begin{equation}
-%Q_\mathrm{\alpha\beta} = ( Q_\alpha^\mathrm{1-16} + Q_\beta^\mathrm{20-35} + Q_\mathrm{tert} ) / 3
-%\end{equation}
+Fold stabilities are calculated as in Ref.~\citenum{Holzgrafe2015} and also described briefly below.
+We define first two structural similarity measures $\QIA$ and $\QIB$ for folds IA and IB, respectively, indicating what fraction of the fold-specific backbone-backbone hydrogen bonds have been formed.
+The fold IA-hydrogen bonds are (2,6), (3,7), (4,8), (5,9), (6,10), (7,11), (8,12), (9,13), (10,14), (11,15) and the fold IB-bonds are (3,14), (5,12), (7,10), (10,7), (12,5), (14,3), where (i,j) indicates a hydrogen bond between the CO group of amino acid i and the NH group of amino acid j.
+The  stabilities of fold IA and IB are then defined as the probabilities $\PIA = P(\QIA\ge0.8)$ and $\PIB = P(\QIB\ge0.80)$, respectively, i.e., stability is the probability that at least 80\% of the fold's hydrogen bonds are formed.
+$\PIA$ and $\PIB$ depend on both sequence $\sbar$ and temperature $T$.
+For example, $\PIA=0.xx$ for A1 and $\PIB=0.xx$ for N1 at $\kb T = 0.43$.
+Structural similarity measures for 35-amino acid folds IIA and IIB are defined as $\QIIA = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IA}^\mathrm{20-35} + Q_\mathrm{tert}) ) / 3$ and $\QIIB = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IB}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$, respectively, where superscripts on $\QIA$ and $\QIB$ indicate over which amino acid positions those measures are applied to within the longer 35 amino acid sequences and $Q_\mathrm{tert}$ is a measure that counts the number of $C_\beta$-$C_\beta$ contacts between the two secondary structure elements of these folds.~\cite{Holzgrafe2015} As for IA and IB, the stabilities of folds IIA and IIB are defined as $\PIIA = P(\QIIA\ge0.8)$ and $\PIIB = P(\QIIB\ge0.80)$, respectively.
+%In calculating $Q$ values, a hydrogen bond is considered formed if (i) the NHO and COH angles are both $>90^\circ$ and (ii) the HO distance is $<2.7$~{\AA}.
 
@@
 \noindent
+
 We now apply the MS algorithm to the sets {\SI} and {\SII} which represent the full binary sequence space for the A1-N1 and A2-TN systems, respectively.
@@
 Using MS simulations, we determined the thermodynamic behavior at the temperatures $\TlowI$ and $\TlowII$ for the A1-N1 and A2-TN systems, respectively (see Table~1).
-In particular, this allows us to find the stabilities relative to folds IA and IB ($\PaI$ and $\PbI$) for all sequences in {\SI} and the stabilities relative to folds IIA and IIB ($\PaII$ and $\PbII$) for all sequences in {\SII}.
+In particular, this allows us to find the stabilities relative to folds IA and IB ($\PIA$ and $\PIB$) for all sequences in {\SI} and the stabilities relative to folds IIA and IIB ($\PIIA$ and $\PIIB$) for all sequences in {\SII}.
 
@@
 To this end, we construct graphs in which each stable sequence is represented by a node and any two nodes are connected if their sequences differ at only one amino acid position.
-To determine if a sequence is stable we used the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PaI + \PbI$ and $\PaII + \PbII$ for A1-N1 and A2-TN, respectively.
+To determine if a sequence is stable we used the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PIA + \PIB$ and $\PIIA + \PIIB$ for A1-N1 and A2-TN, respectively.
 The precise network obtained depends, of course, on the cut-off value of $\Pcut$.
@@
 \noindent
-We have tested an algorithm for coarse-grained biomolecular simulations that relies on making the sequence of the molecule a dynamical parameter.
+We have tested a generalized-ensemble algorithm for coarse-grained simulations of biomolecules that works by making the sequence a dynamic parameter, thereby leading to a random walk in sequence space.
+This approach was evaluated for protein folding simulations by comparing with ST.
+The results indicate two main benefits of the MS approach.
+On the one hand, it provides a convenient way to sample the respective canonical distributions for a large set of sequences in a single run and, on the other, it also enhances the sampling of conformational space.
 
-For comparison, we used the well-established generalized-ensemble ST algorithm and focused on 144 model sequences with 16 amino acids ({\SR}).
+The conformational sampling efficiency of the MS algorithm can be assessed from the comparison with ST in Figure~3.
+Although there is no single ``fair" way to compare MS and ST, we chose as a measure of efficiency the estimated statistical error on the total energy, $\sigE$, obtained with roughly the same computational cost per temperature and sequence.
+At the highest studied temperatures, which are well above the folding temperatures of both A1 and N1, the statistical errors $\sigE$ are basically the same for the two methods.
+This is not unexpected because conformational sampling of short polymers at high $T$ does not involve crossing any major energy barriers.
+As a result, successively sampled conformations for a given combination of sequence $\sbar$ and temperature $T$, are therefore likely mostly uncorrelated in both methods leading to similar $\sigE$ values.
+
+At lower temperatures, we find that the MS simulations often yields significantly smaller $\sigE$ than ST simulations.
+It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant temperature.
+Hence, rather than promoting escape from local minima in the free energy landscape by visits to higher temperature, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences for which the minima may not be as deep or are altogether absent.
+For the same reason, the performance of the MS algorithm will depend on the size and character of sequences included.
+In particular, the MS simulation likely hinge on the inclusion of at least some sequences with poor stabilities such that at least partial unfolding of the chain is regularly triggered and promoting transitions to new conformational states.
+
+The MS method should not be seen as a general method to speed up conformational sampling due to the complexity of performing sequence updates in most biomolecular models.
+However, our results indicate that for coarse-grained models that permit the sequence updates as a Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be highly efficient.
+A notable feature of the algorithm is that it provides escape from local minima on the free energy landscape of proteins without resorting to a change in $T$.
+This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in the context or ordered dense phases such as lipid bilayers, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Simulations of lipid bilayers at elevated $T$ must typically be avoided because it can lead to unwanted perturbations of the membrane structure, although it can be a viable approach to speed up conformational sampling in particular cases.~\cite{Ulmschneider2010}
+
+%Overall our results show that, when technically feasible, a multisequence approach can be highly efficient for calculating the thermodynamics of a large number of amino acid sequences even at low temperatures.
+To demonstrate that the method to large sequence sets we applied it to the full binary sequences spaces between the model sequence pairs A1-N1 and A2-TN, which both include 1,024 sequences.
+These calculations would have been difficult to carry out ``in series" using some other standard method such as simulated tempering.
+
+%For comparison, we used the well-established generalized-ensemble ST algorithm and focused on 144 model sequences with 16 amino acids ({\SR}).
 As a measure of computational efficiency, we used the statistical error $\sigE$ in the estimation of the average total energy $E$ for 4 different sequences which are part of {\SR}.
 
-Our results show that at high $T$, well above the folding temperatures of both A1 and N1, we find that the statistical errors $\sigE$ for ST and MS are small and very similar.
-This is not surprising given that conformational sampling of relatively short polymer chains at high $T$ does not involve crossing major energy barriers.
-Therefore, successive sampled conformational states, for a given combination of sequence $\sbar$ and temperature $T$, are likely uncorrelated in both ST and MS simulations which, in turn, leads to similar estimates of $\Eave$ (and $\sigE$).
-At low $T$,...
-
-Hence, we conclude that the computational efficiency of the MS algorithm is comparable to ST.
+%Hence, we conclude that the computational efficiency of the MS algorithm is comparable to ST.
 However, that while the ST simulation requires that the thermodynamic behavior is calculated at a range of temperatures, MS works well on its own directly at low temperatures.
@@
 
-A notable feature of the MS algorithm is that it provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
+%A notable feature of the MS algorithm is that provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
+This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in lipid membranes, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Sampling in membrane simulations at high $T$, such as in replica-exchange simulations, are usually avoided in membrane simulations because it can lead to unwanted perturbation of the lipid bilayer structure although it can be a viable approach in particular cases.~\cite{Ulmschneider2010}
+
+%A notable feature of the MS algorithm is that it provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
 This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in lipid membranes, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Sampling in membrane simulations at high $T$, such as in replica-exchange simulations, are usually avoided because it can lead to unwanted perturbation of the lipid bilayer structure although it can be a viable approach in particular cases.~\cite{Ulmschneider2010}

### 2017-05-23 Manuscript.tex
Introduction

--- 2017-05-22 Manuscript.tex
@@
 \noindent
+A large number of coarse-grained models for biomolecular simulations have been developed the last decade.~\cite{Ingolfsson2014} Several characteristics of such models goal of such models is to speed up simulations through a reduced number of degrees-of-freedom,
+
+In this situation, it is of interest to explore strategies that can simulate multiple sequence variants in an efficient manner.
+
 Modern investigations into biophysical properties of biomolecular systems are routinely carried out not only on a single species but on multiple molecular variants.~\cite{Vukmirovic2000,Nickson2010} For example, protein engineering methods are used to provide mechanistic insight into the specificity of protein-protein and protein-DNA interactions, to compare the  conformational propensities of disease-related mutations of aggregation-prone proteins, and optimize biomolecular properties through directed evolution.

### 2017-05-23 Manuscript.tex

--- 2017-05-23 Manuscript.tex
@@
 \noindent
+A large number of coarse-grained models for biomolecular simulations have been developed the last decade.~\cite{Ingolfsson2014} Several characteristics of such models goal of such models is to speed up simulations through a reduced number of degrees-of-freedom,
+
+In this situation, it is of interest to explore strategies that can simulate multiple sequence variants in an efficient manner.
+
 Modern investigations into biophysical properties of biomolecular systems are routinely carried out not only on a single species but on multiple molecular variants.~\cite{Vukmirovic2000,Nickson2010} For example, protein engineering methods are used to provide mechanistic insight into the specificity of protein-protein and protein-DNA interactions, to compare the  conformational propensities of disease-related mutations of aggregation-prone proteins, and optimize biomolecular properties through directed evolution.
 Moreover, large-scale screening techniques such as phage display and microarray chips are used to provide biophysical properties across entire genomes.
-In short, through the study of a biomolecular system and related variants, molecular variants are studied in order to gain a more complete picture of the biophysical and biochemical properties of the system and provide evolutionary context.
+In short, through the study of a biomolecular system and related variants molecular variants are studied in order to gain a more complete picture of the biophysical and biochemical properties of the system and provide evolutionary context.
 

### 2017-05-29 Manuscript.tex
TableI and TableII in Manuscript.tex Modified. MCTracFolds.png, network.png, Stderr.pdf, meanE.pdf andSupplementary.tex created. Labels also modified on graphs

--- 2017-05-23 Manuscript.tex
@@
 \begin{equation}
-P(k) = \frac{1}{\hat{Z}}\sum_r e^{-\beta_m E(\rbar) + g_m} \,,
+P(m) = \frac{1}{\hat{Z}}\sum_r e^{-\beta_m E(\rbar) + g_m} \,,
 \end{equation}
@@
 A common and convenient choice is $g_m\approx \beta_m F_m$, where $F_m$ is the free energy at temperature $T_m$.
-With this choice, $P(k)$ becomes approximately flat ensuring all temperatures are frequently visited.
+With this choice, $P(m)$ becomes approximately flat ensuring all temperatures are frequently visited.
 
@@
 
-The sequences studied in this work are summarized in Table~I.
+Six model sequences studied in this work are summarized in Table~I.
+In addition, we study two sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs (see text) denoted $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$, respectively, and a restricted set $\mathrm{S16}_{144}$ with 144 sequences taken from $\mathrm{S16}_{1024}$ ($\mathrm{S16}_{144}$ was also studied in Ref.~\protect\citenum{Holzgrafe2014}).
 
 \begin{table}
-\caption{\label{tab1} List of model sequences of different lengths $N$ studied in this work.}
+\caption{\label{tab1} List of 6 model sequences of different lengths $N$ studied in this work.}
 \begin{ruledtabular}
 \begin{tabular}{lcr}
-Name & $N$ & Sequence\footnote{In addition, we study two sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs (see text) denoted $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$, respectively, and a restricted set $\mathrm{S16}_{144}$ with 144 sequences taken from $\mathrm{S16}_{1024}$ ($\mathrm{S16}_{144}$ was also studied in Ref.~\protect\citenum{Holzgrafe2014}).
-} \\
+Name & $N$ & Sequence \\
 \hline
@@
 \begin{tabular}{lccccr}
-Runs & Algorithm & $\kb T$  & MC steps/run &  Sequences\\
+Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes thermalization with $10^6$ MC steps.} &  Sequences\\
 \hline
-32 & ST & 0.43--0.65 & $8\times 10^6$ &A1\\ 
-32 & ST & 0.43--0.65 & $8\times 10^6$ &N1\\ 
-32 & ST & 0.43--0.65 & $8\times 10^6$ &R1\\ 
-32 & ST & 0.43--0.65 & $8\times 10^6$ &R2\\ 
-32$\times$8\footnote{32  runs per temperature for 8 temperatures in the range given.} & MS &0.43--0.65& $144\times 10^6$ & $\mathrm{S16}_{144}$\\
+32 & ST & 0.43--0.65 & $10^7$ &A1\\ 
+32 & ST & 0.43--0.65 & $10^7$ &N1\\ 
+32 & ST & 0.43--0.65 & $10^7$ &R1\\ 
+32 & ST & 0.43--0.65 & $10^7$ &R2\\ 
+32$\times$8\footnote{32  runs per temperature for 8 temperatures in the range given.} & MS &0.43--0.65& $18\times 10^7$ & $\mathrm{S16}_{144}$\\
 16 & MS & 0.43  & $5\times 10^9$ &  $\mathrm{S16}_{1024}$ \\
@@
 \begin{figure*}
-\rotatebox{0}{\includegraphics[width=16.6cm]{ntwk16}}\\
-\rotatebox{0}{\includegraphics[width=12.6cm]{ntwk35}}
-\caption{Networks of the A1-N1 (top) and A2-TN (bottom) binary sequence space showing the possible stable pathways from one fold to another. }
+\rotatebox{0}{\includegraphics[width=16.6cm]{network}}
+\caption{Networks of the A1-N1 (top) and A2-TN (bottom) binary sequence space showing the possible stable pathways from one fold to another.
+(light blue) Fold A, (dark red) Fold B, and (black) Bistable.
+}
 \end{figure*}

### 2017-05-29 Supplement.tex
TableI and TableII in Manuscript.tex Modified. MCTracFolds.png, network.png, Stderr.pdf, meanE.pdf andSupplementary.tex created. Labels also modified on graphs

\section{Supplementary Material}

\begin{figure}
\includegraphics[width=14cm]{meanE}
\caption{ Correlation of average energy as calculated by ST and MS algorithms for our 4 model sequences (A) A1, (B) N1, (C) R1, and (D) R2. Figure shows that  ST and MS are consistent in their calculations. The errors associated with both algorithms are very small such that they are not visible on the scale of the average energy.}
\end{figure}

### 2017-05-29 Manuscript.tex
Intro+abstract+Paths.pdf

--- 2017-05-23 Manuscript.tex
@@
 % Use the file aiptemplate.tex as a template for your document.
-\documentclass[%
+\documentclass[
 aip,
@@
 \noindent
-A large number of coarse-grained models for biomolecular simulations have been developed the last decade.~\cite{Ingolfsson2014} Several characteristics of such models goal of such models is to speed up simulations through a reduced number of degrees-of-freedom,
+Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} These advances have in part triggered major efforts~\cite{McGuffee2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the molecular dynamics of bio-systems of different sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} or a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging, these large-scale simulations also highlight the fact that severe tradeoffs in size and time scales will likely persist for the foreseeable future.
 
-In this situation, it is of interest to explore strategies that can simulate multiple sequence variants in an efficient manner.
+One way to expand the range of biomolecular simulations is to turn to coarse-grained models, with the basic aim to simplify the physical description of interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list four main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that CG representations often open up for alternative sampling schemes beyond traditional molecular dynamics approaches which can further speed up conformational sampling, e.g., activation-relaxation kinetics,~\cite{Beland2011} discrete molecular dynamics~\cite{Proctor2011} and various Monte Carlo-based strategies such as cluster moves.~\cite{Vitalis2009}
 
-Modern investigations into biophysical properties of biomolecular systems are routinely carried out not only on a single species but on multiple molecular variants.~\cite{Vukmirovic2000,Nickson2010} For example, protein engineering methods are used to provide mechanistic insight into the specificity of protein-protein and protein-DNA interactions, to compare the  conformational propensities of disease-related mutations of aggregation-prone proteins, and optimize biomolecular properties through directed evolution.
+The challenges of simulating individual biomolecular systems notwithstanding, a range of systems naturally call for the investigation and comparison of many molecular variants, e.g., understanding the molecular mechanisms of specificity in protein-protein, peptide-protein,~\cite{Zarrinpar2003} or protein-nucleotide interactions, or the role of mutations in disease processes such as protein aggregation.
+Another example is protein folding where unique insight has been achieved by comparing the folding within and between protein families.~\cite{Tzul2017,Wensley2010}
+
+In this work, we evaluate a Monte Carlo (MC)-based algorithm that can in principle calculate the thermodynamics of multiple sequences in a single run and apply it to a coarse-grained model for protein folding.
+This ``multisequence" (MS) method was originally developed in the context of homo- and heteropolymer simulations~\cite{Irback1995} and was later adapted for the characterization of peptide-protein binding specificity,~\cite{Bhattacherjee2013,Wallin2017}, but it has to our knowledge not been previously tested in protein folding simulations.
+The MS algorithm relies on sampling at fixed temperature a generalized ensemble in which the sequence is a dynamic parameter.
+Hence, there are two main types of updates: conformational updates $\rbar\rightarrow\rbar'$ and sequences updates $\sbar\rightarrow\sbar'$.
+As illustrated in Fig.~1, this approach requires that $\rbar$ and $\sbar$ are ``perpendicular" coordinates such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$.
+While this may be hard to achieve for detailed protein models, it is likely possible in many CG models.
+
+As an initial test case, we apply the MS method to a set of 144 model sequences with 16 amino acids.
+This set, {\SR}, was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and $\beta$-hairpin, respectively, and allowed us to demonstrate the existence of A1-N1 mutational pathways which abruptly switch between the two structures and do not proceed through unstable sequences.~\cite{Holzgrafe2014} Here we use this set to validate the MS method and compare its efficiency to a standard generalized-ensemble methods, namely simulated tempering (ST).~\cite{Marinari1992,Lyubartsev1992} We thereafter apply the MS method to a greatly enlarged set with 1,024 sequences, also spanning the A1-N1 space, as well as another set spanning two 35-amino acid sequences, A2 and TN, folding into a two-helical bundle fold and a mixed $\alpha$-$\beta$ fold, respectively.
+Besides hinting at the feasibility of applying the MS method on large sets of sequences, the results allow us to carry out a more systematic analysis of the biophysical properties of sequences along mutational pathways connecting these two pairs of basic protein folds than has been possible with standard methods.~\cite{Holzgrafe2014,Holzgrafe2015}
+
+%We find that sequences become gradually less stable as they approach switch point.
+However, fold switching is observed to be typically abrupt at the switch point.
+Folding pathways usually pass through bistable sequences which populate both folds up to a significant level.
+We show that while bistable sequences become less when selecting more stable pathways for A1-N1 system, they generally remain the same for the A2-TN system.
+
+%We first apply the MS algorithm to a set of 144 model sequences with 16 amino acids and compare with the simulated tempering (ST) algorithm.
+We find that the computational efficiency of the MS algorithm is never less than ST except at the lowest studied temperature for one sequence.
+
+%We then demonstrate that the multisequence algorithm can be used to calculate the thermodynamic behavior for $>1,000$ sequences in a 3-letter model for protein folding~\cite{Bhattacherjee2012} at temperatures well below the typical folding temperatures $T_\mathrm{f}$ of the included sequences.
+
+%Modern investigations into biophysical properties of biomolecular systems are routinely carried out not only on a single species but on multiple molecular variants.~\cite{Vukmirovic2000,Nickson2010} For example, protein engineering methods are used to provide mechanistic insight into the specificity of protein-protein and protein-DNA interactions, to compare the  conformational propensities of disease-related mutations of aggregation-prone proteins, and optimize biomolecular properties through directed evolution.
 Moreover, large-scale screening techniques such as phage display and microarray chips are used to provide biophysical properties across entire genomes.
@@
 
-While protein or DNA sequences are easily changed on the computer, the high computational cost of simulating even a single moderately-sized biomolecule makes systematic explorations of sequence space challenging.
+%While protein or DNA sequences are easily changed on the computer, the high computational cost of simulating even a single moderately-sized biomolecule makes systematic explorations of sequence space challenging.
 Some notable efforts have nonetheless been attempted, including simulations of all 136 unique tetranucleotide sequences in B-DNA oligonucleotides using all-atom explicit water molecular dynamics.~\cite{Beveridge2004}
 
-Here we explore a Monte Carlo (MC)-based algorithm for calculating the folding thermodynamics of multiple protein sequences in a single enhanced run.
-The algorithm relies on sampling at fixed temperature a generalized ensemble in which the sequence is treated as a dynamic parameter.
-Hence, this multisequence MC method includes two main types of updates: conformational updates $\rbar\rightarrow\rbar'$ and sequences updates $\sbar\rightarrow\sbar'$.
-As illustrated in Fig.~1, this approach requires that $\rbar$ and $\sbar$ are ``perpendicular" coordinates such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$.
-While this may be hard to achieve for detailed protein models, it is likely possible for many coarse-grained models.
-
-We first apply the MS algorithm to a set of 144 model sequences with 16 amino acids and compare with the simulated tempering (ST) algorithm.
-We find that the computational efficiency of the MS algorithm is never less than ST except at the lowest studied temperature for one sequence.
-The model folds spontaneously designed sequences into stable native states with diverse protein-like structures, and captures features such as mutational robustness.
-It therefore provides a biophysical basis for the sequence-structure and sequence-stability relationships of proteins at a conceptual level.
-This model has been used, e.g., to study the sequence-structure relationship of proteins in the "marshlands" between protein folds.
-Experimental studies have indicated that single point mutations are capable of changing one protein fold into another.
-The multisequence algorithm allows us to characterize the stability and structure of all 1,024 sequences in the binary space between the two folds.
-The implication of these results for the fold switching and evolution is discussed.
-\\
-
-We then demonstrate that the multisequence algorithm can be used to calculate the thermodynamic behavior for $>1,000$ sequences in a 3-letter model for protein folding~\cite{Bhattacherjee2012} at temperatures well below the typical folding temperatures $T_\mathrm{f}$ of the included sequences.
-
-While the 16-amino acid A1 and N1 sequences fold into $\alpha$-helical and $\beta$-hairpin folds, respectively, the two 35-amino acid sequences A2 and TN adopt the folds denoted IIA and IIB, as shown in Fig.~X.
-
-The large number of sequences considered allows us to carry out a much more systematic analysis than previously~\cite{Holzgrafe2014,Holzgrafe2015} of the sequence-structure-stability relationship in the sequence region that connect these two folds pairs.
-We find that sequences become gradually less stable as they approach switch point.
-However, fold switching is observed to be typically abrupt at the switch point.
-Folding pathways usually pass through bistable sequences which populate both folds up to a significant level.
-We show that while bistable sequences become less when selecting more stable pathways for A1-N1 system, they generally remain the same for the A2-TN system.
+%While the 16-amino acid A1 and N1 sequences fold into $\alpha$-helical and $\beta$-hairpin folds, respectively, the two 35-amino acid sequences A2 and TN adopt the folds denoted IIA and IIB, as shown in Fig.~X.
 
@@
 \includegraphics[width=8.0cm]{Stderr}
-\caption{Computational efficiency of the multisequence and simulated tempering algorithms for 4 sequences (A) A1, (B) N1, (C) R1 and (D) R2.}
+\caption{Comparing MS and ST algorithms.
+The statistical errors $\sigma_{\left < E\right >}$ of the average total energy $\left < E\right >$ are estimated with the two methods for the sequences (A) A1, (B) N1, (C) R1 and (D) R2.
+Simulation lengths are adjusted such that the number of conformations sampled per sequence and temperature is similar between in MS and ST runs.
+}
 \end{figure}

### 2017-06-01 Manuscript.tex
Miscellaneous changes

@@
 \newcommand {\SI}		{${\mathrm{S16}_{1024}}$}	
-\newcommand {\SII}		{${\mathrm{S16}_{1024}}$}
+\newcommand {\SII}		{${\mathrm{S35}_{1024}}$}
 
@@
 \begin{abstract}
-
-We consider a generalized-ensemble algorithm for coarse-grained simulations of biomolecules, which allows the thermodynamic behaviors of two or more sequences to be determined in a single multisequence run.
+We consider a generalized-ensemble algorithm for coarse-grained simulations of biomolecules which allows the thermodynamic behavior of two or more sequences to be determined in a single ``multisequence" run.
 By carrying out a random walk in sequence space, the method also enhances conformational sampling.
 Escape from local energy minima is accelerated by visiting sequences for which the minima are more shallow or absent.
-We apply the method alongside an intermediate-resolution coarse-grained model for protein folding with 7 atoms per amino acids.
+We apply the method alongside an intermediate-resolution coarse-grained model for protein folding with 7 atoms per amino acid and 3 amino acid types.
 The potential for large-scale coverage of sequence space is explored by applying the method to two sets with $>$1,000 sequences each.
 The resulting thermodynamic data is used to explore the sequence-structure and sequence-stability relationships between pairs of protein folds with different secondary structures.
-The potential implications for the evolution of new protein folds are discussed.
-
 \end{abstract}
 
-\pacs{Add  here}
+\pacs{87.14.E; 87.15.A; 05.10.Ln}
 
-\keywords{Generalized ensembles, simulated tempering, protein folding, coarse-graining, evolution}
+\keywords{Monte Carlo, generalized ensembles, simulated tempering, protein folding, protein evolution}
 
@@
 \noindent
-Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} These advances have in part triggered major efforts~\cite{McGuffee2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the molecular dynamics of bio-systems of different sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} or a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging, these large-scale simulations also highlight the fact that severe tradeoffs in size and time scales will likely persist for the foreseeable future.
+Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} These advances have in part triggered major efforts~\cite{McGuffee2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the molecular dynamics of bio-systems of different sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} or a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging and insightful, these large-scale simulations have also highlighted the fact that severe tradeoffs in size and time scales will likely persist for the foreseeable future for detailed simulations.
 
-One way to expand the range of biomolecular simulations is to turn to coarse-grained models, with the basic aim to simplify the physical description of interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list four main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that CG representations often open up for alternative sampling schemes beyond traditional molecular dynamics approaches which can further speed up conformational sampling, e.g., activation-relaxation kinetics,~\cite{Beland2011} discrete molecular dynamics~\cite{Proctor2011} and various Monte Carlo-based strategies such as cluster moves.~\cite{Vitalis2009}
+One way to expand the range of biomolecular simulations is to turn to coarse-grained (CG) models, where the basic aim is to simplify the physical description of interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list 4 main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that a CG representation either of the interaction potential or the molecular geometry often opens up for alternative sampling schemes beyond traditional molecular dynamics approaches which can further speed up conformational sampling, e.g., activation-relaxation kinetics,~\cite{Beland2011} discrete molecular dynamics~\cite{Proctor2011} and various Monte Carlo (MC)-based techniques such as cluster moves.~\cite{Vitalis2009}
 
-The challenges of simulating individual biomolecular systems notwithstanding, a range of systems naturally call for the investigation and comparison of many molecular variants, e.g., understanding the molecular mechanisms of specificity in protein-protein, peptide-protein,~\cite{Zarrinpar2003} or protein-nucleotide interactions, or the role of mutations in disease processes such as protein aggregation.
-Another example is protein folding where unique insight has been achieved by comparing the folding within and between protein families.~\cite{Tzul2017,Wensley2010}
+The challenges of achieving representative conformational sampling of individual biomolecular systems notwithstanding, a range of systems naturally call for the investigation and comparison of many molecular variants, e.g., determining the molecular mechanisms of specificity in protein-protein~\cite{Zarrinpar2003,Hakes2007} or protein-nucleotide interactions,~\cite{Rohs2010} or the role of mutations in disease processes such as protein aggregation.~\cite{Ross2004} Another example is protein folding where unique insight has been achieved by comparing the folding within and between protein families.~\cite{Tzul2017,Wensley2010} With the increasing availability of sequence information,~\cite{Vukmirovic2000} it is of interest to explore ways to efficiently sample multiple sequences in biomolecular simulations.
 
-In this work, we evaluate a Monte Carlo (MC)-based algorithm that can in principle calculate the thermodynamics of multiple sequences in a single run and apply it to a coarse-grained model for protein folding.
+To this end, we evaluate in this work an MC-based algorithm that can in principle calculate the thermodynamics of multiple sequences in a single run and apply it to a coarse-grained model for protein folding.
 This ``multisequence" (MS) method was originally developed in the context of homo- and heteropolymer simulations~\cite{Irback1995} and was later adapted for the characterization of peptide-protein binding specificity,~\cite{Bhattacherjee2013,Wallin2017}, but it has to our knowledge not been previously tested in protein folding simulations.
 The MS algorithm relies on sampling at fixed temperature a generalized ensemble in which the sequence is a dynamic parameter.
-Hence, there are two main types of updates: conformational updates $\rbar\rightarrow\rbar'$ and sequences updates $\sbar\rightarrow\sbar'$.
-As illustrated in Fig.~1, this approach requires that $\rbar$ and $\sbar$ are ``perpendicular" coordinates such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$.
+Hence, there are two main types of updates: conformational updates $\rbar\rightarrow\rbar'$ and sequence updates $\sbar\rightarrow\sbar'$.
+As illustrated in Fig.~1, this approach is easily applied when $\rbar$ and $\sbar$ are ``perpendicular" coordinates such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$.
 While this may be hard to achieve for detailed protein models, it is likely possible in many CG models.
@@
 As an initial test case, we apply the MS method to a set of 144 model sequences with 16 amino acids.
-This set, {\SR}, was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and $\beta$-hairpin, respectively, and allowed us to demonstrate the existence of A1-N1 mutational pathways which abruptly switch between the two structures and do not proceed through unstable sequences.~\cite{Holzgrafe2014} Here we use this set to validate the MS method and compare its efficiency to a standard generalized-ensemble methods, namely simulated tempering (ST).~\cite{Marinari1992,Lyubartsev1992} We thereafter apply the MS method to a greatly enlarged set with 1,024 sequences, also spanning the A1-N1 space, as well as another set spanning two 35-amino acid sequences, A2 and TN, folding into a two-helical bundle fold and a mixed $\alpha$-$\beta$ fold, respectively.
+This set, {\SR}, was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and $\beta$-hairpin, respectively, and allowed us to demonstrate the existence of A1-N1 mutational pathways which abruptly switch between the two structures and do not proceed through unstable sequences.~\cite{Holzgrafe2014} Here we use this set to validate the MS method and compare its efficiency to a standard generalized-ensemble method, namely simulated tempering (ST).~\cite{Marinari1992,Lyubartsev1992} We thereafter apply the MS method to a greatly enlarged set with 1,024 sequences, also spanning the A1-N1 space, as well as another set spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively.
 Besides hinting at the feasibility of applying the MS method on large sets of sequences, the results allow us to carry out a more systematic analysis of the biophysical properties of sequences along mutational pathways connecting these two pairs of basic protein folds than has been possible with standard methods.~\cite{Holzgrafe2014,Holzgrafe2015}
@@
 \subsection{Model sequences}
-
-Six model sequences studied in this work are summarized in Table~I.
-In addition, we study two sets of 1,024 sequences derived from the A1-N1 and A2-TN pairs (see text) denoted $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$, respectively, and a restricted set $\mathrm{S16}_{144}$ with 144 sequences taken from $\mathrm{S16}_{1024}$ ($\mathrm{S16}_{144}$ was also studied in Ref.~\protect\citenum{Holzgrafe2014}).
+\noindent
+Six of the model sequences studied in this work, A1, N1, R1, R2, A2, and TN, are given in Table~I.
+In addition, we study two sequence sets $\mathrm{S16}_{1024}$  and $\mathrm{S35}_{1024}$ with 1,024 sequences each derived from the A1-N1 and A2-TN pairs, respectively, through mutational combinations, as well as the set $\mathrm{S16}_{144}$ taken from  Ref.~\protect\citenum{Holzgrafe2014}.
 
@@
 A sequence move is attempted every $1,000$ MC steps while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
-The simulations carried out in this work are summarized in Tables~II.
+The simulations carried out in this work are summarized in Table~II.
 
@@
 \begin{ruledtabular}
-\begin{tabular}{lccccr}
-Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes thermalization with $10^6$ MC steps.} &  Sequences\\
+\begin{tabular}{lcccr}
+Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes a thermalization step with $10^6$ MC steps/run.} &  Sequences\\
 \hline
-32 & ST & 0.43--0.65 & $10^7$ &A1\\ 
-32 & ST & 0.43--0.65 & $10^7$ &N1\\ 
-32 & ST & 0.43--0.65 & $10^7$ &R1\\ 
-32 & ST & 0.43--0.65 & $10^7$ &R2\\ 
-32$\times$8\footnote{32  runs per temperature for 8 temperatures in the range given.} & MS &0.43--0.65& $18\times 10^7$ & $\mathrm{S16}_{144}$\\
+32 & ST & 0.43--0.65 & $1\times10^7$ &A1\\ 
+32 & ST & 0.43--0.65 & $1\times10^7$ &N1\\ 
+32 & ST & 0.43--0.65 & $1\times10^7$ &R1\\ 
+32 & ST & 0.43--0.65 & $1\times10^7$ &R2\\ 
+32$\times$8\footnote{32 runs per temperature at 8 different temperatures.} & MS &0.43--0.65& $18\times 10^7$ & $\mathrm{S16}_{144}$\\
 16 & MS & 0.43  & $5\times 10^9$ &  $\mathrm{S16}_{1024}$ \\
@@
 \noindent
-Fold stabilities are calculated as in Ref.~\citenum{Holzgrafe2015} and also described briefly below.
-We define first two structural similarity measures $\QIA$ and $\QIB$ for folds IA and IB, respectively, indicating what fraction of the fold-specific backbone-backbone hydrogen bonds have been formed.
+Fold stabilities are calculated as in Ref.~\citenum{Holzgrafe2015} and described briefly below.
+First we define two structural similarity measures $\QIA$ and $\QIB$ for folds IA and IB, respectively, indicating the fraction of the fold-specific backbone-backbone hydrogen bonds that have been formed.
 The fold IA-hydrogen bonds are (2,6), (3,7), (4,8), (5,9), (6,10), (7,11), (8,12), (9,13), (10,14), (11,15) and the fold IB-bonds are (3,14), (5,12), (7,10), (10,7), (12,5), (14,3), where (i,j) indicates a hydrogen bond between the CO group of amino acid i and the NH group of amino acid j.
-The  stabilities of fold IA and IB are then defined as the probabilities $\PIA = P(\QIA\ge0.8)$ and $\PIB = P(\QIB\ge0.80)$, respectively, i.e., stability is the probability that at least 80\% of the fold's hydrogen bonds are formed.
-$\PIA$ and $\PIB$ depend on both sequence $\sbar$ and temperature $T$.
-For example, $\PIA=0.xx$ for A1 and $\PIB=0.xx$ for N1 at $\kb T = 0.43$.
-Structural similarity measures for 35-amino acid folds IIA and IIB are defined as $\QIIA = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IA}^\mathrm{20-35} + Q_\mathrm{tert}) ) / 3$ and $\QIIB = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IB}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$, respectively, where superscripts on $\QIA$ and $\QIB$ indicate over which amino acid positions those measures are applied to within the longer 35 amino acid sequences and $Q_\mathrm{tert}$ is a measure that counts the number of $C_\beta$-$C_\beta$ contacts between the two secondary structure elements of these folds.~\cite{Holzgrafe2015} As for IA and IB, the stabilities of folds IIA and IIB are defined as $\PIIA = P(\QIIA\ge0.8)$ and $\PIIB = P(\QIIB\ge0.80)$, respectively.
-%In calculating $Q$ values, a hydrogen bond is considered formed if (i) the NHO and COH angles are both $>90^\circ$ and (ii) the HO distance is $<2.7$~{\AA}.
+The  stabilities of folds IA and IB are then defined as the probabilities $\PIA = P(\QIA\ge0.8)$ and $\PIB = P(\QIB\ge0.80)$, respectively, i.e., the probability that at least 80\% of the fold's hydrogen bonds are formed.
+$\PIA$ and $\PIB$ thus depend on both sequence $\sbar$ and temperature $T$.
+For example, $\PIA=0.875\pm 0.003$ for A1 and $\PIB=0.785\pm0.008$ for N1 at $\kb T = 0.43$.
+Structural similarity measures for 35-amino acid folds IIA and IIB are defined as $\QIIA = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IA}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$ and $\QIIB = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IB}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$, respectively, where superscripts on $\QIA$ and $\QIB$ indicate over which amino acid positions those measures are applied to within the longer 35 amino acid sequences and $Q_\mathrm{tert}$ is a measure that counts the number of $\mathrm{C}_\beta$-$\mathrm{C}_\beta$ contacts between the two secondary structure elements of these folds.~\cite{Holzgrafe2015} In analogy with $\PIA$ and $\PIB$, we define the stabilities of folds IIA and IIB as $\PIIA = P(\QIIA\ge0.8)$ and $\PIIB = P(\QIIB\ge0.80)$, respectively.
+The root-mean-square-deviation, RMSD, is calculated over all $\mathrm{C}_\alpha$ atoms.
 
@@
 \includegraphics[width=10.0cm]{MCTrajFolds}
-\caption{(A) Example simulation trajectory of the MS algorithm applied to the sequence set $\mathrm{S16}_{144}$.
+\caption{(A) Example of an MS simulation applied to the sequence set $\mathrm{S16}_{144}$.
 The plot shows the MC evolution of the sequence $\sbar$ (numbered 1--144), the total potential energy $E$ and the root-mean-square deviation (RMSD) calculated against fold IA (light blue) and fold IB (dark red) structures shown in (B).
-The simulation is carried out at $\kb T = 0.43$ corresponding to the lowest studied temperature.
+The simulation is carried out at $\kb T = 0.43$.
 (B) Representative structures of fold IA and fold IB, selected to be the minimum-energy conformations found for the sequences A1 and N1, respectively.
@@
 
-%
-%\begin{equation}
-%Q_\mathrm{\alpha\beta\beta} = ( Q_\alpha^\mathrm{1-12} + Q_\beta^\mathrm{XX} +Q_\beta^\mathrm{20-35} + Q_\mathrm{tert} ) / 4
-%\end{equation}
-%
-%
-%\begin{equation}
-%Q^*_\mathrm{\alpha\beta} = [ \max(Q_\alpha^\mathrm{1-16},Q_\beta^\mathrm{XX}) +Q_\beta^\mathrm{20-35} + Q_\mathrm{tert}) ] / 3
-%\end{equation}
-%
-
 \section{Results}
@@
 \noindent
-To assess the computational efficiency and properties of the multisequence algorithm, we used as a test case the set of 144 3-letter model sequences with 16 amino acids studied in Refs.\cite{Holzgrafe2014} Two of the sequences are A1 and N1 which were designed to fold into stable $\alpha$-helix and $\beta$-hairpin structures, respectively.
+We start by testing the MS algorithm on the set {\SR} consisting of the 144 3-letter model sequences with 16 amino acids studied in Ref.~\citenum{Holzgrafe2014}.
+Two of the sequences in {\SR} are A1 and N1 (see Table~I) which were designed to fold into stable $\alpha$-helix and $\beta$-hairpin structures, respectively, as shown in Fig.~2B.
 A1 and N1 differ at 10 positions such that 10 consecutive point mutations can transform A1 into N1, and vice versa.
 The binary sequence space between A1 and N1 in which any combination of these mutations have been carried out, therefore contains $2^{10}=1,024$ sequences.
-The 144 sequences were selected from this binary space with the constraint that the total number of h-amino acids are not too high or too unevenly distributed.~\cite{Holzgrafe2014}
+The 144 sequences were selected from this binary space with the constraint that the total number of h-amino acids are not too high and not too unevenly distributed.~\cite{Holzgrafe2014}
+
+\begin{figure}
+\includegraphics[width=7.8cm]{Pacc}
+\caption{Acceptance rates for $\sbar\rightarrow\sbar'$ updates in MS simulations of the {\SR} sequence set as a function of (A) the number of changed amino acid positions $\Delta h$ and (B) temperature $T$.
+Acceptance rates for 3 different $T$'s are shown in (A).}
+\end{figure}
+
+We apply the multisequence algorithm on the set {\SR} across a range of temperatures (see Table~II).
+Figure~2 illustrates a typical MS simulation trajectory carried out at the lowest temperature which is well below the folding temperature of both A1 and N1~\cite{Holzgrafe2014,Holzgrafe2015}.
+From the MC evolution of the total energy $E$, sequence index $\sbar$, and RMSD values from the representative structures in Fig.~2B, it is evident that visits to various sequences drive transitions into a range of structural states.
+In particular, there are frequent visits to both $\alpha$-helix and $\beta$-hairpin structures.
+Transitions between them are accompanied by a shift in which sequences are preferably visited.
+For example, visits to high $\sbar$-indices, including N1 with index 144, tend to coincide with formation of $\beta$-hairpin structures.
+
+One might have suspected that the MS algorithm would be hampered by poor acceptance rates for sequence updates.
+However, this is not the case.
+We carry out updates $\sbar\rightarrow\sbar'$ by picking a new random sequence $\sbar'\ne\sbar$ from the set of allowed sequences.
+The (average) acceptance rate  depends on both $T$ and the step in sequences space $\Delta h$, i.e., the number of amino acid positions changed, as shown in Fig.~3.
+At the lowest $T$ and high $\Delta h$, acceptance rates are only around 0.1-0.2 but are substantially higher at other $T$ and $\Delta h$ values.
+The overall acceptance rates are in fact above the oft-quoted rule-of-thumb value $0.25$~\cite{Gilks1996} for all temperatures $T$ (see Fig~4B).
+Increased acceptance rates can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer protein chains.
 
@@
 \includegraphics[width=8.0cm]{Stderr}
-\caption{Comparing MS and ST algorithms.
-The statistical errors $\sigma_{\left < E\right >}$ of the average total energy $\left < E\right >$ are estimated with the two methods for the sequences (A) A1, (B) N1, (C) R1 and (D) R2.
-Simulation lengths are adjusted such that the number of conformations sampled per sequence and temperature is similar between in MS and ST runs.
+\caption{Comparing sampling efficiency of the MS and ST algorithms.
+Statistical errors $\sigma_{\left < E\right >}$ of the average total energy $\left < E\right >$ from the two methods for the 16-amino acid sequences (A) A1, (B) N1, (C) R1 and (D) R2.
+Simulation lengths in the two methods are adjusted such that the number of conformations sampled per sequence and temperature is roughly the same (see text).
 }
@@
 
-We apply the multisequence algorithm to this set of 144 sequences across a range of temperatures (see Table~II).
-Figure~3 illustrates a typical simulation carried out at the lowest temperature, $\Tlow$, which is well below the folding temperature of both A1 and N1~\cite{Holzgrafe2014,Holzgrafe2015}.
-From the MC evolution of the total energy $E$, sequence index $\sbar$, and similarity measures $Q_A$ and $Q_B$, it is evident that visits to various sequences drive transitions into a range of structural states.
-For example, visits to A1 ($\sbar=1$) and N1 ($\sbar=144$) tend to coincide with visits to high-$Q_A$ and high-$Q_B$ conformations, respectively, as is necessary to generate the correct conformational ensembles for these sequences.
+We now compare the results from our MS calculations with simulated tempering (ST) simulations carried out on 4 of the 144 sequences, namely A1 and N1 and two sequences, R1 and R2, chosen randomly at distances $h=4$ and $h=6$ from A1, respectively (see Table~I).
+While ST provides the thermodynamics of a given sequence across a range of $T$ in a single run, a MS simulation provides the thermodynamics of all 144 sequences for one $T$.
+We adjust the simulation lengths for ST and MS runs such that roughly the same number of sampled conformations are obtained for each $\sbar$ and $T$ combination, thus ensuring that similar computational resources are used for the two algorithms (see Table~II).
+The MS algorithm is validated by comparing the average total energy, $\left < E \right >$, calculated for different $T$'s and $\sbar$'s, between the two sets of results (see Supplementary Information).
 
-Sequence updates $\sbar\rightarrow\sbar'$ are carried out by picking a new sequence $\sbar'$ randomly from the entire set of allowed sequences.
-Therefore, a sequence update typically changes more than one amino acid position.
-In other words, if the step size in sequence space is $\Delta H$, then typically $\Delta H>1$.
-Interestingly, despite the large $\Delta H$ and low $T$, the overall (average) acceptance rate $P_\mathrm{acc}=0.xx$ for $\sbar\rightarrow\sbar'$ updates which is not unreasonably low.
-Higher $P_\mathrm{acc}$ can be obtained by restricting proposed moves such that $\Delta H\le\Delta H_\mathrm{max}$, as shown in Figure~3.
-For example, an acceptance rate close to the oft-quoted rule-of-thumb value $0.25~\cite{Gilks1996} $ is obtained with $\Delta H_\mathrm{max}=4$, and with $\Delta H_\mathrm{max} = 1$ (only point mutations), $P_\mathrm{acc}=0.xx$.
-For simplicity, in this work, all simulations are carried out without restrictions on how new sequences $\sbar'$ are drawn.
-
-\begin{figure}
-\includegraphics[width=7.8cm]{Pacc}
-\caption{Acceptance rates for $\sbar\rightarrow\sbar'$ updates in MS simulations of the {\SR} sequence set as a function of (A) the number of changed amino acid positions  and (B) temperature. }
-\end{figure}
-
-In order to assess computational efficiency, we compare with results from simulated tempering (ST) simulations carried out on 4 of the 144 sequences, namely A1 and N1 and the two sequences R1 and R2 chosen randomly at $H=4$ and $H=6$ from A1, respectively (see Table~I).
-While ST provides the thermodynamics of a given sequence across the 8 temperatures $T_1, ..., T_8$ in a single run, a single MS simulation provides the thermodynamics of all 144 sequences for one $T_\mathrm{i}$.
-We adjust the simulation lengths for ST and MS runs such that an identical number of sampled conformations are obtained for each $\sbar$ and $T_\mathrm{i}$ combination, ensuring also that similar computational resources are used for the two algorithms.
-
-Figure 4 shows the statistical error, $\sigE$, estimated for the average energy, $\left <E\right >$, for the 4 different sequences using ST and MS algorithms.
-Since an identical number of sampled conformations were obatined for each s and $T_i$ combination, we can therefore compare the statistical errors directly.
-Both MS and ST algorithms give the same small error at the highest T studied which implies consistent result between them.
-It should be noted that the highest T studied is well above the folding temperature for all the 4 sequences.
+As a way to assess sampling efficiency, we compare in Figure~4 the statistical error, $\sigE$, of the average energy $\left <E\right >$ for the 4 different sequences obtained using ST and MS, respectively.
+Since roughly the same number of sampled conformations were obtained for each s and $T$ combination, we can compare the statistical errors directly.
+Both MS and ST algorithms give the same small error at the highest $T$ studied which implies consistent result between them.
+It should be noted that the highest $T$ studied is well above the folding temperature for all the 4 sequences.
 Therefore, this consistent result is not particularly surprising since at high-T the free-energy landscape is relatively smooth.
@@
 
-Insight from the statistical measurements suggests that the MS algorithm has a similar computational efficiency at $high_T$ and even better at lower T's.
-The MS algorithm is able to achieve  better efficency by sampling a wider region of both conformational and sequence space.
+Insight from the statistical measurements suggests that the MS algorithm has a similar computational efficiency at high $T$ and even better at lower $T$'s.
+The MS algorithm is able to achieve  better efficiency by sampling a wider region of both conformational and sequence space.
 Sampling the sequence space in addition to conformational space ensures a decorrelated ensemble of states is generated.
@@
 \rotatebox{0}{\includegraphics[width=16.6cm]{network}}
-\caption{Networks of the A1-N1 (top) and A2-TN (bottom) binary sequence space showing the possible stable pathways from one fold to another.
-(light blue) Fold A, (dark red) Fold B, and (black) Bistable.
-}
+\caption{Networks of sequences connecting folds IA and IB (top) and folds IIA and IIB (bottom).
+Each node represents a stable sequence ($\Ptot\ge\Pcut$ where $\Pcut=0.50$) that folds into either IA or IIA (light blue), IB or IIB (dark red), or is classified as bistable ($B>0.5$, black).
+A line between two nodes indicates that the sequences differ at only one position.
+Graph created using the tool Graphviz~\protect\cite{Graphviz2000} obtained from  www.graphviz.org.}
 \end{figure*}
 
-Hence, we conclude that in the comparison made above the computational efficiency of the MS algorithm is comparable to ST.
+%Hence, we conclude that in the comparison made above the computational efficiency of the MS algorithm is comparable to ST.
 It must be stressed, however, that while the ST simulation requires that the thermodynamic behavior is calculated at a range of temperatures, MS works well on its own directly at low temperatures.
@@
 \noindent
-
-We now apply the MS algorithm to the sets {\SI} and {\SII} which represent the full binary sequence space for the A1-N1 and A2-TN systems, respectively.
-Each set includes $2^{10}=1,024$ sequences.
-Using MS simulations, we determined the thermodynamic behavior at the temperatures $\TlowI$ and $\TlowII$ for the A1-N1 and A2-TN systems, respectively (see Table~1).
-In particular, this allows us to find the stabilities relative to folds IA and IB ($\PIA$ and $\PIB$) for all sequences in {\SI} and the stabilities relative to folds IIA and IIB ($\PIIA$ and $\PIIB$) for all sequences in {\SII}.
+We now turn to the full binary sequence sets {\SI} and {\SII} with 1,024 sequences each.
+By applying the MS method to these two sets separately (see Table~II), we determined the thermodynamic behavior of each included sequence at relatively low temperatures.
+In particular, these simulations allowed us to find the stabilities relative to folds IA and IB ($\PIA$ and $\PIB$) for all sequences in {\SI} and the stabilities relative to folds IIA and IIB ($\PIIA$ and $\PIIB$) for all sequences in {\SII}.
 
@@
 
-\subsection{Stability properties of fold-to-fold mutational pathways}
+\subsection{Stability and bistability properties in of fold-to-fold mutational pathways}
 \noindent 
@@
 \includegraphics[width=7.8cm]{Paths}
-\caption{(Left) A1-N1 System.
-(Right) A2-TN System.
-(Top) Average stability as a function of Hamming distance $h$.
-(Middle) Fraction of pathways as a function of Swicth length $L_s$.
-(Bottom) Average stability as a function of Bistability $B$.}
+\caption{Stability properties of mutational pathways.
+The total stability $\Ptot$ as a function of the distance $h$ from A1 averaged over all (A) IA-IB and (B) IIA-IIB mutational paths obtained with $\Pcut=0.50$.
+Error bars indicate maximum and minimum $\Ptot$ values.
+The distribution of switch lengths $L_\mathrm{s}$ for the (C) IA-IB and (D) IIA-IIB mutational paths ($\Pcut=0.50$).
+C and D insets: Average switch length $\left < L_\mathrm{s}\right >$ across all paths as a function of $\Pcut$.
+Scatter plots of $\Ptot$ versus bistability $B$ for all sequences in (E) {\SI} and (F) {\SII}, where $B=1-\Delta P/\Ptot$ and $\Delta P = |\PIA-\PIB|$ or $|\PIIA-\PIIB|$.}
 \end{figure}
@@
 
-However, the smooth stability trends in Fig.~6A and B belie the real character of the \textit{individual} mutational pathways which tend to exhibit an abrupt switch between the two folds.
+However, the smooth stability trends in Fig.~6A and B belie the real character of the individual mutational pathways which tend to exhibit an abrupt switch between the two folds.
 To examine the statistics of fold transitions in our model, we first make a distinction between two types of stable sequences: those that fold into a single unique fold, thus behaving as classical proteins, and those that exhibit significant stabilities of both folds.
@@
 
-We consider a stable sequence to be bistable if $B>0.5$, where $B$ is a bistability measure (see Fig.~6 legend); otherwise, it is considered a classical protein.
-In principle, a fold transition can occur directly between two classical proteins with unique folds, or it can proceed via one or more intermediate bistable sequences which populate both folds to some extent.
+We consider a stable sequence to be bistable if $B>0.5$, where $B$ is a bistability measure (see Fig.~6 legend).
+In principle, a fold transition can occur directly between two classical proteins with unique folds, or it can proceed via one or more intermediate bistable sequences which populate both folds.
 We  define the switch length of a mutational pathway $L_\mathrm{s}=2+n_\mathrm{B}$, where $n_\mathrm{B}$ is the number of bistable sequences in between the two classical sequences that define the switch point.
 Hence, a path with $L_\mathrm{s}=2$ accomplishes a fold switch in a single mutational step without going through a bistable point.
-From the distributions of  $L_\mathrm{s}$ in Fig.~5C and D taken over all pathways with $\Pcut=0.50$, it can be seen that fold switching along individual pathways are typically completed in 1-2 mutations and often in a single step in the A1-N1 system.
-Hence, fold switching is typically abrupt in our model and, for $\Pcut = 0.5$, it is fairly common that viable pathways pass through one or two bistable sequences.
+From the distributions of  $L_\mathrm{s}$ in Fig.~5C and D taken over all pathways with $\Pcut=0.50$, it can be seen that fold switching along individual pathways are typically completed in only 1-2 mutations and often a single step is enough switch between the IA and IB folds.
+Hence, fold switching is typically abrupt and, for $\Pcut = 0.5$, it is fairly common that viable pathways pass through one or two bistable sequences.
 
-Interestingly, fold switching through bistable sequences become less common as selection for more stable pathways is made in the A1-N1 system, as can be seen from the decrease in $\left <L_\mathrm{s}\right >$ as a function of $\Pcut$ (see Fig.~5C(inset)).
-For $\Pcut=0.70$, there are no longer any viable pathways between the $\alpha$-helix and $\beta$-hairpin folds that passes through a bistable point because $\left <L_\mathrm{s}\right > =  2$.
-The situation for the A2-TN system is different, where $\left <L_\mathrm{s}\right >$ is relatively stable around 3.
-For this system, there are no direct $L_\mathrm{s}=2$ paths available even at $\Pcut=0.50$.
-Therefore, bistable sequences play a crucial role in bridging the folds IIA and IIB even though their stabilities are significantly reduced.
+Interestingly, IA-IB fold switching through one or more bistable sequences become less and less common as selections for more stable pathways are made.
+This can be seen from the decrease in $\left <L_\mathrm{s}\right >$ as a function of $\Pcut$ (see Fig.~5C(inset)).
+For $\Pcut\ge0.70$, there are no longer any viable pathways between the $\alpha$-helix and $\beta$-hairpin that passes through a bistable sequence because $\left <L_\mathrm{s}\right > =  2$.
+An underlying reason for the occurrence of sharper fold switches for more stable mutational pathways is apparent from a comparison between $\Ptot$ and $B$ across all sequences in {\SI}.
+As shown in Fig.~5E, sequences with the highest $\Ptot$ tend to exhibit very little bistability.
+Hence, highly stable mutational pathways are therefore forced to go through abrupt switch points, thereby transitioning between folds in a single mutational step.
+The situation for the IIA-IIB fold switch is more complicated.
+Sequences in {\SII} too follow the trend that high overall stability is achieved for only classical proteins with a single dominant fold.
+However, bistable sequences nonetheless play a crucial role in bridging the IIA-IIB folds because paths without bistable intermediate sequence are not available for $\Pcut\ge0.50$.
+Hence, passing through sequences with bistable properties may still be required to achieve mutation-induced fold switching between certain fold pairs even through it lead to a reduced stability at the switch point.
 
-An underlying reason for the preferential behavior is apparent by comparing the total stability and bistability of all sequences.
-As shown in Fig.~5E and F, there is a clear trend that sequences with high total stability do not exhibit high bistability but instead they tend to have one clearly dominant fold.\\
+%The sequences {\SII} do also follow this same trend (Fig.~5F).
+However, bistable sequences do play a crucial role in bridging the folds IIA and IIB even though their stabilities are significantly reduced, all stable mutational pathways must pass through one bistable sequences.
+\\
 
@@
 \noindent
-We have tested a generalized-ensemble algorithm for coarse-grained simulations of biomolecules that works by making the sequence a dynamic parameter, thereby leading to a random walk in sequence space.
-This approach was evaluated for protein folding simulations by comparing with ST.
-The results indicate two main benefits of the MS approach.
-On the one hand, it provides a convenient way to sample the respective canonical distributions for a large set of sequences in a single run and, on the other, it also enhances the sampling of conformational space.
+The simulation algorithm for biomolecules considered here works by making the sequence a dynamic parameter.
+We tested this approach by applying it to proteins.
+The results indicate that there are two main potential benefits.
+Firstly, it provides a convenient way to sample the respective canonical distributions for a large set of sequences in a single run and, secondly, it enhances the sampling of conformational space such that it can be applied directly at low temperatures.
 
-The conformational sampling efficiency of the MS algorithm can be assessed from the comparison with ST in Figure~3.
-Although there is no single ``fair" way to compare MS and ST, we chose as a measure of efficiency the estimated statistical error on the total energy, $\sigE$, obtained with roughly the same computational cost per temperature and sequence.
+The conformational sampling efficiency can be assessed from the comparison with ST.
+Although there is no single ``fair" way to compare the MS and ST methods, we chose as a measure of efficiency the statistical error of the total energy, $\sigE$, obtained with roughly the same computational cost per temperature and sequence.
 At the highest studied temperatures, which are well above the folding temperatures of both A1 and N1, the statistical errors $\sigE$ are basically the same for the two methods.
-This is not unexpected because conformational sampling of short polymers at high $T$ does not involve crossing any major energy barriers.
+This finding is not unexpected because conformational sampling of short polymers at high $T$ does not involve crossing any major energy barriers.
 As a result, successively sampled conformations for a given combination of sequence $\sbar$ and temperature $T$, are therefore likely mostly uncorrelated in both methods leading to similar $\sigE$ values.
@@
 It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant temperature.
-Hence, rather than promoting escape from local minima in the free energy landscape by visits to higher temperature, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences for which the minima may not be as deep or are altogether absent.
-For the same reason, the performance of the MS algorithm will depend on the size and character of sequences included.
-In particular, the MS simulation likely hinge on the inclusion of at least some sequences with poor stabilities such that at least partial unfolding of the chain is regularly triggered and promoting transitions to new conformational states.
+Hence, rather than promoting escape from local minima by visits to higher temperature, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences where the minima are not as deep or maybe altogether absent.
+For the same reason, we expect the performance of the MS algorithm to depend on the size the sequence set used, as well as their character.
+Specifically, the MS simulation may hinge on the inclusion of at least a few sequences with poor stabilities such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
 
-The MS method should not be seen as a general method to speed up conformational sampling due to the complexity of performing sequence updates in most biomolecular models.
-However, our results indicate that for coarse-grained models that permit the sequence updates as a Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be highly efficient.
-A notable feature of the algorithm is that it provides escape from local minima on the free energy landscape of proteins without resorting to a change in $T$.
-This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in the context or ordered dense phases such as lipid bilayers, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Simulations of lipid bilayers at elevated $T$ must typically be avoided because it can lead to unwanted perturbations of the membrane structure, although it can be a viable approach to speed up conformational sampling in particular cases.~\cite{Ulmschneider2010}
-
-%Overall our results show that, when technically feasible, a multisequence approach can be highly efficient for calculating the thermodynamics of a large number of amino acid sequences even at low temperatures.
-To demonstrate that the method to large sequence sets we applied it to the full binary sequences spaces between the model sequence pairs A1-N1 and A2-TN, which both include 1,024 sequences.
-These calculations would have been difficult to carry out ``in series" using some other standard method such as simulated tempering.
-
-%For comparison, we used the well-established generalized-ensemble ST algorithm and focused on 144 model sequences with 16 amino acids ({\SR}).
-As a measure of computational efficiency, we used the statistical error $\sigE$ in the estimation of the average total energy $E$ for 4 different sequences which are part of {\SR}.
-
-%Hence, we conclude that the computational efficiency of the MS algorithm is comparable to ST.
-However, that while the ST simulation requires that the thermodynamic behavior is calculated at a range of temperatures, MS works well on its own directly at low temperatures.
-ST was introduced as an extension to the standard, fixed-$T$ Metropolis-Hastings algorithm in order to overcome the poor mixing of states at low temperatures, i.e., the tendency for the simulations at low $T$ to become trapped in local free energy minima.
-The MS algorithm, however, efficiently calculates the thermodynamics at low $T$'s without visits to other $T$'s.
-In the MS algorithm, it is visits to sequences with poor native state stability rather than high $T$'s that drive the decorrelation of generated states and escape from local traps.
-This way, calculation of low-$T$ thermodynamics is enhanced in MS simulations while at the same time enhancing coverage in sequence space.
-This makes the MS algorithm ideal for situations in which the finite-$T$ thermodynamic behavior of multiple sequences are of interest.
-
-%A notable feature of the MS algorithm is that provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
-This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in lipid membranes, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Sampling in membrane simulations at high $T$, such as in replica-exchange simulations, are usually avoided in membrane simulations because it can lead to unwanted perturbation of the lipid bilayer structure although it can be a viable approach in particular cases.~\cite{Ulmschneider2010}
-
-%A notable feature of the MS algorithm is that it provides escape from local energy minima on the free energy landscape of proteins without resorting to a change in temperature.
-This feature could make it attractive as an alternative to exploring the equilibrium behavior of proteins and peptides in lipid membranes, where escape from local minima are especially challenging even in CG simulations.~\cite{Bereau2015} Sampling in membrane simulations at high $T$, such as in replica-exchange simulations, are usually avoided because it can lead to unwanted perturbation of the lipid bilayer structure although it can be a viable approach in particular cases.~\cite{Ulmschneider2010}
+The MS method should not be seen as a general method to speed up conformational sampling.
+However, our results indicate that for coarse-grained models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be highly efficient.
+The ability to promote conformational sampling without resorting to an increased $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers, where escape from local minima can be especially challenging~\cite{Bereau2015} and elevated $T$ is typically avoided because it can lead to unwanted perturbations of the membrane structure.~\cite{Ulmschneider2010}
 
@@
 \noindent
-We have evaluated an algorithm for coarse-grained biomolecular simulations that allow the thermodynamics of multiple different sequences to be calculated in a single run.
-Here we showed that with this MS algorithm, the folding thermodynamics of over 1,000 amino acid sequences with each 35 amino acids could be determined in a coarse-grained model with 3 amino acid types.
-In addition to being convenient for calculations of multiple sequences, the method enhances conformational sampling with an efficiency comparable to simulated tempering.
-Random walks across sequence space allow both sampling of many sequences and efficient escape from local free energy minima in the free energy landscape.
-The method might be well suited for coarse-grained simulations of other biomolecular systems, such as phospholipid bilayers.
+We have evaluated an algorithm for biomolecular simulations that allow the thermodynamics of multiple different sequences to be calculated in a single run.
+Here we applied this algorithm to protein folding and showed that the low-temperature thermodynamic behavior of over 1,000 amino acid sequences with each 35 amino acids could be determined in an intermediate-resolution CG model with 3 amino acid types.
+In addition to being convenient for determination of the thermodynamics of multiple sequences, we found that the method also enhances conformational sampling such that it can be applied also at low temperatures.
+A random walk in sequence space allows for efficient escape from the the minima present in the free energy landscapes of individual sequences.
+The method might be well suited for CG simulations of other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
 

### 2017-06-02 Manuscript.tex
Modified computational efficiency section

@@
 As a way to assess sampling efficiency, we compare in Figure~4 the statistical error, $\sigE$, of the average energy $\left <E\right >$ for the 4 different sequences obtained using ST and MS, respectively.
-Since roughly the same number of sampled conformations were obtained for each s and $T$ combination, we can compare the statistical errors directly.
-Both MS and ST algorithms give the same small error at the highest $T$ studied which implies consistent result between them.
+Since roughly the same number of sampled conformations were obtained for each $\sbar$ and $T$ combination, we compare the statistical errors directly.
+The two algorithms give similar errors at the highest $T$ studied which implies consistent result between them.
 It should be noted that the highest $T$ studied is well above the folding temperature for all the 4 sequences.
-Therefore, this consistent result is not particularly surprising since at high-T the free-energy landscape is relatively smooth.
-Hence, conformational space requires little or no difficulty to sample.
+Therefore, this consistency is not particularly surprising since at high-T the free-energy landscape is relatively smooth.
+Hence, conformational space requires little difficulty to sample.
 
-There is a sequence-dependent statistical trend observed in Figure 6.
-As temperature decreases, the statistical error from the MS algorithm becomes smaller than those from ST.
-Even at the lowest T studied, below the folding temperature of all 4 sequences, we notice that the MS algorithm is at least consistent and most times better than ST in its estimates except for N1.
-This departure from the dominant trend is most likely due to statistical flunctuations, in which case, does not have any considerable significance.
-
-Insight from the statistical measurements suggests that the MS algorithm has a similar computational efficiency at high $T$ and even better at lower $T$'s.
-The MS algorithm is able to achieve  better efficiency by sampling a wider region of both conformational and sequence space.
-Sampling the sequence space in addition to conformational space ensures a decorrelated ensemble of states is generated.
+However, as temperature decreases, the statistical error from the MS algorithm generally becomes smaller than those from ST.
+In particular, Fig.~4C illustrates that MS is on average twice as accurate as ST in its error estimates for R1.
+The best estimate occurs at $\kb\approx$ 0.46 where MS is about 450\% better than ST.
+Even at the lowest T studied, MS is at least consistent or better than ST for all 4 sequences.
 

### 2017-06-06 Manuscript.tex
miscellaneous changes

@@
 
-\title[Multisequence Monte Carlo simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: application to protein fold switching}
+\title[Multisequence Monte Carlo simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: exploring the sequence-structure relationship of proteins}
 
@@
 
-\keywords{Monte Carlo, generalized ensembles, simulated tempering, protein folding, protein evolution}
+\keywords{Monte Carlo, generalized ensembles, protein folding, protein fold switching}
 
@@
 \noindent
-Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} These advances have in part triggered major efforts~\cite{McGuffee2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the molecular dynamics of bio-systems of different sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} or a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging and insightful, these large-scale simulations have also highlighted the fact that severe tradeoffs in size and time scales will likely persist for the foreseeable future for detailed simulations.
+Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} Together with the ever-increasing availability of computational resources, these advances have triggered a few major efforts~\cite{McGuffee2010,Miao2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the dynamics of biomolecular systems of various sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} and a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging and insightful, these large-scale simulations have also highlighted the fact that severe tradeoffs in size and time scales will likely persist for the foreseeable future.
 
-One way to expand the range of biomolecular simulations is to turn to coarse-grained (CG) models, where the basic aim is to simplify the physical description of interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list 4 main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that a CG representation either of the interaction potential or the molecular geometry often opens up for alternative sampling schemes beyond traditional molecular dynamics approaches which can further speed up conformational sampling, e.g., activation-relaxation kinetics,~\cite{Beland2011} discrete molecular dynamics~\cite{Proctor2011} and various Monte Carlo (MC)-based techniques such as cluster moves.~\cite{Vitalis2009}
+One way to expand the range of biomolecular simulations is to turn to coarse-grained (CG) models, where the basic aim is to simplify the physical description of interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list 4 main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that a CG representation of either the interaction potential or the molecular geometry often opens up for alternative sampling schemes beyond traditional molecular dynamics approaches, which can further speed up conformational sampling.
+Examples of such sampling schemes include activation-relaxation kinetics,~\cite{Beland2011} discrete molecular dynamics~\cite{Proctor2011} and various Monte Carlo (MC)-based techniques such as cluster moves.~\cite{Vitalis2009}
 
-The challenges of achieving representative conformational sampling of individual biomolecular systems notwithstanding, a range of systems naturally call for the investigation and comparison of many molecular variants, e.g., determining the molecular mechanisms of specificity in protein-protein~\cite{Zarrinpar2003,Hakes2007} or protein-nucleotide interactions,~\cite{Rohs2010} or the role of mutations in disease processes such as protein aggregation.~\cite{Ross2004} Another example is protein folding where unique insight has been achieved by comparing the folding within and between protein families.~\cite{Tzul2017,Wensley2010} With the increasing availability of sequence information,~\cite{Vukmirovic2000} it is of interest to explore ways to efficiently sample multiple sequences in biomolecular simulations.
+The challenges of achieving representative conformational sampling of individual biomolecular systems notwithstanding, many biological processes naturally call for the investigation and comparison of molecular variants.
+This is required, for example, in determining the molecular mechanisms of specificity in protein-protein~\cite{Zarrinpar2003,Hakes2007} or protein-nucleotide interactions,~\cite{Rohs2010} and the role of mutations in disease processes such as protein aggregation.~\cite{Ross2004} Another example is the protein folding process, where unique insight has been achieved by comparing the folding within and between protein families.~\cite{Tzul2017,Wensley2010} In a situation with rapid growth of sequence information,~\cite{Vukmirovic2000} it is of interest to explore ways to efficiently sample multiple sequences in biomolecular simulations.
 
-To this end, we evaluate in this work an MC-based algorithm that can in principle calculate the thermodynamics of multiple sequences in a single run and apply it to a coarse-grained model for protein folding.
-This ``multisequence" (MS) method was originally developed in the context of homo- and heteropolymer simulations~\cite{Irback1995} and was later adapted for the characterization of peptide-protein binding specificity,~\cite{Bhattacherjee2013,Wallin2017}, but it has to our knowledge not been previously tested in protein folding simulations.
-The MS algorithm relies on sampling at fixed temperature a generalized ensemble in which the sequence is a dynamic parameter.
+To this end, we consider in this work an MC-based algorithm that can calculate the thermodynamics of multiple sequences in a single run and apply it to a coarse-grained model for protein folding.
+This multisequence (MS) method was originally developed in the context of homo- and heteropolymer simulations~\cite{Irback1995} and was also later adapted for the characterization of peptide-protein binding specificity.~\cite{Bhattacherjee2013,Wallin2017} To our knowledge, however, it has not been previously tested in realistic protein folding simulations.
+The MS algorithm carries out a simulation in a generalized ensemble that performs a random walk in sequence space.
 Hence, there are two main types of updates: conformational updates $\rbar\rightarrow\rbar'$ and sequence updates $\sbar\rightarrow\sbar'$.
-As illustrated in Fig.~1, this approach is easily applied when $\rbar$ and $\sbar$ are ``perpendicular" coordinates such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$.
-While this may be hard to achieve for detailed protein models, it is likely possible in many CG models.
+As illustrated in Fig.~1, this strategy can be applied when $\rbar$ and $\sbar$ are ``perpendicular" coordinates such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$.
+While this may be hard to achieve for detailed biomolecular models, it is likely possible in many CG models.
 
-As an initial test case, we apply the MS method to a set of 144 model sequences with 16 amino acids.
-This set, {\SR}, was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and $\beta$-hairpin, respectively, and allowed us to demonstrate the existence of A1-N1 mutational pathways which abruptly switch between the two structures and do not proceed through unstable sequences.~\cite{Holzgrafe2014} Here we use this set to validate the MS method and compare its efficiency to a standard generalized-ensemble method, namely simulated tempering (ST).~\cite{Marinari1992,Lyubartsev1992} We thereafter apply the MS method to a greatly enlarged set with 1,024 sequences, also spanning the A1-N1 space, as well as another set spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively.
-Besides hinting at the feasibility of applying the MS method on large sets of sequences, the results allow us to carry out a more systematic analysis of the biophysical properties of sequences along mutational pathways connecting these two pairs of basic protein folds than has been possible with standard methods.~\cite{Holzgrafe2014,Holzgrafe2015}
-
-%We find that sequences become gradually less stable as they approach switch point.
-However, fold switching is observed to be typically abrupt at the switch point.
-Folding pathways usually pass through bistable sequences which populate both folds up to a significant level.
-We show that while bistable sequences become less when selecting more stable pathways for A1-N1 system, they generally remain the same for the A2-TN system.
-
-%We first apply the MS algorithm to a set of 144 model sequences with 16 amino acids and compare with the simulated tempering (ST) algorithm.
-We find that the computational efficiency of the MS algorithm is never less than ST except at the lowest studied temperature for one sequence.
-
-%We then demonstrate that the multisequence algorithm can be used to calculate the thermodynamic behavior for $>1,000$ sequences in a 3-letter model for protein folding~\cite{Bhattacherjee2012} at temperatures well below the typical folding temperatures $T_\mathrm{f}$ of the included sequences.
-
-%Modern investigations into biophysical properties of biomolecular systems are routinely carried out not only on a single species but on multiple molecular variants.~\cite{Vukmirovic2000,Nickson2010} For example, protein engineering methods are used to provide mechanistic insight into the specificity of protein-protein and protein-DNA interactions, to compare the  conformational propensities of disease-related mutations of aggregation-prone proteins, and optimize biomolecular properties through directed evolution.
-Moreover, large-scale screening techniques such as phage display and microarray chips are used to provide biophysical properties across entire genomes.
-In short, through the study of a biomolecular system and related variants molecular variants are studied in order to gain a more complete picture of the biophysical and biochemical properties of the system and provide evolutionary context.
-
-%While protein or DNA sequences are easily changed on the computer, the high computational cost of simulating even a single moderately-sized biomolecule makes systematic explorations of sequence space challenging.
-Some notable efforts have nonetheless been attempted, including simulations of all 136 unique tetranucleotide sequences in B-DNA oligonucleotides using all-atom explicit water molecular dynamics.~\cite{Beveridge2004}
-
-%While the 16-amino acid A1 and N1 sequences fold into $\alpha$-helical and $\beta$-hairpin folds, respectively, the two 35-amino acid sequences A2 and TN adopt the folds denoted IIA and IIB, as shown in Fig.~X.
+As an initial test case, we apply the MS method to a set of 144 model protein sequences with 16 amino acids.
+This set (denoted here {\SR}) was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and a $\beta$-hairpin, respectively.~\cite{Bhattacherjee2012} In a previous work, this set allowed us to demonstrate the existence of A1-N1 mutational pathways which do not proceed through disordered (non-folding) sequences but instead switch abruptly between the two folds in a single mutational step.~\cite{Holzgrafe2014} Moreover, it allowed the exploration of the biophysical factors that underpin this ``fold switching" phenomenon,~\cite{Holzgrafe2015} which was recently demonstrated in a handful of natural and engineered proteins.~\cite{Bryan2010} Here we use the set {\SR} and our CG model for protein folding to validate the MS method and compare its efficiency to a standard generalized-ensemble method.~\cite{Marinari1992,Lyubartsev1992} We thereafter greatly enlarge {\SR} to a set with 1,024 sequences, also spanning the A1-N1 space, as well as another set of the same size spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively.
+Besides showing that the MS method can be applied to large numbers of sequences, the results allow us to carry out a more systematic analysis of the biophysical character of sequences along mutational pathways connecting these two pairs of basic folds than has been previously possible.
 
@@
 
-\subsection{Computational efficiency of the multisequence algorithm}
+\subsection{Computational efficiency}
 \noindent
-We start by testing the MS algorithm on the set {\SR} consisting of the 144 3-letter model sequences with 16 amino acids studied in Ref.~\citenum{Holzgrafe2014}.
-Two of the sequences in {\SR} are A1 and N1 (see Table~I) which were designed to fold into stable $\alpha$-helix and $\beta$-hairpin structures, respectively, as shown in Fig.~2B.
+We start by applying the MS algorithm to the set {\SR} across a range of temperatures $T$ (see Table~II).
+Two of the sequences in {\SR} are A1 and N1 (see Table~I) which fold into stable $\alpha$-helix and $\beta$-hairpin structures, respectively, as shown in Fig.~2B.
 A1 and N1 differ at 10 positions such that 10 consecutive point mutations can transform A1 into N1, and vice versa.
 The binary sequence space between A1 and N1 in which any combination of these mutations have been carried out, therefore contains $2^{10}=1,024$ sequences.
-The 144 sequences were selected from this binary space with the constraint that the total number of h-amino acids are not too high and not too unevenly distributed.~\cite{Holzgrafe2014}
+The 144 sequences in {\SR} were selected from this binary space with the constraints that the total number of hydrophobic amino acids are not too high and that they are not too unevenly distributed.~\cite{Holzgrafe2014}
 
@@
 
-We apply the multisequence algorithm on the set {\SR} across a range of temperatures (see Table~II).
-Figure~2 illustrates a typical MS simulation trajectory carried out at the lowest temperature which is well below the folding temperature of both A1 and N1~\cite{Holzgrafe2014,Holzgrafe2015}.
-From the MC evolution of the total energy $E$, sequence index $\sbar$, and RMSD values from the representative structures in Fig.~2B, it is evident that visits to various sequences drive transitions into a range of structural states.
-In particular, there are frequent visits to both $\alpha$-helix and $\beta$-hairpin structures.
-Transitions between them are accompanied by a shift in which sequences are preferably visited.
-For example, visits to high $\sbar$-indices, including N1 with index 144, tend to coincide with formation of $\beta$-hairpin structures.
+Figure~2 illustrates a typical MS simulation trajectory carried out at the lowest studied temperature which is below the folding temperature of both A1 and N1.~\cite{Holzgrafe2014,Holzgrafe2015} From the MC evolution of the total energy $E$, sequence index $\sbar$, and RMSD values from the representative structures in Fig.~2B, it is evident that visits to various sequences drive transitions into a range of structural states.
+In particular, there are frequent visits to both $\alpha$-helix and $\beta$-hairpin structures and transitions between them are accompanied by a shift in which sequences are preferably visited.
+For example, visits to high $\sbar$-indices, including N1 with index 144, tend to coincide with formation of $\beta$-hairpin structures as required to generate the correct equilibrium conformational ensembles.
 
 One might have suspected that the MS algorithm would be hampered by poor acceptance rates for sequence updates.
-However, this is not the case.
+However, this is not the case in our model.
 We carry out updates $\sbar\rightarrow\sbar'$ by picking a new random sequence $\sbar'\ne\sbar$ from the set of allowed sequences.
 The (average) acceptance rate  depends on both $T$ and the step in sequences space $\Delta h$, i.e., the number of amino acid positions changed, as shown in Fig.~3.
-At the lowest $T$ and high $\Delta h$, acceptance rates are only around 0.1-0.2 but are substantially higher at other $T$ and $\Delta h$ values.
-The overall acceptance rates are in fact above the oft-quoted rule-of-thumb value $0.25$~\cite{Gilks1996} for all temperatures $T$ (see Fig~4B).
-Increased acceptance rates can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer protein chains.
+At the lowest $T$ and highest $\Delta h$, acceptance rates are only around 0.1-0.2.
+However, for most other $T$ and $\Delta h$ the overall acceptance rate is substantially higher and often above the oft-quoted rule-of-thumb value 0.25~\cite{Gilks1996} (see Fig~4B).
+We note therefore that increased acceptance rates can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer chains.
 
@@
 
-We now compare the results from our MS calculations with simulated tempering (ST) simulations carried out on 4 of the 144 sequences, namely A1 and N1 and two sequences, R1 and R2, chosen randomly at distances $h=4$ and $h=6$ from A1, respectively (see Table~I).
-While ST provides the thermodynamics of a given sequence across a range of $T$ in a single run, a MS simulation provides the thermodynamics of all 144 sequences for one $T$.
+We now compare the results from our MS calculations with simulated tempering (ST) simulations carried out on 4 of the 144 sequences, namely A1 and N1 and two random sequences, R1 and R2, chosen at distances $h=4$ and $h=6$ from A1, respectively (see Table~I).
+While ST provides the thermodynamics of a given sequence across a range of $T$ in a single run, an MS simulation provides the thermodynamics of all 144 sequences at one $T$.
 We adjust the simulation lengths for ST and MS runs such that roughly the same number of sampled conformations are obtained for each $\sbar$ and $T$ combination, thus ensuring that similar computational resources are used for the two algorithms (see Table~II).
-The MS algorithm is validated by comparing the average total energy, $\left < E \right >$, calculated for different $T$'s and $\sbar$'s, between the two sets of results (see Supplementary Information).
+We first validate the MS algorithm by comparing the average total energy, $\left < E \right >$, calculated for these 4 sequences with the two different methods (see Supplementary Information).
+The two sets of results are entirely consistent showing that, for a given $\sbar$ and $T$, both MS and ST indeed sample the same distribution.
 
-As a way to assess sampling efficiency, we compare in Figure~4 the statistical error, $\sigE$, of the average energy $\left <E\right >$ for the 4 different sequences obtained using ST and MS, respectively.
-Since roughly the same number of sampled conformations were obtained for each $\sbar$ and $T$ combination, we compare the statistical errors directly.
-The two algorithms give similar errors at the highest $T$ studied which implies consistent result between them.
-It should be noted that the highest $T$ studied is well above the folding temperature for all the 4 sequences.
-Therefore, this consistency is not particularly surprising since at high-T the free-energy landscape is relatively smooth.
-Hence, conformational space requires little difficulty to sample.
+As a way to assess conformational sampling efficiency, we compare in Fig.~4 the statistical error, $\sigE$, of the average energy $\left <E\right >$ for the 4 sequences obtained using ST and MS, respectively.
+Because approximately the same number of sampled conformations were obtained for each combination of $\sbar$ and $T$, we compare the statistical errors directly.
+At the highest studied $T$, which is well above the folding temperature of both A1 and N1, the two algorithms give almost identical statistical errors.
+This can be understood by noting that at high-$T$ the free-energy landscape is relatively smooth and conformational space requires little difficulty to sample.
+The benefit of adding a dynamic parameter, whether $\sbar$ or $T$, is apparently minor under these conditions.
+However, at lower $T$, the $\sigE$ values from MS is often smaller than those from ST and never significantly higher.
+For example, at the lowest $T$, the precision in the estimate of $\left < E\right >$ is roughly twice as high in MS than ST for A1 and R1, and roughly the same for N1 and R2.
 
-However, as temperature decreases, the statistical error from the MS algorithm generally becomes smaller than those from ST.
-In particular, Fig.~4C illustrates that MS is on average twice as accurate as ST in its error estimates for R1.
-The best estimate occurs at $\kb\approx$ 0.46 where MS is about 450\% better than ST.
-Even at the lowest T studied, MS is at least consistent or better than ST for all 4 sequences.
+%The best estimate occurs at $\kb\approx$ 0.46 where MS is about 450\% better than ST. Even at the lowest T studied, MS is at least consistent or better than ST for all 4 sequences.
 
@@
 
-%Hence, we conclude that in the comparison made above the computational efficiency of the MS algorithm is comparable to ST.
-It must be stressed, however, that while the ST simulation requires that the thermodynamic behavior is calculated at a range of temperatures, MS works well on its own directly at low temperatures.
-ST was introduced as an extension to the standard, fixed-$T$ Metropolis-Hastings algorithm in order to overcome the poor mixing of states at low temperatures, i.e., the tendency for the simulations at low $T$ to become trapped in local free energy minima.
-The MS algorithm, however, efficiently calculates the thermodynamics at low $T$'s without visits to other $T$'s.
-In the MS algorithm, it is visits to sequences with poor native state stability rather than high $T$'s that drive the decorrelation of generated states and escape from local traps.
-This way, calculation of low-$T$ thermodynamics is enhanced in MS simulations while at the same time enhancing coverage in sequence space.
-This makes the MS algorithm ideal for situations in which the finite-$T$ thermodynamic behavior of multiple sequences are of interest.
-
 \vspace{12pt}
-\subsection{A1-N1 and A2-TN binary sequence space}
+\subsection{Exploring sequence space: IA/IB and IIA/IIB fold connectivities}
 \noindent
 We now turn to the full binary sequence sets {\SI} and {\SII} with 1,024 sequences each.
-By applying the MS method to these two sets separately (see Table~II), we determined the thermodynamic behavior of each included sequence at relatively low temperatures.
-In particular, these simulations allowed us to find the stabilities relative to folds IA and IB ($\PIA$ and $\PIB$) for all sequences in {\SI} and the stabilities relative to folds IIA and IIB ($\PIIA$ and $\PIIB$) for all sequences in {\SII}.
+By applying the MS method to these two sets (see Table~II), we determine the low-$T$ thermodynamic behavior of each included sequence.
+In particular, we calculate the stabilities of folds IA and IB, $\PIA$ and $\PIB$, for all sequences in {\SI} and the stabilities of folds IIA and IIB, $\PIIA$ and $\PIIB$, for all sequences in {\SII} (see Methods).
+The relative statistical errors on these quantities vary but are only a few percent at the most, despite the large number of sequences included.
 
-Having calculated these fold stabilities, we are in a position to determine if there are pathways in sequence space that lead to IA-IB or IIA-IIB fold changes without passing through an unstable intermediate sequence.
-To this end, we construct graphs in which each stable sequence is represented by a node and any two nodes are connected if their sequences differ at only one amino acid position.
-To determine if a sequence is stable we used the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PIA + \PIB$ and $\PIIA + \PIIB$ for A1-N1 and A2-TN, respectively.
-The precise network obtained depends, of course, on the cut-off value of $\Pcut$.
-In general, a higher $\Pcut$ means a selection of more stable pathways.
-Fig.~5 shows the networks obtained with $\Pcut=0.50$.
-The number of viable mutational pathways that connect the start and end points are 516,972 between A1 and N1 and  57,912 between A2 and TN.
+Having calculated these fold stabilities, we are in a position to determine if there are pathways in sequence space that lead to abrupt IA-IB or IIA-IIB fold changes, i.e., paths that do not pass through an unstable intermediate sequence.
+To this end, we construct graphs in which each stable sequence is represented by a node and any two nodes are connected if their sequences differ at a single amino acid position.
+To determine if a sequence is stable we use the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PIA + \PIB$ and $\PIIA + \PIIB$ for the IA-IB and IIA-IIB fold pairs, respectively; $\Ptot$ thus indicates the total stability of a sequence across the two competing folds.
+The precise network depends, of course, on the cut-off value of $\Pcut$ and a higher $\Pcut$ generally means a selection of more stable pathways.
+
+Fig.~5 illustrates the networks obtained with $\Pcut=0.50$ showing that, at this stability threshold, there are mutational pathways that connect both IA-IB and IIA-IIB.
+A precise analysis shows that there are 516,972 viable IA-IB paths and 57,912 viable IIA-IIB paths.
 Because there are $10!
-=3,628,800$ possible mutational pathways between start and end points in both cases, these pathways represent 14.2 \% and 1.6 \% of all possible pathways in A1-N1 and A2-TN, respectively.
-Hence, A1 and N1 are rather highly connected at this stability threshold.
-For $\Pcut=0.60$, the numbers are 104,640 pathways (2.9\%) for A1-N1 and 22,512 (0.6\%) pathways for A2-TN.
-We find that there are no possible A1-to-N1 or A2-to-TN pathways when $\Pcut\ge0.74$ and $\Pcut\ge 0.66$, respectively.
+=3,628,800$ possible paths between start and end points in both cases, they represent 14.2 \% and 1.6 \% of all possible paths of the A1-N1 and A2-TN sequence pairs, respectively.
+Hence, folds IA and IB are rather highly connected in our model for $\Pcut= 0.50$.
+For $\Pcut=0.60$, the numbers are 104,640 paths (2.9\%) for IA-IB and 22,512 (0.6\%) paths for IIA-IIB.
+We find that there are no possible IA-to-IB or IIA-to-IIB paths when $\Pcut\ge0.74$ and $\ge 0.66$, respectively.
 \\
 
-\subsection{Stability and bistability properties in of fold-to-fold mutational pathways}
+\subsection{Biophysical properties of fold-to-fold mutational pathways}
 \noindent 
-An apparently general characteristic of designed and natural protein folding switching is that sequences near the switch point have reduced stability.~\cite{Bryan2010} The results from our model exhibit a similar trend.
-Fig.~6A and B show the total stability $\Ptot$ at different mutational distances between start and end points.
-Intermediate sequences are less stable than sequences at distances $h=0$ and $h=10$, although there are large variations as indicated by the spread between upper and lower bounds.
-Nonetheless, there is a clear statistical trend that sequences become gradually less stable as successive mutations are applied to any of the 4 start and end points until a minimum is reached.
+An apparently general characteristic of designed and natural proteins that exhibit mutation-induced fold switching is a reduced stability near the switch point.~\cite{Alexander2009,He2012,Kouza2012,Sikosek2016,Sutto2012} Our model proteins exhibit a similar trend.
+Fig.~6A and B show the average total stability $\Ptot$ for sequences found at different Hamming distances $h$ from the starting point.
+Intermediate sequences are less stable than sequences at distances $h=0$ (A1 or A2) and $h=10$ (N1 or TN), although there are large variations between sequences as indicated by the spread between upper and lower bounds.
+There is nonetheless a clear statistical trend that sequences become gradually less stable as successive mutations are applied to any of the 4 start and end points until a minimum is reached.
 %These minima in stability occur closer to N1 and TN in the two systems, respectively, indicating that the IA/IIA folds are more mutationally robust than the IB/IIB folds.
@@
 However, the smooth stability trends in Fig.~6A and B belie the real character of the individual mutational pathways which tend to exhibit an abrupt switch between the two folds.
-To examine the statistics of fold transitions in our model, we first make a distinction between two types of stable sequences: those that fold into a single unique fold, thus behaving as classical proteins, and those that exhibit significant stabilities of both folds.
-Such ``bistable" sequences are interesting from a biophysical perspective in that they alternate between populating two different folds and hence are highly dynamic proteins.
-Indeed, they have been proposed to play an important role in the evolution of new protein folds.~\cite{Sikosek2016}
-
-We consider a stable sequence to be bistable if $B>0.5$, where $B$ is a bistability measure (see Fig.~6 legend).
+To see this and to further examine the character of the fold transitions in our model, we make a distinction between two types of stable sequences: those that fold into a single unique fold, thus behaving as classical proteins, and those that display substantial stabilities of both folds.
+Such ``bistable" sequences are interesting from a biophysical perspective in that they might continuously alternate between two different folds in a highly dynamic and coordinated fashion.
+Indeed, bistable sequences have been proposed to play a role in the evolution of new protein folds.~\cite{Sikosek2012} We consider a stable sequence to be bistable if $B>0.5$, where $B$ is a bistability measure (see Fig.~6 legend).
 In principle, a fold transition can occur directly between two classical proteins with unique folds, or it can proceed via one or more intermediate bistable sequences which populate both folds.
@@
 Hence, a path with $L_\mathrm{s}=2$ accomplishes a fold switch in a single mutational step without going through a bistable point.
-From the distributions of  $L_\mathrm{s}$ in Fig.~5C and D taken over all pathways with $\Pcut=0.50$, it can be seen that fold switching along individual pathways are typically completed in only 1-2 mutations and often a single step is enough switch between the IA and IB folds.
+From the distributions of  $L_\mathrm{s}$ in Fig.~5C and D, taken over all pathways with $\Pcut=0.50$, it can be seen that fold switching along individual pathways are typically completed in only 1-2 mutations and a single step is often sufficient to switch between the IA and IB folds.
 Hence, fold switching is typically abrupt and, for $\Pcut = 0.5$, it is fairly common that viable pathways pass through one or two bistable sequences.
 
-Interestingly, IA-IB fold switching through one or more bistable sequences become less and less common as selections for more stable pathways are made.
+Interestingly, switching between folds IA and IB through one or more bistable sequences become less and less frequent as selections for more stable pathways are made.
 This can be seen from the decrease in $\left <L_\mathrm{s}\right >$ as a function of $\Pcut$ (see Fig.~5C(inset)).
-For $\Pcut\ge0.70$, there are no longer any viable pathways between the $\alpha$-helix and $\beta$-hairpin that passes through a bistable sequence because $\left <L_\mathrm{s}\right > =  2$.
+For $\Pcut\ge0.70$, there are no longer any remaining path between the $\alpha$-helix and $\beta$-hairpin that passes through a bistable sequence because $\left <L_\mathrm{s}\right > =  2$.
 An underlying reason for the occurrence of sharper fold switches for more stable mutational pathways is apparent from a comparison between $\Ptot$ and $B$ across all sequences in {\SI}.
 As shown in Fig.~5E, sequences with the highest $\Ptot$ tend to exhibit very little bistability.
-Hence, highly stable mutational pathways are therefore forced to go through abrupt switch points, thereby transitioning between folds in a single mutational step.
-The situation for the IIA-IIB fold switch is more complicated.
-Sequences in {\SII} too follow the trend that high overall stability is achieved for only classical proteins with a single dominant fold.
-However, bistable sequences nonetheless play a crucial role in bridging the IIA-IIB folds because paths without bistable intermediate sequence are not available for $\Pcut\ge0.50$.
-Hence, passing through sequences with bistable properties may still be required to achieve mutation-induced fold switching between certain fold pairs even through it lead to a reduced stability at the switch point.
+Hence, highly stable paths are therefore forced to go through abrupt switch points where they transition directly between folds in a single step.
+The situation for the IIA-IIB fold pair is more complicated.
+We find that, just as for {\SI}, sequences in {\SII} follow the trend that the highest $\Ptot$ values occur for only classical, low-$B$ proteins.
+One might therefore expect that selection of more stable IIA-IIB paths would decrease $\left <L_\mathrm{s}\right >$, however, this is not the case as such abrupt switch points between the IIA and IIB are not available for $\Pcut\ge0.50$ (cf.
+Fig.~5 bottom).
+As a result, bistable sequences do play a crucial role in bridging the IIA and IIB folds, although passing though these sequences lead to additional reduction in stability at the switch point.
+
+%.
+Hence, passing through sequences displaying substantial bistability may still be required to mutationally drive switching between certain fold pairs as single-step transitions are not available, thereby
 
@@
 \noindent
-The simulation algorithm for biomolecules considered here works by making the sequence a dynamic parameter.
-We tested this approach by applying it to proteins.
-The results indicate that there are two main potential benefits.
-Firstly, it provides a convenient way to sample the respective canonical distributions for a large set of sequences in a single run and, secondly, it enhances the sampling of conformational space such that it can be applied directly at low temperatures.
-
+We have evaluated a biomolecular simulation algorithm that works by making the sequence a dynamic parameter by testing it on a CG model for protein folding.
+The results indicate that there are two main potential benefits of this approach.
+Firstly, it provides a convenient way to sample the canonical distributions of a large set of sequences in a single run and, secondly, it enhances the sampling of conformational space such that it can be applied directly at low temperatures.
 The conformational sampling efficiency can be assessed from the comparison with ST.
 Although there is no single ``fair" way to compare the MS and ST methods, we chose as a measure of efficiency the statistical error of the total energy, $\sigE$, obtained with roughly the same computational cost per temperature and sequence.
-At the highest studied temperatures, which are well above the folding temperatures of both A1 and N1, the statistical errors $\sigE$ are basically the same for the two methods.
+At the highest studied temperatures, we find that the statistical errors $\sigE$ are basically the same for the two methods.
 This finding is not unexpected because conformational sampling of short polymers at high $T$ does not involve crossing any major energy barriers.
@@
 At lower temperatures, we find that the MS simulations often yields significantly smaller $\sigE$ than ST simulations.
-It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant temperature.
-Hence, rather than promoting escape from local minima by visits to higher temperature, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences where the minima are not as deep or maybe altogether absent.
-For the same reason, we expect the performance of the MS algorithm to depend on the size the sequence set used, as well as their character.
-Specifically, the MS simulation may hinge on the inclusion of at least a few sequences with poor stabilities such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
+It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant $T$.
+Hence, rather than promoting escape from local minima by visits to higher $T$, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences where the minima are not as deep or may be altogether absent.
+For the same reason, we expect the performance of the MS algorithm to depend on the size the sequence set used, as well as their conformational properties.
+Specifically, the performance of MS simulations may hinge on the inclusion of at least a few sequences with poor stabilities such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
 
-The MS method should not be seen as a general method to speed up conformational sampling.
-However, our results indicate that for coarse-grained models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be highly efficient.
-The ability to promote conformational sampling without resorting to an increased $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers, where escape from local minima can be especially challenging~\cite{Bereau2015} and elevated $T$ is typically avoided because it can lead to unwanted perturbations of the membrane structure.~\cite{Ulmschneider2010}
+We emphasize that the MS method should not be seen as a general technique to speed up conformational sampling.
+However, our results indicate that for CG models that permit sequence updates to be carried out as a Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be a highly efficient way to sample the equilibrium behavior of many sequences.
+In addition, the ability to promote conformational sampling without resorting to an increased $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
 
@@
 \noindent
-We have evaluated an algorithm for biomolecular simulations that allow the thermodynamics of multiple different sequences to be calculated in a single run.
-Here we applied this algorithm to protein folding and showed that the low-temperature thermodynamic behavior of over 1,000 amino acid sequences with each 35 amino acids could be determined in an intermediate-resolution CG model with 3 amino acid types.
-In addition to being convenient for determination of the thermodynamics of multiple sequences, we found that the method also enhances conformational sampling such that it can be applied also at low temperatures.
-A random walk in sequence space allows for efficient escape from the the minima present in the free energy landscapes of individual sequences.
-The method might be well suited for CG simulations of other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
+We have evaluated an algorithm for biomolecular simulations that allows the thermodynamics of multiple  sequences to be calculated in a single run.
+We applied the algorithm to protein folding and showed that the thermodynamic behavior of $>$1,000 amino acid sequences with up to 35 amino acids could be determined in an intermediate-resolution CG model.
+The method performs a random walk in sequence space which is especially useful at low temperature as it promotes escape from local minima present in the free energy landscapes of individual sequences.
+The method might therefore be well suited for CG simulations of other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
 

### 2017-06-07 Manuscript.tex
Minor final changes

@@
 \begin{abstract}
-We consider a generalized-ensemble algorithm for coarse-grained simulations of biomolecules which allows the thermodynamic behavior of two or more sequences to be determined in a single ``multisequence" run.
+We consider a generalized-ensemble algorithm for coarse-grained simulations of biomolecules which allows the thermodynamic behavior of two or more sequences to be determined in a single multisequence run.
 By carrying out a random walk in sequence space, the method also enhances conformational sampling.
 Escape from local energy minima is accelerated by visiting sequences for which the minima are more shallow or absent.
-We apply the method alongside an intermediate-resolution coarse-grained model for protein folding with 7 atoms per amino acid and 3 amino acid types.
-The potential for large-scale coverage of sequence space is explored by applying the method to two sets with $>$1,000 sequences each.
-The resulting thermodynamic data is used to explore the sequence-structure and sequence-stability relationships between pairs of protein folds with different secondary structures.
+We test the method on an intermediate-resolution coarse-grained model for protein folding with 3 amino acid types and explore the potential for large-scale coverage of sequence space by applying the method to sets of more than 1,000 sequences.
+The resulting thermodynamic data is used to analyze the structures and stability properties of sequences covering the space between folds with different secondary structures.
 \end{abstract}
@@
 
-The challenges of achieving representative conformational sampling of individual biomolecular systems notwithstanding, many biological processes naturally call for the investigation and comparison of molecular variants.
-This is required, for example, in determining the molecular mechanisms of specificity in protein-protein~\cite{Zarrinpar2003,Hakes2007} or protein-nucleotide interactions,~\cite{Rohs2010} and the role of mutations in disease processes such as protein aggregation.~\cite{Ross2004} Another example is the protein folding process, where unique insight has been achieved by comparing the folding within and between protein families.~\cite{Tzul2017,Wensley2010} In a situation with rapid growth of sequence information,~\cite{Vukmirovic2000} it is of interest to explore ways to efficiently sample multiple sequences in biomolecular simulations.
+The challenges of achieving representative conformational sampling of individual biomolecular systems notwithstanding, many biologically motivated problems naturally call for the investigation and comparison of molecular variants, e.g., determining the molecular mechanisms of specificity in protein-protein~\cite{Zarrinpar2003,Hakes2007} and protein-nucleotide~\cite{Rohs2010} interactions, or the role of mutations in molecular disease processes.~\cite{Ross2004} Another example is protein folding, where unique insight has been achieved by comparing sequences within and between protein families.~\cite{Tzul2017,Wensley2010} In a situation with extremely rapid growth of sequence information,~\cite{Vukmirovic2000} it is of interest to explore ways to efficiently sample multiple sequences in biomolecular simulations.
 
-To this end, we consider in this work an MC-based algorithm that can calculate the thermodynamics of multiple sequences in a single run and apply it to a coarse-grained model for protein folding.
-This multisequence (MS) method was originally developed in the context of homo- and heteropolymer simulations~\cite{Irback1995} and was also later adapted for the characterization of peptide-protein binding specificity.~\cite{Bhattacherjee2013,Wallin2017} To our knowledge, however, it has not been previously tested in realistic protein folding simulations.
+To this end, we consider in this work an MC-based algorithm that can calculate the thermodynamics of multiple sequences in a single run and apply it to a coarse-grained model for protein folding.~\cite{Bhattacherjee2012} This multisequence (MS) method was originally developed in the context of homo- and heteropolymer simulations~\cite{Irback1995} and was later adapted for the characterization of peptide-protein binding specificity.~\cite{Bhattacherjee2013,Wallin2017} To our knowledge, it has not been previously tested in realistic protein folding simulations.
 The MS algorithm carries out a simulation in a generalized ensemble that performs a random walk in sequence space.
 Hence, there are two main types of updates: conformational updates $\rbar\rightarrow\rbar'$ and sequence updates $\sbar\rightarrow\sbar'$.
-As illustrated in Fig.~1, this strategy can be applied when $\rbar$ and $\sbar$ are ``perpendicular" coordinates such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$.
-While this may be hard to achieve for detailed biomolecular models, it is likely possible in many CG models.
+This strategy is straightforward when $\rbar$ and $\sbar$ are ``perpendicular" coordinates, as illustrated in Fig.~1,  such that the potential energy of the model can be written in terms of two independent variables, $E(\sbar,\rbar)$.
+
+%While this may be hard to achieve for detailed biomolecular models, it is likely possible in many CG models.
 
 As an initial test case, we apply the MS method to a set of 144 model protein sequences with 16 amino acids.
-This set (denoted here {\SR}) was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and a $\beta$-hairpin, respectively.~\cite{Bhattacherjee2012} In a previous work, this set allowed us to demonstrate the existence of A1-N1 mutational pathways which do not proceed through disordered (non-folding) sequences but instead switch abruptly between the two folds in a single mutational step.~\cite{Holzgrafe2014} Moreover, it allowed the exploration of the biophysical factors that underpin this ``fold switching" phenomenon,~\cite{Holzgrafe2015} which was recently demonstrated in a handful of natural and engineered proteins.~\cite{Bryan2010} Here we use the set {\SR} and our CG model for protein folding to validate the MS method and compare its efficiency to a standard generalized-ensemble method.~\cite{Marinari1992,Lyubartsev1992} We thereafter greatly enlarge {\SR} to a set with 1,024 sequences, also spanning the A1-N1 space, as well as another set of the same size spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively.
-Besides showing that the MS method can be applied to large numbers of sequences, the results allow us to carry out a more systematic analysis of the biophysical character of sequences along mutational pathways connecting these two pairs of basic folds than has been previously possible.
+This set (denoted here {\SR}) was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and a $\beta$-hairpin,~\cite{Bhattacherjee2012} respectively, as shown in Fig.~2.
+In a previous work, this set allowed us to demonstrate the existence of A1-N1 mutational pathways which do not proceed through any disordered (non-folding) sequences but instead switch abruptly between the two folds in a single mutational step.~\cite{Holzgrafe2014} Such an ability to reversibly switch folds was recently demonstrated in a handful of natural and engineered proteins.~\cite{Bryan2010} Here we use the set {\SR} and our CG model for protein folding to validate the MS method and compare its efficiency to a standard generalized-ensemble method.~\cite{Marinari1992,Lyubartsev1992} We thereafter greatly enlarge {\SR} to a set with 1,024 sequences, also spanning the A1-N1 space, as well as another set of the same size spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively (see Fig.~2).
+Besides demonstrating that the MS method can be applied to large numbers of sequences, the results allow us to carry out a more systematic analysis of the biophysical properties of sequences along mutational pathways connecting these two pairs of basic folds than has been previously possible.
 
@@
 Conventional Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
-The basic idea of generalized-ensemble algorithms~\cite{Mitsutake2001} is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters, such that a more efficient random walk in potential energy can be achieved.
+The basic idea of generalized-ensemble algorithms~\cite{Mitsutake2001} is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters.
 
-A well-known generalized-ensemble algorithm is simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} in which it is the temperature $T$ that becomes a dynamic parameter.
-In this scheme, frequent visits to high-$T$ allow simulations to readily escape from local energy traps.
-The ST algorithm thus simulates the joint probability distribution
+A well-known generalized-ensemble algorithm is simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} where it is the temperature that becomes a dynamic parameter.
+In this scheme, frequent visits to high $T$ allow simulations to readily escape from local energy minima.
+The ST algorithm simulates the joint probability distribution
 \begin{equation}
@@
 The basic idea of the MS algorithm for biomolecular simulation is to let the sequence $\sbar$ become a dynamic parameter rather than the temperature as in ST.
-A dynamic $\sbar$ is technically feasible as long as the potential energy can be written as $E(\sbar,\rbar)$, where $\sbar$ and $\rbar$ are independent variables.
-This is the case in our coarse-grained protein model with only backbone degrees of freedom and can also be achieved for more detailed  models~\cite{Bhattacherjee2013,Wallin2017}.
+A dynamic $\sbar$ is technically feasible when the potential energy can be written as $E(\sbar,\rbar)$, where $\sbar$ and $\rbar$ are independent variables.
+This is the case in our coarse-grained protein model which has only backbone degrees of freedom.
+It can also be achieved for in some more detailed models.~\cite{Bhattacherjee2013,Wallin2017}
 
-In analogy with ST, the MS algorithm simulates the joint probability distribution
+Similarly to ST, the MS algorithm simulates the joint probability distribution
 \begin{equation}
-P(\sbar,\rbar) =\dfrac{1}{\tilde{Z}}e^{-\beta E(\sbar,\rbar) + h(\sbar)}\,, 
+P(\sbar,\rbar) =\dfrac{1}{Z}e^{-\beta E(\sbar,\rbar) + h(\sbar)}\,, 
 \label{MS}
@@
 \begin{equation}
-\tilde{Z} = \sum_{\sbar}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)}\,
+Z = \sum_{\sbar}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)}\,
 \end{equation}
 and the first sum goes over a set of allowed sequences $\sbar$.
-The simulation parameters $h(\sbar)$, similar to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=\tilde{Z}^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
+The simulation parameters $h(\sbar)$, similar to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=Z^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
 
@@
 This geometrical description is paired with a simplified but finely tuned energy function with 4 terms: $E= E_\mathrm{ev} + E_\mathrm{loc} + E_\mathrm{hb} + E_\mathrm{hp}$.
-The first two, $E_\mathrm{ev}$ and $E_\mathrm{loc}$, represent excluded-volume effects of all atoms and local electrostatic effects, respectively.
-The hydrogen-bond energy, $E_\mathrm{hb}$, represents directionally dependent interactions between NH and CO groups which are necessary for secondary structure formation.
+The first two, $E_\mathrm{ev}$ and $E_\mathrm{loc}$, represent excluded-volume effects and local electrostatic effects, respectively.
+The hydrogen-bond energy, $E_\mathrm{hb}$, represents directionally dependent interactions between NH and CO groups and is necessary for secondary structure formation.
 Finally, the ``hydrophobicity" term, $E_\mathrm{hp}$, implements pairwise Lennard-Jones-like interactions between the $C_\beta$ atoms of h amino acids which are necessary for driving chain collapse during folding.
@@
 The fold IA-hydrogen bonds are (2,6), (3,7), (4,8), (5,9), (6,10), (7,11), (8,12), (9,13), (10,14), (11,15) and the fold IB-bonds are (3,14), (5,12), (7,10), (10,7), (12,5), (14,3), where (i,j) indicates a hydrogen bond between the CO group of amino acid i and the NH group of amino acid j.
-The  stabilities of folds IA and IB are then defined as the probabilities $\PIA = P(\QIA\ge0.8)$ and $\PIB = P(\QIB\ge0.80)$, respectively, i.e., the probability that at least 80\% of the fold's hydrogen bonds are formed.
+The stabilities of folds IA and IB are defined as the probabilities $\PIA = P(\QIA\ge0.8)$ and $\PIB = P(\QIB\ge0.8)$, respectively, i.e., the probability that at least 80\% of the fold's hydrogen bonds are formed.
 $\PIA$ and $\PIB$ thus depend on both sequence $\sbar$ and temperature $T$.
 For example, $\PIA=0.875\pm 0.003$ for A1 and $\PIB=0.785\pm0.008$ for N1 at $\kb T = 0.43$.
-Structural similarity measures for 35-amino acid folds IIA and IIB are defined as $\QIIA = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IA}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$ and $\QIIB = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IB}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$, respectively, where superscripts on $\QIA$ and $\QIB$ indicate over which amino acid positions those measures are applied to within the longer 35 amino acid sequences and $Q_\mathrm{tert}$ is a measure that counts the number of $\mathrm{C}_\beta$-$\mathrm{C}_\beta$ contacts between the two secondary structure elements of these folds.~\cite{Holzgrafe2015} In analogy with $\PIA$ and $\PIB$, we define the stabilities of folds IIA and IIB as $\PIIA = P(\QIIA\ge0.8)$ and $\PIIB = P(\QIIB\ge0.80)$, respectively.
+Structural similarity measures for 35-amino acid folds IIA and IIB are defined as $\QIIA = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IA}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$ and $\QIIB = ( Q_\mathrm{IA}^\mathrm{1-16} + Q_\mathrm{IB}^\mathrm{20-35} + Q_\mathrm{tert} ) / 3$, respectively, where superscripts on $\QIA$ and $\QIB$ indicate over which amino acid positions those measures are applied to within the longer 35 amino acid sequences and $Q_\mathrm{tert}$ is a measure that counts the number of $\mathrm{C}_\beta$-$\mathrm{C}_\beta$ contacts between the two secondary structure elements of these folds.~\cite{Holzgrafe2015} In analogy with $\PIA$ and $\PIB$, we define the stabilities of folds IIA and IIB as $\PIIA = P(\QIIA\ge0.8)$ and $\PIIB = P(\QIIB\ge0.8)$, respectively.
 The root-mean-square-deviation, RMSD, is calculated over all $\mathrm{C}_\alpha$ atoms.
@@
 \begin{figure}
-\includegraphics[width=10.0cm]{MCTrajFolds}
+\includegraphics[width=8.3cm]{Fig2}
 \caption{(A) Example of an MS simulation applied to the sequence set $\mathrm{S16}_{144}$.
@@
 Two of the sequences in {\SR} are A1 and N1 (see Table~I) which fold into stable $\alpha$-helix and $\beta$-hairpin structures, respectively, as shown in Fig.~2B.
-A1 and N1 differ at 10 positions such that 10 consecutive point mutations can transform A1 into N1, and vice versa.
+Because A1 and N1 differ at 10 positions, 10 consecutive point mutations can transform A1 into N1, and vice versa.
 The binary sequence space between A1 and N1 in which any combination of these mutations have been carried out, therefore contains $2^{10}=1,024$ sequences.
-The 144 sequences in {\SR} were selected from this binary space with the constraints that the total number of hydrophobic amino acids are not too high and that they are not too unevenly distributed.~\cite{Holzgrafe2014}
+The 144 sequences in {\SR} were selected from this binary space with the constraints that the total number of hydrophobic amino acids are not too high and that they are not too unevenly distributed along the sequence.~\cite{Holzgrafe2014}
 
 \begin{figure}
-\includegraphics[width=7.8cm]{Pacc}
+\includegraphics[width=7.8cm]{Fig3}
 \caption{Acceptance rates for $\sbar\rightarrow\sbar'$ updates in MS simulations of the {\SR} sequence set as a function of (A) the number of changed amino acid positions $\Delta h$ and (B) temperature $T$.
@@
 Figure~2 illustrates a typical MS simulation trajectory carried out at the lowest studied temperature which is below the folding temperature of both A1 and N1.~\cite{Holzgrafe2014,Holzgrafe2015} From the MC evolution of the total energy $E$, sequence index $\sbar$, and RMSD values from the representative structures in Fig.~2B, it is evident that visits to various sequences drive transitions into a range of structural states.
-In particular, there are frequent visits to both $\alpha$-helix and $\beta$-hairpin structures and transitions between them are accompanied by a shift in which sequences are preferably visited.
-For example, visits to high $\sbar$-indices, including N1 with index 144, tend to coincide with formation of $\beta$-hairpin structures as required to generate the correct equilibrium conformational ensembles.
+In particular, there are frequent visits to $\alpha$-helix and $\beta$-hairpin structures and transitions between them are accompanied by a shift in which sequences are preferably visited.
+For example, visits to high $\sbar$-indices, including N1 with index 144, tend to coincide with formation of $\beta$-hairpin structures as required to generate correct equilibrium conformational ensembles.
 
@@
 However, for most other $T$ and $\Delta h$ the overall acceptance rate is substantially higher and often above the oft-quoted rule-of-thumb value 0.25~\cite{Gilks1996} (see Fig~4B).
-We note therefore that increased acceptance rates can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer chains.
+An increased acceptance rate can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer chains.
 
 \begin{figure}
-\includegraphics[width=8.0cm]{Stderr}
+\includegraphics[width=8.0cm]{Fig4}
 \caption{Comparing sampling efficiency of the MS and ST algorithms.
@@
 We first validate the MS algorithm by comparing the average total energy, $\left < E \right >$, calculated for these 4 sequences with the two different methods (see Supplementary Information).
-The two sets of results are entirely consistent showing that, for a given $\sbar$ and $T$, both MS and ST indeed sample the same distribution.
+The two sets of results are entirely consistent showing that, for a given $\sbar$ and $T$, the MS and ST algorithms indeed sample the same distribution.
 
@@
 This can be understood by noting that at high-$T$ the free-energy landscape is relatively smooth and conformational space requires little difficulty to sample.
-The benefit of adding a dynamic parameter, whether $\sbar$ or $T$, is apparently minor under these conditions.
+The benefit of adding a dynamic parameter, whether $\sbar$ or $T$, is apparently minimal under these conditions.
 However, at lower $T$, the $\sigE$ values from MS is often smaller than those from ST and never significantly higher.
@@
 \begin{figure*}
-\rotatebox{0}{\includegraphics[width=16.6cm]{network}}
+\rotatebox{0}{\includegraphics[width=16.6cm]{Fig5}}
 \caption{Networks of sequences connecting folds IA and IB (top) and folds IIA and IIB (bottom).
@@
 
-Having calculated these fold stabilities, we are in a position to determine if there are pathways in sequence space that lead to abrupt IA-IB or IIA-IIB fold changes, i.e., paths that do not pass through an unstable intermediate sequence.
+Having calculated these fold stabilities, we are in a position to determine if there are pathways in sequence space that lead to abrupt IA-IB or IIA-IIB fold changes, i.e., paths that do not pass through any unstable intermediate sequence.
 To this end, we construct graphs in which each stable sequence is represented by a node and any two nodes are connected if their sequences differ at a single amino acid position.
 To determine if a sequence is stable we use the criterion $P_\mathrm{tot}>\Pcut$, where  $\Ptot = \PIA + \PIB$ and $\PIIA + \PIIB$ for the IA-IB and IIA-IIB fold pairs, respectively; $\Ptot$ thus indicates the total stability of a sequence across the two competing folds.
-The precise network depends, of course, on the cut-off value of $\Pcut$ and a higher $\Pcut$ generally means a selection of more stable pathways.
+The precise network depends, of course, on the cut-off value $\Pcut$ and a higher $\Pcut$ generally means a selection of more stable pathways.
 
-Fig.~5 illustrates the networks obtained with $\Pcut=0.50$ showing that, at this stability threshold, there are mutational pathways that connect both IA-IB and IIA-IIB.
+Fig.~5 illustrates the networks obtained with $\Pcut=0.50$ showing that both the IA-IB and IIA-IIB fold pairs are connected in sequence space at this stability threshold.
 A precise analysis shows that there are 516,972 viable IA-IB paths and 57,912 viable IIA-IIB paths.
-Because there are $10!
-=3,628,800$ possible paths between start and end points in both cases, they represent 14.2 \% and 1.6 \% of all possible paths of the A1-N1 and A2-TN sequence pairs, respectively.
+These paths represent 14.2 \% and 1.6 \% of all possible paths, respectively, because there are a total of $10!
+=3,628,800$ possible paths between start and end points in both cases.
 Hence, folds IA and IB are rather highly connected in our model for $\Pcut= 0.50$.
@@
 Fig.~6A and B show the average total stability $\Ptot$ for sequences found at different Hamming distances $h$ from the starting point.
-Intermediate sequences are less stable than sequences at distances $h=0$ (A1 or A2) and $h=10$ (N1 or TN), although there are large variations between sequences as indicated by the spread between upper and lower bounds.
+Intermediate sequences are less stable than sequences at distances $h=0$ (A1 or A2) and $h=10$ (N1 or TN), although there are large variations between sequences as indicated by the upper and lower bounds.
 There is nonetheless a clear statistical trend that sequences become gradually less stable as successive mutations are applied to any of the 4 start and end points until a minimum is reached.
@@
 \begin{figure}
-\includegraphics[width=7.8cm]{Paths}
+\includegraphics[width=7.8cm]{Fig6}
 \caption{Stability properties of mutational pathways.
@@
 To see this and to further examine the character of the fold transitions in our model, we make a distinction between two types of stable sequences: those that fold into a single unique fold, thus behaving as classical proteins, and those that display substantial stabilities of both folds.
-Such ``bistable" sequences are interesting from a biophysical perspective in that they might continuously alternate between two different folds in a highly dynamic and coordinated fashion.
-Indeed, bistable sequences have been proposed to play a role in the evolution of new protein folds.~\cite{Sikosek2012} We consider a stable sequence to be bistable if $B>0.5$, where $B$ is a bistability measure (see Fig.~6 legend).
-In principle, a fold transition can occur directly between two classical proteins with unique folds, or it can proceed via one or more intermediate bistable sequences which populate both folds.
+Such ``bistable" sequences are interesting from a biophysical perspective in that they are able to fold into two alternative folds.
+Indeed, bistable sequences have been proposed to play a role in the evolution of new protein folds.~\cite{Sikosek2012} We consider a sequence to be bistable if $B>0.5$, where $B$ is a bistability measure (see Fig.~6 legend).
+In principle, a fold transition can then occur directly between two classical proteins with unique native folds, or it can proceed via one or more intermediate bistable sequences which populate both folds.
 We define the switch length of a mutational pathway $L_\mathrm{s}=2+n_\mathrm{B}$, where $n_\mathrm{B}$ is the number of bistable sequences in between the two classical sequences that define the switch point.
@@
 
-%.
-Hence, passing through sequences displaying substantial bistability may still be required to mutationally drive switching between certain fold pairs as single-step transitions are not available, thereby
-
-%The sequences {\SII} do also follow this same trend (Fig.~5F).
-However, bistable sequences do play a crucial role in bridging the folds IIA and IIB even though their stabilities are significantly reduced, all stable mutational pathways must pass through one bistable sequences.
-\\
-
-%\subsection{Super-highways in sequence space}
-%
-%While there are a large number possible pathways (for $\Pcut<0.60$) in both systems, there are relatively few ways to pass between the two neutral nets.
-For $\Pcut=0.50$, we find that there are 109 nodes in the IA neutral net and 34 nodes in the IIB neutral net through which all viable pathways pass.
-For $\Pcut=0.60$, these numbers reduce to 14 and 10 nodes in the IA and IIA neutral nets, respectively.
-%
-
 \section{Discussion}
 \noindent
-We have evaluated a biomolecular simulation algorithm that works by making the sequence a dynamic parameter by testing it on a CG model for protein folding.
-The results indicate that there are two main potential benefits of this approach.
-Firstly, it provides a convenient way to sample the canonical distributions of a large set of sequences in a single run and, secondly, it enhances the sampling of conformational space such that it can be applied directly at low temperatures.
+We have evaluated a biomolecular simulation algorithm that works by making the biological sequence a dynamic parameter.
+As a test, we applied it on a CG model for protein folding.
+The results indicate that there are two main benefits of this approach.
+Firstly, it provides a convenient way to sample the canonical distributions of large numbers of sequences in a single run and, secondly, it enhances the sampling of conformational space meaning it can be applied directly at low temperatures.
 The conformational sampling efficiency can be assessed from the comparison with ST.
-Although there is no single ``fair" way to compare the MS and ST methods, we chose as a measure of efficiency the statistical error of the total energy, $\sigE$, obtained with roughly the same computational cost per temperature and sequence.
-At the highest studied temperatures, we find that the statistical errors $\sigE$ are basically the same for the two methods.
+Although there is no single ``fair" way to compare the two methods, we chose as a measure of efficiency the statistical error of the total energy, $\sigE$, obtained with roughly the same computational cost per temperature and sequence.
+At the highest studied temperatures, we find that the statistical errors $\sigE$ are basically the same.
 This finding is not unexpected because conformational sampling of short polymers at high $T$ does not involve crossing any major energy barriers.
-As a result, successively sampled conformations for a given combination of sequence $\sbar$ and temperature $T$, are therefore likely mostly uncorrelated in both methods leading to similar $\sigE$ values.
+As a result, successively sampled conformations for a given combination of sequence $\sbar$ and temperature $T$, are likely uncorrelated in both methods which leads to similar $\sigE$ values.
 
@@
 It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant $T$.
-Hence, rather than promoting escape from local minima by visits to higher $T$, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences where the minima are not as deep or may be altogether absent.
+Hence, rather than promoting escape from local minima by visits to higher $T$, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences where the minima are not as deep or even altogether absent.
 For the same reason, we expect the performance of the MS algorithm to depend on the size the sequence set used, as well as their conformational properties.
-Specifically, the performance of MS simulations may hinge on the inclusion of at least a few sequences with poor stabilities such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
+Specifically, the performance of MS simulations at low $T$ may hinge on the inclusion of at least a few sequences with poor stabilities such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
 
 We emphasize that the MS method should not be seen as a general technique to speed up conformational sampling.
-However, our results indicate that for CG models that permit sequence updates to be carried out as a Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be a highly efficient way to sample the equilibrium behavior of many sequences.
-In addition, the ability to promote conformational sampling without resorting to an increased $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
+However, our results indicate that for CG models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be a highly efficient way to sample the equilibrium behavior of many sequences.
+In addition, the ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
 
@@
 The method performs a random walk in sequence space which is especially useful at low temperature as it promotes escape from local minima present in the free energy landscapes of individual sequences.
-The method might therefore be well suited for CG simulations of other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
+The method might be suitable for CG simulations of other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
 

### 2017-06-07 Supplement.tex
Supplement minor change

@@
-\section{Supplementary Material}
+\documentclass[12]{article}
 
-\begin{figure}
-\includegraphics[width=14cm]{meanE}
-\caption{ Correlation of average energy as calculated by ST and MS algorithms for our 4 model sequences (A) A1, (B) N1, (C) R1, and (D) R2.
-Figure shows that  ST and MS are consistent in their calculations.
-The errors associated with both algorithms are very small such that they are not visible on the scale of the average energy.}
+\usepackage[left=3cm, right=2.5cm, top=2.5cm, bottom=3cm]{geometry}
+%\usepackage{graphicx}% Include figure files
+\usepackage[export]{adjustbox}
+
+\begin{document}
+
+\section*{\LARGE{\textbf{Supplementary Material}}}
+
+\begin{figure}[h]
+\includegraphics[width=0.8\textwidth,center]{meanE}
+\caption{ Comparison of the average energy, $\left < E\right >$, as calculated at 8 different temperatures by the ST and MS algorithms for the 4 model sequences (A) A1, (B) N1, (C) R1, and (D) R2.
+Statistical 1$\sigma$ errors, estimated from 32 independent runs for each method (see Table~II), are shown but are smaller than the plot symbols for all points.}
 \end{figure}
+
+\end{document}

### 2017-06-07 Supplement.tex
final changes

@@
 \usepackage[export]{adjustbox}
+\usepackage{caption}
+
+\renewcommand{\figurename}{FIG.~S}
 
@@
 \includegraphics[width=0.8\textwidth,center]{meanE}
-\caption{ Comparison of the average energy, $\left < E\right >$, as calculated at 8 different temperatures by the ST and MS algorithms for the 4 model sequences (A) A1, (B) N1, (C) R1, and (D) R2.
+\caption*{FIG.~S1.
+Comparison of the average energy, $\left < E\right >$, as calculated at 8 different temperatures by the ST and MS algorithms for the 4 model sequences (A) A1, (B) N1, (C) R1, and (D) R2.
 Statistical 1$\sigma$ errors, estimated from 32 independent runs for each method (see Table~II), are shown but are smaller than the plot symbols for all points.}

### 2017-07-18 Manuscript.tex
Revision for JCP

@@
 \usepackage{bm}% bold math
+\usepackage{color}
 %\usepackage[mathlines]{lineno}% Enable numbering of text and display math
@@
 \section{Introduction}
-\noindent
-Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} Together with the ever-increasing availability of computational resources, these advances have triggered a few major efforts~\cite{McGuffee2010,Miao2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the dynamics of biomolecular systems of various sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} and a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging and insightful, these large-scale simulations have also highlighted the fact that severe tradeoffs in size and time scales will likely persist for the foreseeable future.
 
-One way to expand the range of biomolecular simulations is to turn to coarse-grained (CG) models, where the basic aim is to simplify the physical description of interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list 4 main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that a CG representation of either the interaction potential or the molecular geometry often opens up for alternative sampling schemes beyond traditional molecular dynamics approaches, which can further speed up conformational sampling.
+Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} Together with the ever-increasing availability of computational resources, these advances have triggered a few major efforts~\cite{McGuffee2010,Miao2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the dynamics of biomolecular systems of various sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} and a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging and insightful, these large-scale simulations have also highlighted that severe tradeoffs in size and time scales will likely persist for the foreseeable future.
+
+One way to expand the range of biomolecular simulations is to turn to coarse-grained (CG) models, where the basic aim is to simplify the description of physical interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list 4 main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that a CG representation of either the interaction potential or the molecular geometry often opens up for alternative sampling schemes beyond traditional molecular dynamics approaches, which can further speed up conformational sampling.
 Examples of such sampling schemes include activation-relaxation kinetics,~\cite{Beland2011} discrete molecular dynamics~\cite{Proctor2011} and various Monte Carlo (MC)-based techniques such as cluster moves.~\cite{Vitalis2009}
 
-The challenges of achieving representative conformational sampling of individual biomolecular systems notwithstanding, many biologically motivated problems naturally call for the investigation and comparison of molecular variants, e.g., determining the molecular mechanisms of specificity in protein-protein~\cite{Zarrinpar2003,Hakes2007} and protein-nucleotide~\cite{Rohs2010} interactions, or the role of mutations in molecular disease processes.~\cite{Ross2004} Another example is protein folding, where unique insight has been achieved by comparing sequences within and between protein families.~\cite{Tzul2017,Wensley2010} In a situation with extremely rapid growth of sequence information,~\cite{Vukmirovic2000} it is of interest to explore ways to efficiently sample multiple sequences in biomolecular simulations.
+The challenges of achieving representative conformational sampling of individual biomolecular systems notwithstanding, many biologically motivated problems naturally call for the investigation and comparison of molecular variants, e.g., determining the molecular mechanisms of specificity in protein-protein~\cite{Zarrinpar2003,Hakes2007} or protein-nucleotide~\cite{Rohs2010} interactions, or the role of mutations in molecular disease processes.~\cite{Ross2004} Another example is protein folding, where unique insight has been achieved by comparing sequences within and between protein families.~\cite{Tzul2017,Wensley2010} In a situation with extremely rapid growth of sequence information,~\cite{Vukmirovic2000} it is of interest to explore ways to efficiently sample multiple sequences in biomolecular simulations.
 
@@
 
-%While this may be hard to achieve for detailed biomolecular models, it is likely possible in many CG models.
+{\color{red} As a test case for the MS algorithm, we selected the phenomenon of protein fold switching~\cite{Bryan2010} which has been recently demonstrated in a handful of natural and engineered proteins.
+These special proteins exhibit a unique ability to reversibly switch between different folds with accompanying changes in secondary structure, hydrophobic core packing and overall shape.
+Fold switching in proteins often plays a functional role.
+For example, rare transitions to an alternative fold in the protein KaiB provide a crucial time delay mechanism in the circadian clock cycle of cyanobacteria.~\cite{LiWang2015} Fold switching can be triggered by various events including changes to solution conditions,~\cite{Kuloglu2002} subdomain detachment,~\cite{Burmann2012} ligand binding~\cite{Luo2004} and point mutations.~\cite{Alexander2009} }
 
-As an initial test case, we apply the MS method to a set of 144 model protein sequences with 16 amino acids.
+{\color{red} The discovery that abrupt and major changes in the 3-dimensional structures of proteins can be driven by the accumulation of point mutations, in particular, holds implications for protein evolution as it suggests a simple mechanism of fold evolution.
+The A and B domains of protein G fold into 3$\alpha$ and 4$\beta+\alpha$ structures, respectively,~\cite{Alexander2009} and are thus unrelated in both structure and sequence.
+Yet Alexander \textit{et al.}~\cite{Alexander2009} demonstrated that, after extensive initial mutations to both domains that left their respective folds undisturbed, a single additional mutation triggered an abrupt fold switch.
+This possibility suggests a potential evolutionary relationship between these two domains despite the lack of detectable similarity in both sequence or structure in wild-type protein G.~\cite{He2012} It is still unclear, however, how common such fold-to-fold mutational pathways are in evolutionary processes.
+We showed previously~\cite{Holzgrafe2014,Holzgrafe2015} that, within the model framework provided by our CG model,~\cite{Bhattacherjee2012} there are mutational paths between elementary protein folds that do not proceed through any disordered (non-folding) sequences.
+Instead, these paths exhibit abruptly switching between folds in a single, or very few, mutational steps similarly to the A and B domains of protein G.
+}
+
+In demonstrating this fold-switching behavior in our model we characterized the folding of a set of 144 different model sequences with 16 amino acids.
+This set (denoted here {\SR}) was constructed to sparsely span the sequence space between two ideally designed sequences, A1 and N1, folding into an $\alpha$-helix and a $\beta$-hairpin, respectively, as shown in Fig.~2.
+Here we use the set {\SR} to validate the MS method and compare its computational efficiency to a standard generalized-ensemble method.~\cite{Marinari1992,Lyubartsev1992} We thereafter greatly enlarge {\SR} to a set with 1,024 sequences as well as another set of the same size spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively (see Fig.~2).
+Besides demonstrating that the MS method can be applied to large numbers of sequences, the results allow us to carry out a more systematic analysis of the biophysical properties of sequences along mutational pathways connecting these two pairs of basic folds than has been previously possible.
+
+%As an initial test case, we apply the MS method to a set of 144 model protein sequences with 16 amino acids.
 This set (denoted here {\SR}) was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and a $\beta$-hairpin,~\cite{Bhattacherjee2012} respectively, as shown in Fig.~2.
@@
 \begin{figure}
-\includegraphics[width=4.2cm]{Fig1}
+\includegraphics[width=4.2cm]{./Fig1}
 \caption{The two types of Monte Carlo updates in the multisequence Monte Carlo algorithm.}
@@
 \subsection{Generalized-ensemble algorithms and simulated tempering}
-\noindent
+
 Conventional Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
@@
 \subsection{Multisequence algorithm}
-\noindent 
+
 The basic idea of the MS algorithm for biomolecular simulation is to let the sequence $\sbar$ become a dynamic parameter rather than the temperature as in ST.
@@
 This is the case in our coarse-grained protein model which has only backbone degrees of freedom.
-It can also be achieved for in some more detailed models.~\cite{Bhattacherjee2013,Wallin2017}
+It can also be achieved in some more detailed models.~\cite{Bhattacherjee2013,Wallin2017}
 
@@
 \end{equation}
-where  
+where  {\color{red}
 \begin{equation}
-Z = \sum_{\sbar}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)}\,
+Z = \sum_{\sbar\in\mathrm{S}}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)}\,
+\label{eq:ZMS}
 \end{equation}
-and the first sum goes over a set of allowed sequences $\sbar$.
-The simulation parameters $h(\sbar)$, similar to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=Z^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
+and S is a set of pre-selected sequences, i.e., the sequences to which visits are allowed during a simulation.} The simulation parameters $h(\sbar)$, similar to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=Z^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
 
-Two types of MC updates are required to sample from the distrubution in Eq.~\ref{MS}, ordinary conformational update $\rbar\rightarrow\rbar'$ and mutational updates $\sbar\rightarrow\sbar'$.
+Two types of MC updates are required to sample from the distrubution in Eq.~\ref{MS}, ordinary conformational updates $\rbar\rightarrow\rbar'$ and {\color{red} sequence} updates $\sbar\rightarrow\sbar'$.
 The acceptance probabilities for the latter  becomes
@@
 \subsection{Coarse-grained 3-letter model for protein folding}
-\noindent
+
 All calculations were carried out using the coarse-grained model for protein folding developed in Ref.~\citenum{Bhattacherjee2012}.
@@
 \subsection{Model sequences}
-\noindent
+
 Six of the model sequences studied in this work, A1, N1, R1, R2, A2, and TN, are given in Table~I.
@@
 
-\subsection{Monte Carlo simulation parameters}
-\noindent
-Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assign a new value between $-\pi$ and $\pi$; and (2) a semi-local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2001} In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out by randomly picking a new sequence $\sbar'\ne\sbar$ and applying the accept-reject criterion in Eq.~\ref{accrej}.
-A sequence move is attempted every $1,000$ MC steps while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
-The simulations carried out in this work are summarized in Table~II.
+\subsection{Monte Carlo simulation parameters and updates}
+
+Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assign a new value between $-\pi$ and $\pi$; and (2) a semi-local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2001}
+
+%, and where S is the set of allowed (pre-selected) sequences, S (see Eq.~\ref{eq:ZMS})
+%i.e., given by the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$ angles,
+
+{\color{red}In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out in the following way.
+First, a new random sequence $\sbar$ is picked from the set of pre-selected (allowed) sequences S, such that $\sbar'\ne\sbar$.
+The new sequence $\sbar'$ therefore differs from $\sbar$ in one or more amino acid positions.
+Thereafter, the sidechains of the protein, which remains in an unchanged (backbone) conformation $\rbar$, is re-built according to the new sequence $\sbar'$.
+Practically this means that, at the position(s) where the amino acid type has changed, the sidechain is altered according to the type change.
+For example, if $\mathrm{p}\rightarrow\mathrm{t}$, the $\mathrm{C}_\beta$ atom is removed and replaced with an $\mathrm{H}_{\alpha 2}$ atom or, if $\mathrm{p}\rightarrow\mathrm{h}$, the $\mathrm{C}_\beta$ remains in place but its character is changed to hydrophobic.
+Finally, the change in total energy $\Delta E$ is calculated and the accept-reject criterion in Eq.~\ref{accrej} is applied.
+If rejected, the old state $(\sbar,\rbar)$ is restored.}
+
+A sequence update is attempted every $1,000$ MC steps while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
+{\color{red}The computational cost for sequence updates is somewhat higher than for temperature updates, which does not require any energy calculations and is thus extremely rapid.
+For the purpose of comparing computational efficiencies between ST and MS, we therefore chose sequence updates to be slightly less frequent than temperature updates.} All simulations carried out in this work are summarized in Table~II.
 
@@
 \subsection{Observables}
-\noindent
+
 Fold stabilities are calculated as in Ref.~\citenum{Holzgrafe2015} and described briefly below.
@@
 \begin{figure}
-\includegraphics[width=8.3cm]{Fig2}
-\caption{(A) Example of an MS simulation applied to the sequence set $\mathrm{S16}_{144}$.
-The plot shows the MC evolution of the sequence $\sbar$ (numbered 1--144), the total potential energy $E$ and the root-mean-square deviation (RMSD) calculated against fold IA (light blue) and fold IB (dark red) structures shown in (B).
-The simulation is carried out at $\kb T = 0.43$.
-(B) Representative structures of fold IA and fold IB, selected to be the minimum-energy conformations found for the sequences A1 and N1, respectively.
-(C) Representative fold IIA and fold IIB structures selected to be the minimum-energy conformations of A2 and TN, respectively.
+\includegraphics[width=8.5cm]{Fig2}
+\caption{(A) Example of an MS simulation of the sequence set $\mathrm{S16}_{144}$ carried out at $\kb T = 0.43$.
+The plot shows the MC evolution of the sequence $\sbar$ (numbered 1--144), the total potential energy $E$ and the root-mean-square deviation (RMSD) calculated against the representative fold IA (light blue) and fold IB (dark red) structures in (B).
+Representative structures of folds (B) IA, IB, (C) IIA and IIB, chosen to be the minimum-energy conformations found for the sequences A1, N1, A2 and TN, respectively.
 }
@@
 \subsection{Computational efficiency}
-\noindent
+
 We start by applying the MS algorithm to the set {\SR} across a range of temperatures $T$ (see Table~II).
@@
 An increased acceptance rate can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer chains.
+{\color{red} For example, $\Delta h_\mathrm{max} = 1$ would be equivalent to allowing only (single-site) ``mutational" updates, i.e., picking randomly a position and changing the amino acid type at that position.}
 
@@
 \includegraphics[width=8.0cm]{Fig4}
-\caption{Comparing sampling efficiency of the MS and ST algorithms.
-Statistical errors $\sigma_{\left < E\right >}$ of the average total energy $\left < E\right >$ from the two methods for the 16-amino acid sequences (A) A1, (B) N1, (C) R1 and (D) R2.
+\caption{Comparing sampling efficiencies of the MS and ST algorithms.
+Statistical errors $\sigma_{\left < E\right >}$ of the average total energy $\left < E\right >$ obtained for the sequences (A) A1, (B) N1, (C) R1 and (D) R2 (see Table~I) at different temperatures $T$.
 Simulation lengths in the two methods are adjusted such that the number of conformations sampled per sequence and temperature is roughly the same (see text).
@@
 \begin{figure*}
-\rotatebox{0}{\includegraphics[width=16.6cm]{Fig5}}
+\includegraphics[width=17.0cm]{Fig5}
 \caption{Networks of sequences connecting folds IA and IB (top) and folds IIA and IIB (bottom).
@@
 \subsection{Exploring sequence space: IA/IB and IIA/IIB fold connectivities}
-\noindent
+
 We now turn to the full binary sequence sets {\SI} and {\SII} with 1,024 sequences each.
@@
 \subsection{Biophysical properties of fold-to-fold mutational pathways}
-\noindent 
+
 An apparently general characteristic of designed and natural proteins that exhibit mutation-induced fold switching is a reduced stability near the switch point.~\cite{Alexander2009,He2012,Kouza2012,Sikosek2016,Sutto2012} Our model proteins exhibit a similar trend.
@@
 \section{Discussion}
-\noindent
+
 We have evaluated a biomolecular simulation algorithm that works by making the biological sequence a dynamic parameter.
@@
 It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant $T$.
-Hence, rather than promoting escape from local minima by visits to higher $T$, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences where the minima are not as deep or even altogether absent.
-For the same reason, we expect the performance of the MS algorithm to depend on the size the sequence set used, as well as their conformational properties.
-Specifically, the performance of MS simulations at low $T$ may hinge on the inclusion of at least a few sequences with poor stabilities such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
+{\color{red} Therefore, rather than promoting escape from local minima by visits to higher $T$, as in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986} MS simulations must escape the minima occurring for one sequence through visits to other sequences.
+How this process works can be envisioned by considering an MS simulation that is visiting a sequence $s$ and is trapped in a local minimum, requiring that a high free-energy barrier is overcome for escape.
+The trapped state could be, e.g., a compact $\beta$-sheet rich state with a particular register.
+Eventually, sequence updates will carry the simulation to other sequences $\hat{s}$ while still remaining in the trapped state.
+However, the free energy barrier of this trapped state might be substantially lower for $\hat{s}$ than for $s$, or the barrier may even be altogether absent if the trapped state is unstable for $\hat{s}$, leading to rapid escape from the minimum.
+} The above reasoning also implies that the performance of the MS algorithm likely depends on the size the sequence set S as well as the conformational properties of the sequences.
+Specifically, the performance of MS simulations of proteins at low $T$ may benefit from the inclusion of at least a few sequences with poor stability properties, such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
+
+{\color{red} Individual protein sequences that exhibit spontaneous transitions between widely different competing conformational states, such as fold switching proteins, are especially challenging to molecular simulation methods.
+One potential approach is to use Hamiltonian replica exchange to couple Go-like terms, i.e., artificial energy terms with a bias towards a given structure, with a physical force field.~\cite{Meinke2007,Chen2014,Bernhardt2016} In this approach, exchange moves ``feed" conformations into the replica of the physical force field, thereby enhancing sampling without introducing a bias in the simulation.~\cite{Bernhardt2016} Although we did not specifically study sampling efficiency of bistable sequences, which exhibit two basins of attractions, it is possible that a similar confomational ``feeding" take place in MS simulations, where the coupling occurs instead between different sequences rather than with different Go-type energy terms.
+}
 
 We emphasize that the MS method should not be seen as a general technique to speed up conformational sampling.
-However, our results indicate that for CG models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be a highly efficient way to sample the equilibrium behavior of many sequences.
-In addition, the ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
+However, our results indicate that for CG models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), our method can be a highly efficient way to sample the equilibrium behavior of many sequences.
+{\color{red} This opens up for various applications.
+For example, it might be used to aid computational studies into the sequence determinants of conformational properties of intrinsically disordered~\cite{Das2013} or  denatured~\cite{Meng2013} proteins.
+Moreover, the ability of the MS algorithm to efficiently probe sequence effects on native structure and stability, as demonstrated in this work, may make it useful for efforts underway to merge molecular simulation and population genetics techniques.~\cite{Serohijos2014} We note also that, while applied to proteins here, the MS algorithm is equally valid for other bio-macromolecules, including DNA and RNA.} The ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
 
 \section{Conclusion}
-\noindent
+
 We have evaluated an algorithm for biomolecular simulations that allows the thermodynamics of multiple  sequences to be calculated in a single run.
@@
 The method performs a random walk in sequence space which is especially useful at low temperature as it promotes escape from local minima present in the free energy landscapes of individual sequences.
-The method might be suitable for CG simulations of other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
+The method might be suitable for CG simulations of {\color{red} various} other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
+
+\section{Supplementary Materia}
+
+Validation of the MS algorithm.
+
+\section{Acknowledgments}
+
+This work was supported by grants from Memorial University and National Sciences and Engineering Research Council of Canada (NSERC).
 
 \bibliography{/Users/stefan/Documents/Manuscripts/Jshort,/Users/stefan/Documents/Manuscripts/MyRefs}
+%\input{Manuscript.bbl}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2017-07-19 Manuscript.tex
JCP revisions

@@
 
-{\color{red} As a test case for the MS algorithm, we selected the phenomenon of protein fold switching~\cite{Bryan2010} which has been recently demonstrated in a handful of natural and engineered proteins.
-These special proteins exhibit a unique ability to reversibly switch between different folds with accompanying changes in secondary structure, hydrophobic core packing and overall shape.
-Fold switching in proteins often plays a functional role.
-For example, rare transitions to an alternative fold in the protein KaiB provide a crucial time delay mechanism in the circadian clock cycle of cyanobacteria.~\cite{LiWang2015} Fold switching can be triggered by various events including changes to solution conditions,~\cite{Kuloglu2002} subdomain detachment,~\cite{Burmann2012} ligand binding~\cite{Luo2004} and point mutations.~\cite{Alexander2009} }
+{\color{red} As a test application of the MS algorithm, we selected the phenomenon of protein fold switching which recently was demonstrated in a handful of natural and engineered proteins.
+These special proteins exhibit a unique ability to reversibly switch between entirely different folds, with accompanying changes in secondary structure, hydrophobic core packing and overall shape.~\cite{Bryan2010} The fold switching transitions found in natural proteins typically play a functional role.
+For example, rare transitions to an alternative fold in the protein KaiB provide a crucial time delay mechanism in the circadian clock cycle of cyanobacteria.~\cite{LiWang2015} Fold switching can occur either spontaneously~\cite{Meier2007} or be triggered by various signals including changes to solution conditions,~\cite{Kuloglu2002} subdomain detachment,~\cite{Burmann2012} ligand binding~\cite{Luo2004} and point mutations.~\cite{Alexander2009} Computational studies, using  coarse-grained~\cite{Holzgrafe2014,Holzgrafe2015,Kouza2012,Sutto2012} or atomistic~\cite{Ramirez-Sarmiento2015,Sikosek2016,Hansen2013} models, have attempted to explain how proteins can exhibit multiple folding funnels and how they are altered in response to binding events or changes in sequence.}
 
-{\color{red} The discovery that abrupt and major changes in the 3-dimensional structures of proteins can be driven by the accumulation of point mutations, in particular, holds implications for protein evolution as it suggests a simple mechanism of fold evolution.
-The A and B domains of protein G fold into 3$\alpha$ and 4$\beta+\alpha$ structures, respectively,~\cite{Alexander2009} and are thus unrelated in both structure and sequence.
-Yet Alexander \textit{et al.}~\cite{Alexander2009} demonstrated that, after extensive initial mutations to both domains that left their respective folds undisturbed, a single additional mutation triggered an abrupt fold switch.
-This possibility suggests a potential evolutionary relationship between these two domains despite the lack of detectable similarity in both sequence or structure in wild-type protein G.~\cite{He2012} It is still unclear, however, how common such fold-to-fold mutational pathways are in evolutionary processes.
-We showed previously~\cite{Holzgrafe2014,Holzgrafe2015} that, within the model framework provided by our CG model,~\cite{Bhattacherjee2012} there are mutational paths between elementary protein folds that do not proceed through any disordered (non-folding) sequences.
-Instead, these paths exhibit abruptly switching between folds in a single, or very few, mutational steps similarly to the A and B domains of protein G.
-}
+{\color{red} The discovery that proteins can be driven to switch folds through an accumulation of point mutations, in particular, holds implications for protein evolution as it suggests a simple mechanism of fold evolution.~\cite{Sikosek2014} Alexander \textit{et al.} demonstrated that the similarly sized but structurally distinct A ($3\alpha$) and B ($4\beta+\alpha$) domains of protein G could, after extensive mutations leaving their respective folds undisturbed, be triggered to switch folds by applying one additional mutation, Y45L, located at the edge of the hydrophobic core in the B domain.~\cite{Alexander2009} This remarkable discovery suggests the possibility that the two domains might be evolutionary related despite a lack of detectable similarity in either sequence or structure in wild-type protein G,~\cite{He2012} although this has yet to be proven.
+Moreover, it is unclear how common such fold-to-fold transitions are and how they might occur in evolutionary processes.~\cite{Porter2015} In previous work~\cite{Holzgrafe2014,Holzgrafe2015} we showed that mutational paths with abrupt fold switching exist between two other pairs of smaller protein folds within the model framework provided by our CG model.~\cite{Bhattacherjee2012} }
 
-In demonstrating this fold-switching behavior in our model we characterized the folding of a set of 144 different model sequences with 16 amino acids.
+In demonstrating mutation-induced fold switching in our model we characterized the folding of a set of 144 different model sequences with 16 amino acids.
 This set (denoted here {\SR}) was constructed to sparsely span the sequence space between two ideally designed sequences, A1 and N1, folding into an $\alpha$-helix and a $\beta$-hairpin, respectively, as shown in Fig.~2.
 Here we use the set {\SR} to validate the MS method and compare its computational efficiency to a standard generalized-ensemble method.~\cite{Marinari1992,Lyubartsev1992} We thereafter greatly enlarge {\SR} to a set with 1,024 sequences as well as another set of the same size spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively (see Fig.~2).
-Besides demonstrating that the MS method can be applied to large numbers of sequences, the results allow us to carry out a more systematic analysis of the biophysical properties of sequences along mutational pathways connecting these two pairs of basic folds than has been previously possible.
-
-%As an initial test case, we apply the MS method to a set of 144 model protein sequences with 16 amino acids.
-This set (denoted here {\SR}) was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and a $\beta$-hairpin,~\cite{Bhattacherjee2012} respectively, as shown in Fig.~2.
-In a previous work, this set allowed us to demonstrate the existence of A1-N1 mutational pathways which do not proceed through any disordered (non-folding) sequences but instead switch abruptly between the two folds in a single mutational step.~\cite{Holzgrafe2014} Such an ability to reversibly switch folds was recently demonstrated in a handful of natural and engineered proteins.~\cite{Bryan2010} Here we use the set {\SR} and our CG model for protein folding to validate the MS method and compare its efficiency to a standard generalized-ensemble method.~\cite{Marinari1992,Lyubartsev1992} We thereafter greatly enlarge {\SR} to a set with 1,024 sequences, also spanning the A1-N1 space, as well as another set of the same size spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively (see Fig.~2).
 Besides demonstrating that the MS method can be applied to large numbers of sequences, the results allow us to carry out a more systematic analysis of the biophysical properties of sequences along mutational pathways connecting these two pairs of basic folds than has been previously possible.
@@
 
-Conventional Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
-The basic idea of generalized-ensemble algorithms~\cite{Mitsutake2001} is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters.
+Conventional {\color{red} Metropolis} Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
+The basic idea of generalized-ensemble algorithms is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters~\cite{Mitsutake2001} {\color{red} (for a recent historical account see Ref.~\citenum{Berg2017}).
+Generalized-ensemble methods that have been frequently used for biomolecular simulations include simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} replica exchange~\cite{Swendsen1986}  or parallel tempering,~\cite{Hansmann1997} and metadynamics.~\cite{Laio2002}}
 
-A well-known generalized-ensemble algorithm is simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} where it is the temperature that becomes a dynamic parameter.
-In this scheme, frequent visits to high $T$ allow simulations to readily escape from local energy minima.
-The ST algorithm simulates the joint probability distribution
+ST is a direct extension to the Metropolis algorithm in which the temperature $T$ becomes a dynamic parameter.
+In this way, frequent visits to high $T$ allow simulations to readily escape from local energy minima.
+The algorithm thus simulates the joint probability distribution
 \begin{equation}
@@
 A sequence update is attempted every $1,000$ MC steps while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
-{\color{red}The computational cost for sequence updates is somewhat higher than for temperature updates, which does not require any energy calculations and is thus extremely rapid.
-For the purpose of comparing computational efficiencies between ST and MS, we therefore chose sequence updates to be slightly less frequent than temperature updates.} All simulations carried out in this work are summarized in Table~II.
+{\color{red}The computational cost for sequence updates is somewhat higher than for temperature updates.
+The latter does not require any energy calculation and is thus extremely rapid.
+For the purpose of comparing computational efficiencies of ST and MS, we therefore chose sequence updates to be slightly less frequent than temperature updates while both are fairly frequent.} All simulations carried out in this work are summarized in Table~II.
 
@@
 \begin{tabular}{lcccr}
-Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes a thermalization step with $10^6$ MC steps/run.} &  Sequences\\
+Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes a thermalization step with {\color{red} $10^7$} MC steps/run.} &  Sequences\\
 \hline
@@
 At the lowest $T$ and highest $\Delta h$, acceptance rates are only around 0.1-0.2.
-However, for most other $T$ and $\Delta h$ the overall acceptance rate is substantially higher and often above the oft-quoted rule-of-thumb value 0.25~\cite{Gilks1996} (see Fig~4B).
+However, for most other $T$ and $\Delta h$ the overall acceptance rates are substantially higher and often above the oft-quoted rule-of-thumb value 0.25~\cite{Gilks1996} (see Fig~4B).
 An increased acceptance rate can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer chains.
-{\color{red} For example, $\Delta h_\mathrm{max} = 1$ would be equivalent to allowing only (single-site) ``mutational" updates, i.e., picking randomly a position and changing the amino acid type at that position.}
+{\color{red} For example, $\Delta h_\mathrm{max} = 1$ would be equivalent to applying only a ``mutational" update, i.e., picking a random (allowed) position and changing the amino acid type at that position.}
 
@@
 {\color{red} Individual protein sequences that exhibit spontaneous transitions between widely different competing conformational states, such as fold switching proteins, are especially challenging to molecular simulation methods.
-One potential approach is to use Hamiltonian replica exchange to couple Go-like terms, i.e., artificial energy terms with a bias towards a given structure, with a physical force field.~\cite{Meinke2007,Chen2014,Bernhardt2016} In this approach, exchange moves ``feed" conformations into the replica of the physical force field, thereby enhancing sampling without introducing a bias in the simulation.~\cite{Bernhardt2016} Although we did not specifically study sampling efficiency of bistable sequences, which exhibit two basins of attractions, it is possible that a similar confomational ``feeding" take place in MS simulations, where the coupling occurs instead between different sequences rather than with different Go-type energy terms.
+This problem has been addressed by using Hamiltonian replica-exchange techniques to couple G\={o}-like terms, i.e., energy terms with an artificial bias towards a given structure, with a physical force field.~\cite{Meinke2007,Chen2014,Bernhardt2016} This way exchange moves ``feed" conformations into the replica corresponding to the physical force field and enhance sampling.~\cite{Bernhardt2016} We did not specifically study sampling efficiency for sequences exhibiting competing states.
+It appears likely, however, that a very similar type of ``feeding" of conformations would take place in MS simulations although the coupling occurs instead with other sequences rather than with G\={o}-type terms.
 }
@@
 However, our results indicate that for CG models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), our method can be a highly efficient way to sample the equilibrium behavior of many sequences.
-{\color{red} This opens up for various applications.
-For example, it might be used to aid computational studies into the sequence determinants of conformational properties of intrinsically disordered~\cite{Das2013} or  denatured~\cite{Meng2013} proteins.
-Moreover, the ability of the MS algorithm to efficiently probe sequence effects on native structure and stability, as demonstrated in this work, may make it useful for efforts underway to merge molecular simulation and population genetics techniques.~\cite{Serohijos2014} We note also that, while applied to proteins here, the MS algorithm is equally valid for other bio-macromolecules, including DNA and RNA.} The ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
+{\color{red} This opens up for various applications, such as exploring sequence effects on the conformational properties of disordered~\cite{Das2013} or denatured~\cite{Meng2013} proteins, or as a tool in efforts to combine population genetics and molecular simulations.~\cite{Serohijos2014} While applied to proteins in this work, we note that the MS algorithm is equally valid for other bio-macromolecules, including DNA and RNA.} The ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
 
@@
 
-This work was supported by grants from Memorial University and National Sciences and Engineering Research Council of Canada (NSERC).
+This work was supported by grants from Memorial University and National Sciences and Engineering Research Council of Canada (RGPIN-2016-05014).
 

### 2017-10-05 JCP_revision/Manuscript.tex
Final submission and revision

--- 2017-07-19 Manuscript.tex
@@
 amsmath,amssymb,
-%preprint,%
-reprint,%
+preprint,%
+%reprint,%
 %author-year,%
@@
 
-\title[Multisequence Monte Carlo simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: exploring the sequence-structure relationship of proteins}
+\title[Multisequence biomolecular simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: exploring the sequence-structure relationship of proteins}
 
@@
 
-{\color{red} As a test application of the MS algorithm, we selected the phenomenon of protein fold switching which recently was demonstrated in a handful of natural and engineered proteins.
+As a test application of the MS algorithm, we selected the phenomenon of protein fold switching which recently was demonstrated in a handful of natural and engineered proteins.
 These special proteins exhibit a unique ability to reversibly switch between entirely different folds, with accompanying changes in secondary structure, hydrophobic core packing and overall shape.~\cite{Bryan2010} The fold switching transitions found in natural proteins typically play a functional role.
-For example, rare transitions to an alternative fold in the protein KaiB provide a crucial time delay mechanism in the circadian clock cycle of cyanobacteria.~\cite{LiWang2015} Fold switching can occur either spontaneously~\cite{Meier2007} or be triggered by various signals including changes to solution conditions,~\cite{Kuloglu2002} subdomain detachment,~\cite{Burmann2012} ligand binding~\cite{Luo2004} and point mutations.~\cite{Alexander2009} Computational studies, using  coarse-grained~\cite{Holzgrafe2014,Holzgrafe2015,Kouza2012,Sutto2012} or atomistic~\cite{Ramirez-Sarmiento2015,Sikosek2016,Hansen2013} models, have attempted to explain how proteins can exhibit multiple folding funnels and how they are altered in response to binding events or changes in sequence.}
+For example, rare transitions to an alternative fold in the protein KaiB provide a crucial time delay mechanism in the circadian clock cycle of cyanobacteria.~\cite{LiWang2015} Fold switching can occur either spontaneously~\cite{Meier2007} or be triggered by various signals including changes to solution conditions,~\cite{Kuloglu2002} subdomain detachment,~\cite{Burmann2012} ligand binding~\cite{Luo2004} and point mutations.~\cite{Alexander2009} Computational studies, using  coarse-grained~\cite{Holzgrafe2014,Holzgrafe2015,Kouza2012,Sutto2012} or atomistic~\cite{Ramirez-Sarmiento2015,Sikosek2016,Hansen2013} models, have attempted to explain how proteins can exhibit multiple folding funnels and how they are altered in response to binding events or changes in sequence.
 
-{\color{red} The discovery that proteins can be driven to switch folds through an accumulation of point mutations, in particular, holds implications for protein evolution as it suggests a simple mechanism of fold evolution.~\cite{Sikosek2014} Alexander \textit{et al.} demonstrated that the similarly sized but structurally distinct A ($3\alpha$) and B ($4\beta+\alpha$) domains of protein G could, after extensive mutations leaving their respective folds undisturbed, be triggered to switch folds by applying one additional mutation, Y45L, located at the edge of the hydrophobic core in the B domain.~\cite{Alexander2009} This remarkable discovery suggests the possibility that the two domains might be evolutionary related despite a lack of detectable similarity in either sequence or structure in wild-type protein G,~\cite{He2012} although this has yet to be proven.
-Moreover, it is unclear how common such fold-to-fold transitions are and how they might occur in evolutionary processes.~\cite{Porter2015} In previous work~\cite{Holzgrafe2014,Holzgrafe2015} we showed that mutational paths with abrupt fold switching exist between two other pairs of smaller protein folds within the model framework provided by our CG model.~\cite{Bhattacherjee2012} }
+The discovery that proteins can be driven to switch folds through an accumulation of point mutations, in particular, holds implications for protein evolution as it suggests a simple mechanism of fold evolution.~\cite{Sikosek2014} Alexander \textit{et al.} demonstrated that the similarly sized but structurally distinct A ($3\alpha$) and B ($4\beta+\alpha$) domains of protein G could, after extensive mutations leaving their respective folds undisturbed, be triggered to switch folds by applying one additional mutation, Y45L, located at the edge of the hydrophobic core in the B domain.~\cite{Alexander2009} This remarkable discovery suggests the possibility that the two domains might be evolutionary related despite a lack of detectable similarity in either sequence or structure in wild-type protein G,~\cite{He2012} although this has yet to be proven.
+Moreover, it is unclear how common such fold-to-fold transitions are and how they might occur in evolutionary processes.~\cite{Porter2015} In previous work~\cite{Holzgrafe2014,Holzgrafe2015} we showed that mutational paths with abrupt fold switching exist between two other pairs of smaller protein folds within the framework of our CG model.~\cite{Bhattacherjee2012}
 
@@
 
-Conventional {\color{red} Metropolis} Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
-The basic idea of generalized-ensemble algorithms is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters~\cite{Mitsutake2001} {\color{red} (for a recent historical account see Ref.~\citenum{Berg2017}).
-Generalized-ensemble methods that have been frequently used for biomolecular simulations include simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} replica exchange~\cite{Swendsen1986}  or parallel tempering,~\cite{Hansmann1997} and metadynamics.~\cite{Laio2002}}
+Conventional Metropolis Monte Carlo simulations of the canonical distribution is problematic at low temperatures for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
+The basic idea of generalized-ensemble algorithms is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters~\cite{Mitsutake2001} (for a recent historical account see Ref.~\citenum{Berg2017}).
+Generalized-ensemble methods that have been frequently used for biomolecular simulations include simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} replica exchange~\cite{Swendsen1986}  or parallel tempering,~\cite{Hansmann1997} and metadynamics.~\cite{Laio2002}
 
@@
 \begin{equation}
-P(\rbar,m) =\dfrac{1}{\hat{Z}} e^{-\beta_m E(\rbar) + g_m}\,,
+P(m,\rbar) =\dfrac{1}{\hat{Z}} e^{-\beta_m E(\rbar) + g_m}\,,
 \label{ST}
@@
 \end{equation}
-where  {\color{red}
+where  
 \begin{equation}
@@
 \end{equation}
-and S is a set of pre-selected sequences, i.e., the sequences to which visits are allowed during a simulation.} The simulation parameters $h(\sbar)$, similar to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=Z^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
+and S is a set of pre-selected sequences, i.e., the sequences to which visits are allowed during a simulation.
+The simulation parameters $h(\sbar)$, similar to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=Z^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
 
-Two types of MC updates are required to sample from the distrubution in Eq.~\ref{MS}, ordinary conformational updates $\rbar\rightarrow\rbar'$ and {\color{red} sequence} updates $\sbar\rightarrow\sbar'$.
-The acceptance probabilities for the latter  becomes
+Two types of MC updates are required to sample from the distrubution in Eq.~\ref{MS}, ordinary conformational updates $\rbar\rightarrow\rbar'$ and sequence updates $\sbar\rightarrow\sbar'$.
+The acceptance probability for the latter  becomes
 \begin{equation}
@@
 where $\Delta E = E(\sbar',\rbar)-E(\sbar,\rbar)$ and $\Delta h = h({\sbar'})-h(\sbar)$.
+
+{\color{red}Picking a new sequence $\sbar'$ in a sequence update $\sbar\rightarrow\sbar'$ can be done in several ways.
+One possibility is to draw $\sbar'$ randomly from the set S, such than $\sbar'\ne\sbar$.
+Alternatively, a type of ``mutational" move can be used where an amino acid position is first picked and then assigned a new amino acid type.
+The selection of position and type would have to be chosen such that $\sbar'$ does not end up outside S.
+In this work, we use the former update which is general and guarantees that ergodicity is fulfilled for any S.
+Importantly, both updates fulfill detailed balance and therefore lead to the same estimates of equilibrium quantities, such as native state stabilities, for the different sequences in S.}
 
@@
 
-{\color{red}In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out in the following way.
-First, a new random sequence $\sbar$ is picked from the set of pre-selected (allowed) sequences S, such that $\sbar'\ne\sbar$.
-The new sequence $\sbar'$ therefore differs from $\sbar$ in one or more amino acid positions.
+In our MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out in the following way.
+First, a new sequence $\sbar'$ is picked randomly from the set of pre-selected (allowed) sequences S, such that $\sbar'\ne\sbar$.
+This new sequence $\sbar'$ therefore differs from $\sbar$ in one or more amino acid positions.
 Thereafter, the sidechains of the protein, which remains in an unchanged (backbone) conformation $\rbar$, is re-built according to the new sequence $\sbar'$.
@@
 Finally, the change in total energy $\Delta E$ is calculated and the accept-reject criterion in Eq.~\ref{accrej} is applied.
-If rejected, the old state $(\sbar,\rbar)$ is restored.}
+If rejected, the old state $(\sbar,\rbar)$ is restored.
 
 A sequence update is attempted every $1,000$ MC steps while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
-{\color{red}The computational cost for sequence updates is somewhat higher than for temperature updates.
-The latter does not require any energy calculation and is thus extremely rapid.
-For the purpose of comparing computational efficiencies of ST and MS, we therefore chose sequence updates to be slightly less frequent than temperature updates while both are fairly frequent.} All simulations carried out in this work are summarized in Table~II.
+The computational cost for sequence updates is somewhat higher than for temperature updates.
+The latter update does not require any energy calculation and is thus extremely rapid.
+For the purpose of comparing computational efficiencies of ST and MS, we therefore chose sequence updates to be slightly less frequent than temperature updates while both are fairly frequent.
+All simulations carried out in this work are summarized in Table~II.
 
@@
 \begin{tabular}{lcccr}
-Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes a thermalization step with {\color{red} $10^7$} MC steps/run.} &  Sequences\\
+Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes a thermalization step with $10^7$ MC steps/run.} &  Sequences\\
 \hline
@@
 At the lowest $T$ and highest $\Delta h$, acceptance rates are only around 0.1-0.2.
-However, for most other $T$ and $\Delta h$ the overall acceptance rates are substantially higher and often above the oft-quoted rule-of-thumb value 0.25~\cite{Gilks1996} (see Fig~4B).
+However, for most other $T$ and $\Delta h$ the overall acceptance rates are substantially higher and often above the oft-quoted rule-of-thumb value 0.25~\cite{Gilks1996} (see Fig~3B).
 An increased acceptance rate can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer chains.
-{\color{red} For example, $\Delta h_\mathrm{max} = 1$ would be equivalent to applying only a ``mutational" update, i.e., picking a random (allowed) position and changing the amino acid type at that position.}
+For example, $\Delta h_\mathrm{max} = 1$ would be equivalent to applying only a ``mutational" update, i.e., picking a random (allowed) position and changing the amino acid type at that position.
 
@@
 We first validate the MS algorithm by comparing the average total energy, $\left < E \right >$, calculated for these 4 sequences with the two different methods (see Supplementary Information).
-The two sets of results are entirely consistent showing that, for a given $\sbar$ and $T$, the MS and ST algorithms indeed sample the same distribution.
+The two sets of results are entirely consistent showing that, for a given $\sbar$ and $T$, the MS and ST algorithms indeed sample the same (Boltzmann) distribution.
 
@@
 The results indicate that there are two main benefits of this approach.
-Firstly, it provides a convenient way to sample the canonical distributions of large numbers of sequences in a single run and, secondly, it enhances the sampling of conformational space meaning it can be applied directly at low temperatures.
+Firstly, it provides a convenient way to sample the canonical distributions of large numbers of sequences in a single run and, secondly, it enhances the sampling of conformational space meaning it can be applied directly to low temperatures.
 The conformational sampling efficiency can be assessed from the comparison with ST.
@@
 It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant $T$.
-{\color{red} Therefore, rather than promoting escape from local minima by visits to higher $T$, as in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986} MS simulations must escape the minima occurring for one sequence through visits to other sequences.
+Therefore, rather than promoting escape from local minima by visits to higher $T$, as in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986} MS simulations must escape the minima occurring for one sequence through visits to other sequences.
 How this process works can be envisioned by considering an MS simulation that is visiting a sequence $s$ and is trapped in a local minimum, requiring that a high free-energy barrier is overcome for escape.
 The trapped state could be, e.g., a compact $\beta$-sheet rich state with a particular register.
-Eventually, sequence updates will carry the simulation to other sequences $\hat{s}$ while still remaining in the trapped state.
-However, the free energy barrier of this trapped state might be substantially lower for $\hat{s}$ than for $s$, or the barrier may even be altogether absent if the trapped state is unstable for $\hat{s}$, leading to rapid escape from the minimum.
-} The above reasoning also implies that the performance of the MS algorithm likely depends on the size the sequence set S as well as the conformational properties of the sequences.
+Eventually, sequence updates will carry the simulation to other sequences $\hat{s}$ while it conformationally still remains in the trapped state.
+However, the free energy barrier of escape might be substantially lower for $\hat{s}$ than for $s$, or the barrier may even be altogether absent if the trapped state is unstable for $\hat{s}$, leading to rapid escape from the minimum.
+The above reasoning also implies that the performance of the MS algorithm likely depends on the size of the sequence set S as well as the conformational properties of the sequences.
 Specifically, the performance of MS simulations of proteins at low $T$ may benefit from the inclusion of at least a few sequences with poor stability properties, such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
 
-{\color{red} Individual protein sequences that exhibit spontaneous transitions between widely different competing conformational states, such as fold switching proteins, are especially challenging to molecular simulation methods.
-This problem has been addressed by using Hamiltonian replica-exchange techniques to couple G\={o}-like terms, i.e., energy terms with an artificial bias towards a given structure, with a physical force field.~\cite{Meinke2007,Chen2014,Bernhardt2016} This way exchange moves ``feed" conformations into the replica corresponding to the physical force field and enhance sampling.~\cite{Bernhardt2016} We did not specifically study sampling efficiency for sequences exhibiting competing states.
+Individual protein sequences that exhibit spontaneous transitions between widely different competing conformational states, such as fold switching proteins, are especially challenging to molecular simulation methods.
+Representative sampling in such cases requires multiple transitions between highly different states, which can be a slow process.
+This problem has been addressed by using Hamiltonian replica-exchange techniques to couple G\={o}-like terms, i.e., energy terms with an artificial bias towards a given structure, with a physical force field.~\cite{Meinke2007,Chen2014,Bernhardt2016} This way exchange moves ``feed" diverse conformations into the replica corresponding to the physical force field and enhance sampling.~\cite{Bernhardt2016} We did not specifically study sampling efficiency for sequences exhibiting competing states, such as bistable sequences.
 It appears likely, however, that a very similar type of ``feeding" of conformations would take place in MS simulations although the coupling occurs instead with other sequences rather than with G\={o}-type terms.
-}
 
@@
 However, our results indicate that for CG models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), our method can be a highly efficient way to sample the equilibrium behavior of many sequences.
-{\color{red} This opens up for various applications, such as exploring sequence effects on the conformational properties of disordered~\cite{Das2013} or denatured~\cite{Meng2013} proteins, or as a tool in efforts to combine population genetics and molecular simulations.~\cite{Serohijos2014} While applied to proteins in this work, we note that the MS algorithm is equally valid for other bio-macromolecules, including DNA and RNA.} The ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
+This opens up for various applications, such as exploring sequence effects on the conformational properties of disordered~\cite{Das2013} or denatured~\cite{Meng2013} proteins, or as a tool in efforts to combine population genetics and molecular simulations.~\cite{Serohijos2014} While applied to proteins in this work, we note that the theoretical framework of the MS algorithm is equally valid for other bio-macromolecules, including DNA and RNA.
+The ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unraveling of the basic underlying structure.
 
@@
 The method performs a random walk in sequence space which is especially useful at low temperature as it promotes escape from local minima present in the free energy landscapes of individual sequences.
-The method might be suitable for CG simulations of {\color{red} various} other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
+The method might be suitable for CG simulations of various other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
 
-\section{Supplementary Materia}
+\section{Supplementary Material}
 
-Validation of the MS algorithm.
+See supplementary material for the validation of the MS algorithm.
 
@@
 
-\bibliography{/Users/stefan/Documents/Manuscripts/Jshort,/Users/stefan/Documents/Manuscripts/MyRefs}
+%\bibliography{/Users/stefan/Documents/Manuscripts/Jshort,/Users/stefan/Documents/Manuscripts/MyRefs}
 %\input{Manuscript.bbl}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%merlin.mbs aipnum4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
+%Control: key (0)
+%Control: author (8) initials jnrlst
+%Control: editor formatted (1) identically to author
+%Control: production of article title (-1) disabled
+%Control: page (0) single
+%Control: year (1) truncated
+%Control: production of eprint (0) enabled
+\begin{thebibliography}{63}%
+\makeatletter
+\providecommand \@ifxundefined [1]{%
+ \@ifx{#1\undefined}
+}%
+\providecommand \@ifnum [1]{%
+ \ifnum #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \@ifx [1]{%
+ \ifx #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \natexlab [1]{#1}%
+\providecommand \enquote  [1]{``#1''}%
+\providecommand \bibnamefont  [1]{#1}%
+\providecommand \bibfnamefont [1]{#1}%
+\providecommand \citenamefont [1]{#1}%
+\providecommand \href@noop [0]{\@secondoftwo}%
+\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
+\providecommand \@href[1]{\@@startlink{#1}\@@href}%
+\providecommand \@@href[1]{\endgroup#1\@@endlink}%
+\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
+  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
+\providecommand \@@startlink[1]{}%
+\providecommand \@@endlink[0]{}%
+\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
+\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
+\providecommand \urlprefix  [0]{URL }%
+\providecommand \Eprint [0]{\href }%
+\providecommand \doibase [0]{http://dx.doi.org/}%
+\providecommand \selectlanguage [0]{\@gobble}%
+\providecommand \bibinfo  [0]{\@secondoftwo}%
+\providecommand \bibfield  [0]{\@secondoftwo}%
+\providecommand \translation [1]{[#1]}%
+\providecommand \BibitemOpen [0]{}%
+\providecommand \bibitemStop [0]{}%
+\providecommand \bibitemNoStop [0]{.\EOS\space}%
+\providecommand \EOS [0]{\spacefactor3000\relax}%
+\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
+\let\auto@bib@innerbib\@empty
+%</preamble>
+\bibitem [{\citenamefont {Piana}, \citenamefont {Klepeis},\ and\ \citenamefont
+  {Shaw}(2014)}]{Piana2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Piana}}, \bibinfo {author} {\bibfnamefont {J.~L.}\ \bibnamefont {Klepeis}},
+  \ and\ \bibinfo {author} {\bibfnamefont {D.~E.}\ \bibnamefont {Shaw}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Curr Opin Struct
+  Biol}\ }\textbf {\bibinfo {volume} {24}},\ \bibinfo {pages} {98} (\bibinfo
+  {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Ding}\ \emph {et~al.}(2008)\citenamefont {Ding},
+  \citenamefont {Tsao}, \citenamefont {Nie},\ and\ \citenamefont
+  {Dokholyan}}]{Ding2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
+  {Ding}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Tsao}}, \bibinfo
+  {author} {\bibfnamefont {H.}~\bibnamefont {Nie}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {N.~V.}\ \bibnamefont {Dokholyan}},\ }\href@noop {} {\bibfield
+   {journal} {\bibinfo  {journal} {Structure}\ }\textbf {\bibinfo {volume}
+  {16}},\ \bibinfo {pages} {1010} (\bibinfo {year} {2008})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Irb{\"a}ck}\ and\ \citenamefont
+  {Mohanty}(2006)}]{Irback2006}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Irb{\"a}ck}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Mohanty}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
+  Comput Chem}\ }\textbf {\bibinfo {volume} {27}},\ \bibinfo {pages} {1548}
+  (\bibinfo {year} {2006})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Verma}\ and\ \citenamefont
+  {Wenzel}(2009)}]{Verma2009}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Verma}}\ and\ \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont {Wenzel}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Biophys J}\
+  }\textbf {\bibinfo {volume} {96}},\ \bibinfo {pages} {3483} (\bibinfo {year}
+  {2009})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Yang}\ \emph {et~al.}(2007)\citenamefont {Yang},
+  \citenamefont {Chen}, \citenamefont {Skolnick},\ and\ \citenamefont
+  {Shakhnovich}}]{Yang2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont
+  {Yang}}, \bibinfo {author} {\bibfnamefont {W.~W.}\ \bibnamefont {Chen}},
+  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Skolnick}}, \ and\
+  \bibinfo {author} {\bibfnamefont {E.~I.}\ \bibnamefont {Shakhnovich}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Structure}\
+  }\textbf {\bibinfo {volume} {15}},\ \bibinfo {pages} {53} (\bibinfo {year}
+  {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bernardi}, \citenamefont {Melo},\ and\ \citenamefont
+  {Schulten}(2015)}]{Bernardi2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~C.}\ \bibnamefont
+  {Bernardi}}, \bibinfo {author} {\bibfnamefont {M.~C.}\ \bibnamefont {Melo}},
+  \ and\ \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Schulten}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Biochim Biophys
+  Acta}\ }\textbf {\bibinfo {volume} {1850}},\ \bibinfo {pages} {872} (\bibinfo
+  {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {McGuffee}\ and\ \citenamefont
+  {Elcock}(2010)}]{McGuffee2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.~R.}\ \bibnamefont
+  {McGuffee}}\ and\ \bibinfo {author} {\bibfnamefont {A.~H.}\ \bibnamefont
+  {Elcock}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {PLOS
+  Comput Biol}\ }\textbf {\bibinfo {volume} {6}},\ \bibinfo {pages} {e1000694}
+  (\bibinfo {year} {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Miao}, \citenamefont {Johnson},\ and\ \citenamefont
+  {Ortoleva}(2010)}]{Miao2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Miao}}, \bibinfo {author} {\bibfnamefont {J.~E.}\ \bibnamefont {Johnson}}, \
+  and\ \bibinfo {author} {\bibfnamefont {P.~J.}\ \bibnamefont {Ortoleva}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J Phys Chem B}\
+  }\textbf {\bibinfo {volume} {114}},\ \bibinfo {pages} {11181} (\bibinfo
+  {year} {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Perilla}\ \emph {et~al.}(2016)\citenamefont
+  {Perilla}, \citenamefont {Hadden}, \citenamefont {Goh}, \citenamefont
+  {Mayne},\ and\ \citenamefont {Schulten}}]{Perilla2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~R.}\ \bibnamefont
+  {Perilla}}, \bibinfo {author} {\bibfnamefont {J.~A.}\ \bibnamefont {Hadden}},
+  \bibinfo {author} {\bibfnamefont {B.~C.}\ \bibnamefont {Goh}}, \bibinfo
+  {author} {\bibfnamefont {C.~G.}\ \bibnamefont {Mayne}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {K.}~\bibnamefont {Schulten}},\ }\href@noop {}
+  {\bibfield  {journal} {\bibinfo  {journal} {J Phys Chem Lett}\ }\textbf
+  {\bibinfo {volume} {7}},\ \bibinfo {pages} {1836} (\bibinfo {year}
+  {2016})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Lindorff-Larsen}\ \emph {et~al.}(2011)\citenamefont
+  {Lindorff-Larsen}, \citenamefont {Piana}, \citenamefont {Dror},\ and\
+  \citenamefont {Shaw}}]{Lindorff-Larsen2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
+  {Lindorff-Larsen}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Piana}}, \bibinfo {author} {\bibfnamefont {R.~O.}\ \bibnamefont {Dror}}, \
+  and\ \bibinfo {author} {\bibfnamefont {D.~E.}\ \bibnamefont {Shaw}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Science}\ }\textbf
+  {\bibinfo {volume} {334}},\ \bibinfo {pages} {517} (\bibinfo {year}
+  {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Yu}\ \emph {et~al.}(2016)\citenamefont {Yu},
+  \citenamefont {Mori}, \citenamefont {Ando}, \citenamefont {Harada},
+  \citenamefont {Jung}, \citenamefont {Sugita},\ and\ \citenamefont
+  {Feig}}]{Yu2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {I.}~\bibnamefont
+  {Yu}}, \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Mori}}, \bibinfo
+  {author} {\bibfnamefont {T.}~\bibnamefont {Ando}}, \bibinfo {author}
+  {\bibfnamefont {R.}~\bibnamefont {Harada}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Jung}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Sugita}}, \ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Feig}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {ELife}\
+  }\textbf {\bibinfo {volume} {5}} (\bibinfo {year} {2016})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Riniker}, \citenamefont {Allison},\ and\
+  \citenamefont {van Gunsteren}(2012)}]{Riniker2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Riniker}}, \bibinfo {author} {\bibfnamefont {J.~R.}\ \bibnamefont
+  {Allison}}, \ and\ \bibinfo {author} {\bibfnamefont {W.~F.}\ \bibnamefont
+  {van Gunsteren}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys Chem Chem Phys}\ }\textbf {\bibinfo {volume} {14}},\ \bibinfo {pages}
+  {12423} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Ing{\'{o}}lfsson}\ \emph {et~al.}(2014)\citenamefont
+  {Ing{\'{o}}lfsson}, \citenamefont {Lopez}, \citenamefont {Uusitalo},
+  \citenamefont {{de Jong}}, \citenamefont {Gopal}, \citenamefont {Periole},\
+  and\ \citenamefont {Marrink}}]{Ingolfsson2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.~I.}\ \bibnamefont
+  {Ing{\'{o}}lfsson}}, \bibinfo {author} {\bibfnamefont {C.~A.}\ \bibnamefont
+  {Lopez}}, \bibinfo {author} {\bibfnamefont {J.~J.}\ \bibnamefont {Uusitalo}},
+  \bibinfo {author} {\bibfnamefont {D.~H.}\ \bibnamefont {{de Jong}}}, \bibinfo
+  {author} {\bibfnamefont {S.~M.}\ \bibnamefont {Gopal}}, \bibinfo {author}
+  {\bibfnamefont {X.}~\bibnamefont {Periole}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {S.~J.}\ \bibnamefont {Marrink}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {WIREs Comput Mol Sci}\ }\textbf {\bibinfo
+  {volume} {4}},\ \bibinfo {pages} {225} (\bibinfo {year} {2014})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Beland}\ \emph {et~al.}(2011)\citenamefont {Beland},
+  \citenamefont {Brommer}, \citenamefont {El-Mellouhi}, \citenamefont {Joly},\
+  and\ \citenamefont {Mousseau}}]{Beland2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.~K.}\ \bibnamefont
+  {Beland}}, \bibinfo {author} {\bibfnamefont {P.}~\bibnamefont {Brommer}},
+  \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {El-Mellouhi}}, \bibinfo
+  {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Joly}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {N.}~\bibnamefont {Mousseau}},\ }\href@noop {}
+  {\bibfield  {journal} {\bibinfo  {journal} {Phys Rev E}\ }\textbf {\bibinfo
+  {volume} {84}},\ \bibinfo {pages} {046704} (\bibinfo {year}
+  {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Proctor}, \citenamefont {Ding},\ and\ \citenamefont
+  {Dokholyan}(2011)}]{Proctor2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.~A.}\ \bibnamefont
+  {Proctor}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Ding}}, \
+  and\ \bibinfo {author} {\bibfnamefont {N.~V.}\ \bibnamefont {Dokholyan}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {WIREs Comput Mol
+  Sci}\ }\textbf {\bibinfo {volume} {1}},\ \bibinfo {pages} {80} (\bibinfo
+  {year} {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Vitalis}\ and\ \citenamefont
+  {Pappu}(2009)}]{Vitalis2009}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Vitalis}}\ and\ \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont
+  {Pappu}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Annu
+  Rep Comput Chem}\ }\textbf {\bibinfo {volume} {5}},\ \bibinfo {pages} {49}
+  (\bibinfo {year} {2009})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zarrinpar}, \citenamefont {Park},\ and\ \citenamefont
+  {Lim}(2003)}]{Zarrinpar2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Zarrinpar}}, \bibinfo {author} {\bibfnamefont {S.~H.}\ \bibnamefont {Park}},
+  \ and\ \bibinfo {author} {\bibfnamefont {W.~A.}\ \bibnamefont {Lim}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Nature}\ }\textbf
+  {\bibinfo {volume} {426}},\ \bibinfo {pages} {676} (\bibinfo {year}
+  {2003})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Hakes}\ \emph {et~al.}(2007)\citenamefont {Hakes},
+  \citenamefont {Lovell}, \citenamefont {Oliver},\ and\ \citenamefont
+  {Robertson}}]{Hakes2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Hakes}}, \bibinfo {author} {\bibfnamefont {S.~C.}\ \bibnamefont {Lovell}},
+  \bibinfo {author} {\bibfnamefont {S.~G.}\ \bibnamefont {Oliver}}, \ and\
+  \bibinfo {author} {\bibfnamefont {D.~L.}\ \bibnamefont {Robertson}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Proc Natl Acad Sci
+  USA}\ }\textbf {\bibinfo {volume} {104}},\ \bibinfo {pages} {7999} (\bibinfo
+  {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Rohs}\ \emph {et~al.}(2010)\citenamefont {Rohs},
+  \citenamefont {Jin}, \citenamefont {West}, \citenamefont {Joshi},
+  \citenamefont {Honig},\ and\ \citenamefont {Mann}}]{Rohs2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {Rohs}}, \bibinfo {author} {\bibfnamefont {X.}~\bibnamefont {Jin}}, \bibinfo
+  {author} {\bibfnamefont {S.~M.}\ \bibnamefont {West}}, \bibinfo {author}
+  {\bibfnamefont {R.}~\bibnamefont {Joshi}}, \bibinfo {author} {\bibfnamefont
+  {B.}~\bibnamefont {Honig}}, \ and\ \bibinfo {author} {\bibfnamefont {R.~S.}\
+  \bibnamefont {Mann}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
+  {journal} {Annu Rev Biochem}\ }\textbf {\bibinfo {volume} {79}},\ \bibinfo
+  {pages} {233} (\bibinfo {year} {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Ross}\ and\ \citenamefont
+  {Poirier}(2004)}]{Ross2004}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.~A.}\ \bibnamefont
+  {Ross}}\ and\ \bibinfo {author} {\bibfnamefont {M.~A.}\ \bibnamefont
+  {Poirier}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Nat
+  Med}\ }\textbf {\bibinfo {volume} {10 Suppl}},\ \bibinfo {pages} {S10}
+  (\bibinfo {year} {2004})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Tzul}, \citenamefont {Vasilchuk},\ and\ \citenamefont
+  {Makhatadze}(2017)}]{Tzul2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.~O.}\ \bibnamefont
+  {Tzul}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Vasilchuk}}, \
+  and\ \bibinfo {author} {\bibfnamefont {G.~I.}\ \bibnamefont {Makhatadze}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Proc Natl Acad Sci
+  USA}\ }\textbf {\bibinfo {volume} {114}},\ \bibinfo {pages} {E1627} (\bibinfo
+  {year} {2017})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Wensley}\ \emph {et~al.}(2010)\citenamefont
+  {Wensley}, \citenamefont {Batey}, \citenamefont {Bone}, \citenamefont {Chan},
+  \citenamefont {Tumelty}, \citenamefont {Steward}, \citenamefont {Kwa},
+  \citenamefont {Borgia},\ and\ \citenamefont {Clarke}}]{Wensley2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.~G.}\ \bibnamefont
+  {Wensley}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Batey}},
+  \bibinfo {author} {\bibfnamefont {F.~A.}\ \bibnamefont {Bone}}, \bibinfo
+  {author} {\bibfnamefont {Z.~M.}\ \bibnamefont {Chan}}, \bibinfo {author}
+  {\bibfnamefont {N.~R.}\ \bibnamefont {Tumelty}}, \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Steward}}, \bibinfo {author} {\bibfnamefont
+  {L.~G.}\ \bibnamefont {Kwa}}, \bibinfo {author} {\bibfnamefont
+  {A.}~\bibnamefont {Borgia}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Clarke}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
+  {journal} {Nature}\ }\textbf {\bibinfo {volume} {463}},\ \bibinfo {pages}
+  {685} (\bibinfo {year} {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Vukmirovic}\ and\ \citenamefont
+  {Tilghman}(2000)}]{Vukmirovic2000}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~G.}\ \bibnamefont
+  {Vukmirovic}}\ and\ \bibinfo {author} {\bibfnamefont {S.~M.}\ \bibnamefont
+  {Tilghman}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Nature}\ }\textbf {\bibinfo {volume} {405}},\ \bibinfo {pages} {820}
+  (\bibinfo {year} {2000})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bhattacherjee}\ and\ \citenamefont
+  {Wallin}(2012)}]{Bhattacherjee2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Bhattacherjee}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Biophys J}\ }\textbf {\bibinfo {volume} {102}},\ \bibinfo {pages} {569}
+  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Irb{\"a}ck}\ and\ \citenamefont
+  {Potthast}(1995)}]{Irback1995}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Irb{\"a}ck}}\ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
+  {Potthast}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Comp. Phys.}\ }\textbf {\bibinfo {volume} {103}},\ \bibinfo {pages} {10298}
+  (\bibinfo {year} {1995})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bhattacherjee}\ and\ \citenamefont
+  {Wallin}(2013)}]{Bhattacherjee2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Bhattacherjee}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {PLOS
+  Comput Biol}\ }\textbf {\bibinfo {volume} {9}},\ \bibinfo {pages} {e1003277}
+  (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Wallin}(2017)}]{Wallin2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Methods Mol Biol}\ }\textbf {\bibinfo {volume} {1561}},\ \bibinfo {pages}
+  {201} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bryan}\ and\ \citenamefont
+  {Orban}(2010)}]{Bryan2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.~N.}\ \bibnamefont
+  {Bryan}}\ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Orban}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Curr Opin Struct
+  Biol}\ }\textbf {\bibinfo {volume} {20}},\ \bibinfo {pages} {482} (\bibinfo
+  {year} {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Chang}\ \emph {et~al.}(2015)\citenamefont {Chang},
+  \citenamefont {Cohen}, \citenamefont {Phong}, \citenamefont {Myers},
+  \citenamefont {Kim}, \citenamefont {Tseng}, \citenamefont {Lin},
+  \citenamefont {Zhang}, \citenamefont {Boyd}, \citenamefont {Lee},
+  \citenamefont {Kang}, \citenamefont {Lee}, \citenamefont {Li}, \citenamefont
+  {Britt}, \citenamefont {Rust}, \citenamefont {Golden},\ and\ \citenamefont
+  {LiWang}}]{LiWang2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.~G.}\ \bibnamefont
+  {Chang}}, \bibinfo {author} {\bibfnamefont {S.~E.}\ \bibnamefont {Cohen}},
+  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Phong}}, \bibinfo
+  {author} {\bibfnamefont {W.~K.}\ \bibnamefont {Myers}}, \bibinfo {author}
+  {\bibfnamefont {Y.~I.}\ \bibnamefont {Kim}}, \bibinfo {author} {\bibfnamefont
+  {R.}~\bibnamefont {Tseng}}, \bibinfo {author} {\bibfnamefont
+  {J.}~\bibnamefont {Lin}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Zhang}}, \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont {Boyd}},
+  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Lee}}, \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Kang}}, \bibinfo {author} {\bibfnamefont
+  {D.}~\bibnamefont {Lee}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Li}}, \bibinfo {author} {\bibfnamefont {R.~D.}\ \bibnamefont {Britt}},
+  \bibinfo {author} {\bibfnamefont {M.~J.}\ \bibnamefont {Rust}}, \bibinfo
+  {author} {\bibfnamefont {S.~S.}\ \bibnamefont {Golden}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {A.}~\bibnamefont {LiWang}},\ }\href@noop {}
+  {\bibfield  {journal} {\bibinfo  {journal} {Science}\ }\textbf {\bibinfo
+  {volume} {349}},\ \bibinfo {pages} {324} (\bibinfo {year}
+  {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Meier}\ \emph {et~al.}(2007)\citenamefont {Meier},
+  \citenamefont {Jensen}, \citenamefont {David}, \citenamefont {Chapman},
+  \citenamefont {Holstein}, \citenamefont {Grzesiek},\ and\ \citenamefont
+  {Ozbek}}]{Meier2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Meier}}, \bibinfo {author} {\bibfnamefont {P.~R.}\ \bibnamefont {Jensen}},
+  \bibinfo {author} {\bibfnamefont {C.~N.}\ \bibnamefont {David}}, \bibinfo
+  {author} {\bibfnamefont {J.}~\bibnamefont {Chapman}}, \bibinfo {author}
+  {\bibfnamefont {T.~W.}\ \bibnamefont {Holstein}}, \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Grzesiek}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {S.}~\bibnamefont {Ozbek}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Curr Biol}\ }\textbf {\bibinfo {volume}
+  {17}},\ \bibinfo {pages} {173} (\bibinfo {year} {2007})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Kuloglu}\ \emph {et~al.}(2002)\citenamefont
+  {Kuloglu}, \citenamefont {McCaslin}, \citenamefont {Markley},\ and\
+  \citenamefont {Volkman}}]{Kuloglu2002}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.~S.}\ \bibnamefont
+  {Kuloglu}}, \bibinfo {author} {\bibfnamefont {D.~R.}\ \bibnamefont
+  {McCaslin}}, \bibinfo {author} {\bibfnamefont {J.~L.}\ \bibnamefont
+  {Markley}}, \ and\ \bibinfo {author} {\bibfnamefont {B.~F.}\ \bibnamefont
+  {Volkman}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
+  Biol Chem}\ }\textbf {\bibinfo {volume} {277}},\ \bibinfo {pages} {17863}
+  (\bibinfo {year} {2002})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Burmann}\ \emph {et~al.}(2012)\citenamefont
+  {Burmann}, \citenamefont {Knauer}, \citenamefont {Sevostyanova},
+  \citenamefont {Schweimer}, \citenamefont {Mooney}, \citenamefont {Landick},
+  \citenamefont {Artsimovitch},\ and\ \citenamefont {R{\"o}sch}}]{Burmann2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.~M.}\ \bibnamefont
+  {Burmann}}, \bibinfo {author} {\bibfnamefont {S.~H.}\ \bibnamefont {Knauer}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Sevostyanova}}, \bibinfo
+  {author} {\bibfnamefont {K.}~\bibnamefont {Schweimer}}, \bibinfo {author}
+  {\bibfnamefont {R.~A.}\ \bibnamefont {Mooney}}, \bibinfo {author}
+  {\bibfnamefont {R.}~\bibnamefont {Landick}}, \bibinfo {author} {\bibfnamefont
+  {I.}~\bibnamefont {Artsimovitch}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {P.}~\bibnamefont {R{\"o}sch}},\ }\href@noop {} {\bibfield  {journal}
+  {\bibinfo  {journal} {Cell}\ }\textbf {\bibinfo {volume} {150}},\ \bibinfo
+  {pages} {291} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Luo}\ \emph {et~al.}(2004)\citenamefont {Luo},
+  \citenamefont {Tang}, \citenamefont {Xia}, \citenamefont {Wassmann},
+  \citenamefont {Matsumoto}, \citenamefont {Rizo},\ and\ \citenamefont
+  {Yu}}]{Luo2004}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
+  {Luo}}, \bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont {Tang}}, \bibinfo
+  {author} {\bibfnamefont {G.}~\bibnamefont {Xia}}, \bibinfo {author}
+  {\bibfnamefont {K.}~\bibnamefont {Wassmann}}, \bibinfo {author}
+  {\bibfnamefont {T.}~\bibnamefont {Matsumoto}}, \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Rizo}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {H.}~\bibnamefont {Yu}},\ }\href@noop {} {\bibfield  {journal}
+  {\bibinfo  {journal} {Nat Struct Mol Biol}\ }\textbf {\bibinfo {volume}
+  {11}},\ \bibinfo {pages} {338} (\bibinfo {year} {2004})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Alexander}\ \emph {et~al.}(2009)\citenamefont
+  {Alexander}, \citenamefont {He}, \citenamefont {Chen}, \citenamefont
+  {Orban},\ and\ \citenamefont {Bryan}}]{Alexander2009}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.~A.}\ \bibnamefont
+  {Alexander}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {He}},
+  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Chen}}, \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Orban}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {P.~N.}\ \bibnamefont {Bryan}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Proc Natl Acad Sci USA}\ }\textbf {\bibinfo
+  {volume} {106}},\ \bibinfo {pages} {21149} (\bibinfo {year}
+  {2009})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Holzgr{\"a}fe}\ and\ \citenamefont
+  {Wallin}(2014)}]{Holzgrafe2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Holzgr{\"a}fe}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Biophys J}\ }\textbf {\bibinfo {volume} {107}},\ \bibinfo {pages} {1217}
+  (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Holzgr{\"a}fe}\ and\ \citenamefont
+  {Wallin}(2015)}]{Holzgrafe2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Holzgr{\"a}fe}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys
+  Biol}\ }\textbf {\bibinfo {volume} {12}},\ \bibinfo {pages} {026002}
+  (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kouza}\ and\ \citenamefont
+  {Hansmann}(2012)}]{Kouza2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Kouza}}\ and\ \bibinfo {author} {\bibfnamefont {U.~H.}\ \bibnamefont
+  {Hansmann}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
+  Phys Chem B}\ }\textbf {\bibinfo {volume} {116}},\ \bibinfo {pages} {6645}
+  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Sutto}\ and\ \citenamefont
+  {Camilloni}(2012)}]{Sutto2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Sutto}}\ and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Camilloni}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Comp. Phys.}\ }\textbf {\bibinfo {volume} {136}},\ \bibinfo {pages} {185101}
+  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Ramirez-Sarmiento}\ \emph {et~al.}(2015)\citenamefont
+  {Ramirez-Sarmiento}, \citenamefont {Noel}, \citenamefont {Valenzuela},\ and\
+  \citenamefont {Artsimovitch}}]{Ramirez-Sarmiento2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.~A.}\ \bibnamefont
+  {Ramirez-Sarmiento}}, \bibinfo {author} {\bibfnamefont {J.~K.}\ \bibnamefont
+  {Noel}}, \bibinfo {author} {\bibfnamefont {S.~L.}\ \bibnamefont
+  {Valenzuela}}, \ and\ \bibinfo {author} {\bibfnamefont {I.}~\bibnamefont
+  {Artsimovitch}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {PLOS Comput Biol}\ }\textbf {\bibinfo {volume} {11}},\ \bibinfo {pages}
+  {e1004379} (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Sikosek}, \citenamefont {Krobath},\ and\
+  \citenamefont {Chan}(2016)}]{Sikosek2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Sikosek}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Krobath}}, \
+  and\ \bibinfo {author} {\bibfnamefont {H.~S.}\ \bibnamefont {Chan}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {PLOS Comput Biol}\
+  }\textbf {\bibinfo {volume} {12}},\ \bibinfo {pages} {e1004960} (\bibinfo
+  {year} {2016})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Hansen}\ \emph {et~al.}(2013)\citenamefont {Hansen},
+  \citenamefont {Allison}, \citenamefont {Hodel},\ and\ \citenamefont {van
+  Gunsteren}}]{Hansen2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
+  {Hansen}}, \bibinfo {author} {\bibfnamefont {J.~R.}\ \bibnamefont {Allison}},
+  \bibinfo {author} {\bibfnamefont {F.~H.}\ \bibnamefont {Hodel}}, \ and\
+  \bibinfo {author} {\bibfnamefont {W.~F.}\ \bibnamefont {van Gunsteren}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Biochemistry}\
+  }\textbf {\bibinfo {volume} {52}},\ \bibinfo {pages} {4962} (\bibinfo {year}
+  {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Sikosek}\ and\ \citenamefont
+  {Chan}(2014)}]{Sikosek2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Sikosek}}\ and\ \bibinfo {author} {\bibfnamefont {H.~S.}\ \bibnamefont
+  {Chan}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J R Soc
+  Interface}\ }\textbf {\bibinfo {volume} {11}},\ \bibinfo {pages} {20140419}
+  (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {He}\ \emph {et~al.}(2012)\citenamefont {He},
+  \citenamefont {Chen}, \citenamefont {Alexander}, \citenamefont {Bryan},\ and\
+  \citenamefont {Orban}}]{He2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {He}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Chen}}, \bibinfo
+  {author} {\bibfnamefont {P.~A.}\ \bibnamefont {Alexander}}, \bibinfo {author}
+  {\bibfnamefont {P.~N.}\ \bibnamefont {Bryan}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Orban}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Structure}\ }\textbf {\bibinfo {volume}
+  {20}},\ \bibinfo {pages} {283} (\bibinfo {year} {2012})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Porter}\ \emph {et~al.}(2015)\citenamefont {Porter},
+  \citenamefont {He}, \citenamefont {Chen}, \citenamefont {Orban},\ and\
+  \citenamefont {Bryan}}]{Porter2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.~L.}\ \bibnamefont
+  {Porter}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {He}}, \bibinfo
+  {author} {\bibfnamefont {Y.}~\bibnamefont {Chen}}, \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Orban}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {P.~N.}\ \bibnamefont {Bryan}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Biophys J}\ }\textbf {\bibinfo {volume}
+  {108}},\ \bibinfo {pages} {154} (\bibinfo {year} {2015})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Marinari}\ and\ \citenamefont
+  {Parisi}(1992)}]{Marinari1992}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
+  {Marinari}}\ and\ \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Parisi}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Europhys Lett}\ }\textbf {\bibinfo {volume} {19}} (\bibinfo {year}
+  {1992})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Lyubartsev}\ \emph {et~al.}(1992)\citenamefont
+  {Lyubartsev}, \citenamefont {Martsinovski}, \citenamefont {Shevkunov},\ and\
+  \citenamefont {Vorontsov-Velyaminov}}]{Lyubartsev1992}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~P.}\ \bibnamefont
+  {Lyubartsev}}, \bibinfo {author} {\bibfnamefont {A.~A.}\ \bibnamefont
+  {Martsinovski}}, \bibinfo {author} {\bibfnamefont {S.~V.}\ \bibnamefont
+  {Shevkunov}}, \ and\ \bibinfo {author} {\bibfnamefont {P.~N.}\ \bibnamefont
+  {Vorontsov-Velyaminov}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
+  {journal} {J. Comp. Phys.}\ }\textbf {\bibinfo {volume} {96}},\ \bibinfo
+  {pages} {1776} (\bibinfo {year} {1992})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Mitsutake}, \citenamefont {Sugita},\ and\
+  \citenamefont {Okamoto}(2001)}]{Mitsutake2001}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Mitsutake}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Sugita}}, \
+  and\ \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Okamoto}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Biopolymers}\
+  }\textbf {\bibinfo {volume} {60}},\ \bibinfo {pages} {96} (\bibinfo {year}
+  {2001})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Berg}(2017)}]{Berg2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.~A.}\ \bibnamefont
+  {Berg}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Eur Phys
+  J Spec Top}\ }\textbf {\bibinfo {volume} {226}},\ \bibinfo {pages} {,
+  551} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Swendsen}\ and\ \citenamefont
+  {Wang}(1986)}]{Swendsen1986}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~H.}\ \bibnamefont
+  {Swendsen}}\ and\ \bibinfo {author} {\bibfnamefont {J.-S.}\ \bibnamefont
+  {Wang}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. Lett.}\ }\textbf {\bibinfo {volume} {57}},\ \bibinfo {pages} {2607}
+  (\bibinfo {year} {1986})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Hansmann}(1997)}]{Hansmann1997}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {U.~H.~E.}\
+  \bibnamefont {Hansmann}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
+  {journal} {Chem Phys Lett}\ }\textbf {\bibinfo {volume} {281}},\ \bibinfo
+  {pages} {140} (\bibinfo {year} {1997})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Laio}\ and\ \citenamefont
+  {Parrinello}(2002)}]{Laio2002}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Laio}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Parrinello}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Proc Natl Acad Sci USA}\ }\textbf {\bibinfo {volume} {99}},\ \bibinfo
+  {pages} {12562} (\bibinfo {year} {2002})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Favrin}, \citenamefont {Irb{\"a}ck},\ and\
+  \citenamefont {Sjunnesson}(2001)}]{Favrin2001}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Favrin}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Irb{\"a}ck}},
+  \ and\ \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Sjunnesson}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J. Comp. Phys.}\
+  }\textbf {\bibinfo {volume} {114}},\ \bibinfo {pages} {8154} (\bibinfo {year}
+  {2001})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Gelman}, \citenamefont {Roberts},\ and\ \citenamefont
+  {Gilks}(1996)}]{Gilks1996}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Gelman}}, \bibinfo {author} {\bibfnamefont {G.~O.}\ \bibnamefont {Roberts}},
+  \ and\ \bibinfo {author} {\bibfnamefont {W.~R.}\ \bibnamefont {Gilks}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Bayesian Statist}\
+  }\textbf {\bibinfo {volume} {5}},\ \bibinfo {pages} {599} (\bibinfo {year}
+  {1996})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Gansner}\ and\ \citenamefont
+  {North}(2000)}]{Graphviz2000}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.~R.}\ \bibnamefont
+  {Gansner}}\ and\ \bibinfo {author} {\bibfnamefont {S.~C.}\ \bibnamefont
+  {North}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Software -- practice and experience}\ }\textbf {\bibinfo {volume} {30}},\
+  \bibinfo {pages} {1203} (\bibinfo {year} {2000})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Sikosek}, \citenamefont {Bornberg-Bauer},\ and\
+  \citenamefont {Chan}(2012)}]{Sikosek2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Sikosek}}, \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
+  {Bornberg-Bauer}}, \ and\ \bibinfo {author} {\bibfnamefont {H.~S.}\
+  \bibnamefont {Chan}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
+  {journal} {PLOS Comput Biol}\ }\textbf {\bibinfo {volume} {8}},\ \bibinfo
+  {pages} {e1002659} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Meinke}\ and\ \citenamefont
+  {Hansmann}(2007)}]{Meinke2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~H.}\ \bibnamefont
+  {Meinke}}\ and\ \bibinfo {author} {\bibfnamefont {U.~H.~E.}\ \bibnamefont
+  {Hansmann}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
+  Phys Condens Matter}\ }\textbf {\bibinfo {volume} {19}},\ \bibinfo {pages}
+  {285215} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zhang}\ and\ \citenamefont {Chen}(2014)}]{Chen2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
+  {Zhang}}\ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Chen}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J Chem Theory
+  Comput}\ }\textbf {\bibinfo {volume} {10}},\ \bibinfo {pages} {918} (\bibinfo
+  {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bernhardt}\ \emph {et~al.}(2016)\citenamefont
+  {Bernhardt}, \citenamefont {Xi}, \citenamefont {Wang},\ and\ \citenamefont
+  {Hansmann}}]{Bernhardt2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~A.}\ \bibnamefont
+  {Bernhardt}}, \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont {Xi}},
+  \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont {Wang}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {U.~H.}\ \bibnamefont {Hansmann}},\ }\href@noop {}
+  {\bibfield  {journal} {\bibinfo  {journal} {J Chem Theory Comput}\ }\textbf
+  {\bibinfo {volume} {12}},\ \bibinfo {pages} {5656} (\bibinfo {year}
+  {2016})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Das}\ and\ \citenamefont {Pappu}(2013)}]{Das2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~K.}\ \bibnamefont
+  {Das}}\ and\ \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont
+  {Pappu}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Proc
+  Natl Acad Sci USA}\ }\textbf {\bibinfo {volume} {110}},\ \bibinfo {pages}
+  {13392} (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Meng}\ \emph {et~al.}(2013)\citenamefont {Meng},
+  \citenamefont {Luan}, \citenamefont {Lyle}, \citenamefont {Pappu},\ and\
+  \citenamefont {Raleigh}}]{Meng2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
+  {Meng}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Luan}}, \bibinfo
+  {author} {\bibfnamefont {N.}~\bibnamefont {Lyle}}, \bibinfo {author}
+  {\bibfnamefont {R.~V.}\ \bibnamefont {Pappu}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {D.~P.}\ \bibnamefont {Raleigh}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Biochemistry}\ }\textbf {\bibinfo {volume}
+  {52}},\ \bibinfo {pages} {2662} (\bibinfo {year} {2013})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Serohijos}\ and\ \citenamefont
+  {Shakhnovich}(2014)}]{Serohijos2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~W.}\ \bibnamefont
+  {Serohijos}}\ and\ \bibinfo {author} {\bibfnamefont {E.~I.}\ \bibnamefont
+  {Shakhnovich}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Curr Opin Struct Biol}\ }\textbf {\bibinfo {volume} {26}},\ \bibinfo {pages}
+  {84} (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bereau}\ and\ \citenamefont
+  {Deserno}(2015)}]{Bereau2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Bereau}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Deserno}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
+  Membr Biol}\ }\textbf {\bibinfo {volume} {248}},\ \bibinfo {pages} {395}
+  (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Curuksu}\ and\ \citenamefont
+  {Zacharias}(2009)}]{Curuksu2009}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Curuksu}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Zacharias}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Comp. Phys.}\ }\textbf {\bibinfo {volume} {130}},\ \bibinfo {pages} {104110}
+  (\bibinfo {year} {2009})}\BibitemShut {NoStop}%
+\end{thebibliography}%
 

### 2017-10-05 JCP_submit/Manuscript.tex
Final submission and revision

--- 2017-10-05 JCP_revision/Manuscript.tex
@@
 amsmath,amssymb,
-preprint,%
-%reprint,%
+%preprint,%
+reprint,%
 %author-year,%
@@
 \usepackage{bm}% bold math
-\usepackage{color}
 %\usepackage[mathlines]{lineno}% Enable numbering of text and display math
@@
 
-\title[Multisequence biomolecular simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: exploring the sequence-structure relationship of proteins}
+\title[Multisequence Monte Carlo simulations]{Multisequence algorithm for coarse-grained biomolecular simulations: exploring the sequence-structure relationship of proteins}
 
@@
 
-Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} Together with the ever-increasing availability of computational resources, these advances have triggered a few major efforts~\cite{McGuffee2010,Miao2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the dynamics of biomolecular systems of various sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} and a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging and insightful, these large-scale simulations have also highlighted that severe tradeoffs in size and time scales will likely persist for the foreseeable future.
-
-One way to expand the range of biomolecular simulations is to turn to coarse-grained (CG) models, where the basic aim is to simplify the description of physical interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list 4 main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that a CG representation of either the interaction potential or the molecular geometry often opens up for alternative sampling schemes beyond traditional molecular dynamics approaches, which can further speed up conformational sampling.
+Recent years have seen important advances in biomolecular simulation methods, including improvements to standard molecular dynamics force fields,~\cite{Piana2014} the advent of several alternative atomistic simulation approaches,~\cite{Ding2008,Irback2006,Verma2009,Yang2007} and new techniques for  conformational sampling.~\cite{Bernardi2015} Together with the ever-increasing availability of computational resources, these advances have triggered a few major efforts~\cite{McGuffee2010,Miao2010,Perilla2016,Lindorff-Larsen2011,Yu2016} to characterize the dynamics of biomolecular systems of various sizes, e.g., a small native protein on the millisecond scale~\cite{Lindorff-Larsen2011} and a comprehensive model cytoplasm on the nanosecond scale.~\cite{Yu2016} While encouraging and insightful, these large-scale simulations have also highlighted the fact that severe tradeoffs in size and time scales will likely persist for the foreseeable future.
+
+One way to expand the range of biomolecular simulations is to turn to coarse-grained (CG) models, where the basic aim is to simplify the physical description of interactions while retaining the essential physics of the system under study.~\cite{Riniker2012} Ingolfsson \textit{et al.} list 4 main factors that make CG models computationally fast: reduction in the number of degrees of freedom, faster simulation dynamics, emphasis on short-range interactions and the ability of using larger integration time steps.~\cite{Ingolfsson2014} To this list can be added that a CG representation of either the interaction potential or the molecular geometry often opens up for alternative sampling schemes beyond traditional molecular dynamics approaches, which can further speed up conformational sampling.
 Examples of such sampling schemes include activation-relaxation kinetics,~\cite{Beland2011} discrete molecular dynamics~\cite{Proctor2011} and various Monte Carlo (MC)-based techniques such as cluster moves.~\cite{Vitalis2009}
@@
 
-As a test application of the MS algorithm, we selected the phenomenon of protein fold switching which recently was demonstrated in a handful of natural and engineered proteins.
-These special proteins exhibit a unique ability to reversibly switch between entirely different folds, with accompanying changes in secondary structure, hydrophobic core packing and overall shape.~\cite{Bryan2010} The fold switching transitions found in natural proteins typically play a functional role.
-For example, rare transitions to an alternative fold in the protein KaiB provide a crucial time delay mechanism in the circadian clock cycle of cyanobacteria.~\cite{LiWang2015} Fold switching can occur either spontaneously~\cite{Meier2007} or be triggered by various signals including changes to solution conditions,~\cite{Kuloglu2002} subdomain detachment,~\cite{Burmann2012} ligand binding~\cite{Luo2004} and point mutations.~\cite{Alexander2009} Computational studies, using  coarse-grained~\cite{Holzgrafe2014,Holzgrafe2015,Kouza2012,Sutto2012} or atomistic~\cite{Ramirez-Sarmiento2015,Sikosek2016,Hansen2013} models, have attempted to explain how proteins can exhibit multiple folding funnels and how they are altered in response to binding events or changes in sequence.
-
-The discovery that proteins can be driven to switch folds through an accumulation of point mutations, in particular, holds implications for protein evolution as it suggests a simple mechanism of fold evolution.~\cite{Sikosek2014} Alexander \textit{et al.} demonstrated that the similarly sized but structurally distinct A ($3\alpha$) and B ($4\beta+\alpha$) domains of protein G could, after extensive mutations leaving their respective folds undisturbed, be triggered to switch folds by applying one additional mutation, Y45L, located at the edge of the hydrophobic core in the B domain.~\cite{Alexander2009} This remarkable discovery suggests the possibility that the two domains might be evolutionary related despite a lack of detectable similarity in either sequence or structure in wild-type protein G,~\cite{He2012} although this has yet to be proven.
-Moreover, it is unclear how common such fold-to-fold transitions are and how they might occur in evolutionary processes.~\cite{Porter2015} In previous work~\cite{Holzgrafe2014,Holzgrafe2015} we showed that mutational paths with abrupt fold switching exist between two other pairs of smaller protein folds within the framework of our CG model.~\cite{Bhattacherjee2012}
-
-In demonstrating mutation-induced fold switching in our model we characterized the folding of a set of 144 different model sequences with 16 amino acids.
-This set (denoted here {\SR}) was constructed to sparsely span the sequence space between two ideally designed sequences, A1 and N1, folding into an $\alpha$-helix and a $\beta$-hairpin, respectively, as shown in Fig.~2.
-Here we use the set {\SR} to validate the MS method and compare its computational efficiency to a standard generalized-ensemble method.~\cite{Marinari1992,Lyubartsev1992} We thereafter greatly enlarge {\SR} to a set with 1,024 sequences as well as another set of the same size spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively (see Fig.~2).
+%While this may be hard to achieve for detailed biomolecular models, it is likely possible in many CG models.
+
+As an initial test case, we apply the MS method to a set of 144 model protein sequences with 16 amino acids.
+This set (denoted here {\SR}) was constructed to sparsely cover the sequence space between two ideally designed sequences, A1 and N1, which fold into an $\alpha$-helix and a $\beta$-hairpin,~\cite{Bhattacherjee2012} respectively, as shown in Fig.~2.
+In a previous work, this set allowed us to demonstrate the existence of A1-N1 mutational pathways which do not proceed through any disordered (non-folding) sequences but instead switch abruptly between the two folds in a single mutational step.~\cite{Holzgrafe2014} Such an ability to reversibly switch folds was recently demonstrated in a handful of natural and engineered proteins.~\cite{Bryan2010} Here we use the set {\SR} and our CG model for protein folding to validate the MS method and compare its efficiency to a standard generalized-ensemble method.~\cite{Marinari1992,Lyubartsev1992} We thereafter greatly enlarge {\SR} to a set with 1,024 sequences, also spanning the A1-N1 space, as well as another set of the same size spanning two 35-amino acid sequences, A2 and TN, that fold into two-helical bundle and mixed $\alpha$-$\beta$ structures, respectively (see Fig.~2).
 Besides demonstrating that the MS method can be applied to large numbers of sequences, the results allow us to carry out a more systematic analysis of the biophysical properties of sequences along mutational pathways connecting these two pairs of basic folds than has been previously possible.
@@
 
-Conventional Metropolis Monte Carlo simulations of the canonical distribution is problematic at low temperatures for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
-The basic idea of generalized-ensemble algorithms is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters~\cite{Mitsutake2001} (for a recent historical account see Ref.~\citenum{Berg2017}).
-Generalized-ensemble methods that have been frequently used for biomolecular simulations include simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} replica exchange~\cite{Swendsen1986}  or parallel tempering,~\cite{Hansmann1997} and metadynamics.~\cite{Laio2002}
-
-ST is a direct extension to the Metropolis algorithm in which the temperature $T$ becomes a dynamic parameter.
-In this way, frequent visits to high $T$ allow simulations to readily escape from local energy minima.
-The algorithm thus simulates the joint probability distribution
+Conventional Monte Carlo simulations of the canonical distribution is problematic at low temperatures $T$ for many physical systems because simulations tend to become trapped in local energy minima and hamper representative sampling of configurational space.
+The basic idea of generalized-ensemble algorithms~\cite{Mitsutake2001} is to alleviate this trapping problem by sampling states using a non-Boltzmann weight factor and/or expand the state space with additional dynamical parameters.
+
+A well-known generalized-ensemble algorithm is simulated tempering (ST),~\cite{Marinari1992,Lyubartsev1992} where it is the temperature that becomes a dynamic parameter.
+In this scheme, frequent visits to high $T$ allow simulations to readily escape from local energy minima.
+The ST algorithm simulates the joint probability distribution
 \begin{equation}
-P(m,\rbar) =\dfrac{1}{\hat{Z}} e^{-\beta_m E(\rbar) + g_m}\,,
+P(\rbar,m) =\dfrac{1}{\hat{Z}} e^{-\beta_m E(\rbar) + g_m}\,,
 \label{ST}
@@
 This is the case in our coarse-grained protein model which has only backbone degrees of freedom.
-It can also be achieved in some more detailed models.~\cite{Bhattacherjee2013,Wallin2017}
+It can also be achieved for in some more detailed models.~\cite{Bhattacherjee2013,Wallin2017}
 
@@
 \begin{equation}
-Z = \sum_{\sbar\in\mathrm{S}}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)}\,
-\label{eq:ZMS}
+Z = \sum_{\sbar}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)}\,
 \end{equation}
-and S is a set of pre-selected sequences, i.e., the sequences to which visits are allowed during a simulation.
+and the first sum goes over a set of allowed sequences $\sbar$.
 The simulation parameters $h(\sbar)$, similar to the parameters $g_m$ in ST, control the marginal distribution $P(\sbar)=Z^{-1}\sum_{\rbar} e^{-\beta E(\sbar,\rbar)+ h(\sbar)} = \tilde{Z}^{-1}e^{-\beta F(\sbar)+ h(\sbar)}$ and a roughly flat $P(\sbar)$ can be achieved by choosing $h(\sbar) \approx \beta F(\sbar)$, where $F(\sbar)$ is the free energy of sequence $\sbar$ at temperature $T$.
 
-Two types of MC updates are required to sample from the distrubution in Eq.~\ref{MS}, ordinary conformational updates $\rbar\rightarrow\rbar'$ and sequence updates $\sbar\rightarrow\sbar'$.
-The acceptance probability for the latter  becomes
+Two types of MC updates are required to sample from the distrubution in Eq.~\ref{MS}, ordinary conformational update $\rbar\rightarrow\rbar'$ and mutational updates $\sbar\rightarrow\sbar'$.
+The acceptance probabilities for the latter  becomes
 \begin{equation}
@@
 where $\Delta E = E(\sbar',\rbar)-E(\sbar,\rbar)$ and $\Delta h = h({\sbar'})-h(\sbar)$.
-
-{\color{red}Picking a new sequence $\sbar'$ in a sequence update $\sbar\rightarrow\sbar'$ can be done in several ways.
-One possibility is to draw $\sbar'$ randomly from the set S, such than $\sbar'\ne\sbar$.
-Alternatively, a type of ``mutational" move can be used where an amino acid position is first picked and then assigned a new amino acid type.
-The selection of position and type would have to be chosen such that $\sbar'$ does not end up outside S.
-In this work, we use the former update which is general and guarantees that ergodicity is fulfilled for any S.
-Importantly, both updates fulfill detailed balance and therefore lead to the same estimates of equilibrium quantities, such as native state stabilities, for the different sequences in S.}
 
@@
 
-\subsection{Monte Carlo simulation parameters and updates}
-
-Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assign a new value between $-\pi$ and $\pi$; and (2) a semi-local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2001}
-
-%, and where S is the set of allowed (pre-selected) sequences, S (see Eq.~\ref{eq:ZMS})
-%i.e., given by the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$ angles,
-
-In our MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out in the following way.
-First, a new sequence $\sbar'$ is picked randomly from the set of pre-selected (allowed) sequences S, such that $\sbar'\ne\sbar$.
-This new sequence $\sbar'$ therefore differs from $\sbar$ in one or more amino acid positions.
-Thereafter, the sidechains of the protein, which remains in an unchanged (backbone) conformation $\rbar$, is re-built according to the new sequence $\sbar'$.
-Practically this means that, at the position(s) where the amino acid type has changed, the sidechain is altered according to the type change.
-For example, if $\mathrm{p}\rightarrow\mathrm{t}$, the $\mathrm{C}_\beta$ atom is removed and replaced with an $\mathrm{H}_{\alpha 2}$ atom or, if $\mathrm{p}\rightarrow\mathrm{h}$, the $\mathrm{C}_\beta$ remains in place but its character is changed to hydrophobic.
-Finally, the change in total energy $\Delta E$ is calculated and the accept-reject criterion in Eq.~\ref{accrej} is applied.
-If rejected, the old state $(\sbar,\rbar)$ is restored.
-
-A sequence update is attempted every $1,000$ MC steps while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
-The computational cost for sequence updates is somewhat higher than for temperature updates.
-The latter update does not require any energy calculation and is thus extremely rapid.
-For the purpose of comparing computational efficiencies of ST and MS, we therefore chose sequence updates to be slightly less frequent than temperature updates while both are fairly frequent.
-All simulations carried out in this work are summarized in Table~II.
+\subsection{Monte Carlo simulation parameters}
+
+Both ST and MS simulations are carried out with two types of conformational updates $\rbar\rightarrow\rbar'$: (1) a global pivot move (20\%) which randomly picks a $\phi_\mathrm{i}$ angle or $\psi_\mathrm{i}$ angle and assign a new value between $-\pi$ and $\pi$; and (2) a semi-local move (80\%) which turns the $\phi_\mathrm{i}$ and $\psi_\mathrm{i}$-angles of 4 consecutive amino acids in a coordinated manner.~\cite{Favrin2001} In MS simulations, sequence updates $\sbar\rightarrow\sbar'$ are carried out by randomly picking a new sequence $\sbar'\ne\sbar$ and applying the accept-reject criterion in Eq.~\ref{accrej}.
+A sequence move is attempted every $1,000$ MC steps while temperature updates $m\rightarrow m'$ are attempted every 100 steps.
+The simulations carried out in this work are summarized in Table~II.
 
@@
 \begin{tabular}{lcccr}
-Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes a thermalization step with $10^7$ MC steps/run.} &  Sequences\\
+Runs & Algorithm & $\kb T$  & MC steps/run\footnote{Excludes a thermalization step with $10^6$ MC steps/run.} &  Sequences\\
 \hline
@@
 At the lowest $T$ and highest $\Delta h$, acceptance rates are only around 0.1-0.2.
-However, for most other $T$ and $\Delta h$ the overall acceptance rates are substantially higher and often above the oft-quoted rule-of-thumb value 0.25~\cite{Gilks1996} (see Fig~3B).
+However, for most other $T$ and $\Delta h$ the overall acceptance rate is substantially higher and often above the oft-quoted rule-of-thumb value 0.25~\cite{Gilks1996} (see Fig~4B).
 An increased acceptance rate can easily be achieved by restricting proposed updates such that $\Delta h\le\Delta h_\mathrm{max}$, where $\Delta h_\mathrm{max}$ is a maximum step size, which might be necessary for longer chains.
-For example, $\Delta h_\mathrm{max} = 1$ would be equivalent to applying only a ``mutational" update, i.e., picking a random (allowed) position and changing the amino acid type at that position.
 
@@
 We first validate the MS algorithm by comparing the average total energy, $\left < E \right >$, calculated for these 4 sequences with the two different methods (see Supplementary Information).
-The two sets of results are entirely consistent showing that, for a given $\sbar$ and $T$, the MS and ST algorithms indeed sample the same (Boltzmann) distribution.
+The two sets of results are entirely consistent showing that, for a given $\sbar$ and $T$, the MS and ST algorithms indeed sample the same distribution.
 
@@
 The results indicate that there are two main benefits of this approach.
-Firstly, it provides a convenient way to sample the canonical distributions of large numbers of sequences in a single run and, secondly, it enhances the sampling of conformational space meaning it can be applied directly to low temperatures.
+Firstly, it provides a convenient way to sample the canonical distributions of large numbers of sequences in a single run and, secondly, it enhances the sampling of conformational space meaning it can be applied directly at low temperatures.
 The conformational sampling efficiency can be assessed from the comparison with ST.
@@
 It is notable that this acceleration in conformational sampling vis-{\'a}-vis ST is achieved despite that simulations are carried out at constant $T$.
-Therefore, rather than promoting escape from local minima by visits to higher $T$, as in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986} MS simulations must escape the minima occurring for one sequence through visits to other sequences.
-How this process works can be envisioned by considering an MS simulation that is visiting a sequence $s$ and is trapped in a local minimum, requiring that a high free-energy barrier is overcome for escape.
-The trapped state could be, e.g., a compact $\beta$-sheet rich state with a particular register.
-Eventually, sequence updates will carry the simulation to other sequences $\hat{s}$ while it conformationally still remains in the trapped state.
-However, the free energy barrier of escape might be substantially lower for $\hat{s}$ than for $s$, or the barrier may even be altogether absent if the trapped state is unstable for $\hat{s}$, leading to rapid escape from the minimum.
-The above reasoning also implies that the performance of the MS algorithm likely depends on the size of the sequence set S as well as the conformational properties of the sequences.
-Specifically, the performance of MS simulations of proteins at low $T$ may benefit from the inclusion of at least a few sequences with poor stability properties, such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
-
-Individual protein sequences that exhibit spontaneous transitions between widely different competing conformational states, such as fold switching proteins, are especially challenging to molecular simulation methods.
-Representative sampling in such cases requires multiple transitions between highly different states, which can be a slow process.
-This problem has been addressed by using Hamiltonian replica-exchange techniques to couple G\={o}-like terms, i.e., energy terms with an artificial bias towards a given structure, with a physical force field.~\cite{Meinke2007,Chen2014,Bernhardt2016} This way exchange moves ``feed" diverse conformations into the replica corresponding to the physical force field and enhance sampling.~\cite{Bernhardt2016} We did not specifically study sampling efficiency for sequences exhibiting competing states, such as bistable sequences.
-It appears likely, however, that a very similar type of ``feeding" of conformations would take place in MS simulations although the coupling occurs instead with other sequences rather than with G\={o}-type terms.
+Hence, rather than promoting escape from local minima by visits to higher $T$, as is the case in  ST~\cite{Marinari1992,Lyubartsev1992} or temperature replica-exchange,~\cite{Swendsen1986}  MS simulations escape local minima through visits to sequences where the minima are not as deep or even altogether absent.
+For the same reason, we expect the performance of the MS algorithm to depend on the size the sequence set used, as well as their conformational properties.
+Specifically, the performance of MS simulations at low $T$ may hinge on the inclusion of at least a few sequences with poor stabilities such that partial unfolding of the chain is regularly triggered and thus promoting transitions to new conformational states.
 
 We emphasize that the MS method should not be seen as a general technique to speed up conformational sampling.
-However, our results indicate that for CG models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), our method can be a highly efficient way to sample the equilibrium behavior of many sequences.
-This opens up for various applications, such as exploring sequence effects on the conformational properties of disordered~\cite{Das2013} or denatured~\cite{Meng2013} proteins, or as a tool in efforts to combine population genetics and molecular simulations.~\cite{Serohijos2014} While applied to proteins in this work, we note that the theoretical framework of the MS algorithm is equally valid for other bio-macromolecules, including DNA and RNA.
-The ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unraveling of the basic underlying structure.
+However, our results indicate that for CG models that permit sequence updates to be carried out as a simple Metropolis step and when visits to higher $T$ is unwanted (or unnecessary), the MS method can be a highly efficient way to sample the equilibrium behavior of many sequences.
+In addition, the ability to promote conformational sampling without resorting to an increase in $T$ may make the method useful in simulations of biomolecules in ordered phases, such as lipid bilayers or double-stranded DNA, where escape from local minima can be especially challenging~\cite{Bereau2015,Curuksu2009} and elevated $T$ is typically avoided in simulations because it may lead to unwanted perturbations or unfolding of the basic underlying structure.
 
@@
 The method performs a random walk in sequence space which is especially useful at low temperature as it promotes escape from local minima present in the free energy landscapes of individual sequences.
-The method might be suitable for CG simulations of various other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
-
-\section{Supplementary Material}
-
-See supplementary material for the validation of the MS algorithm.
+The method might be suitable for CG simulations of other biomolecular systems, such as peptides in phospholipid bilayers, where sampling at elevated temperatures is not desirable.
 
@@
 
-This work was supported by grants from Memorial University and National Sciences and Engineering Research Council of Canada (RGPIN-2016-05014).
+This work was supported by grants from Memorial University and National Sciences and Engineering Research Council of Canada (NSERC).
 
@@
 %\input{Manuscript.bbl}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %merlin.mbs aipnum4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
@@
 %Control: production of eprint (0) enabled
-\begin{thebibliography}{63}%
+\begin{thebibliography}{45}%
 \makeatletter
@@
   {201} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Holzgr{\"a}fe}\ and\ \citenamefont
+  {Wallin}(2014)}]{Holzgrafe2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Holzgr{\"a}fe}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Biophys J}\ }\textbf {\bibinfo {volume} {107}},\ \bibinfo {pages} {1217}
+  (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Bryan}\ and\ \citenamefont
@@
   {year} {2010})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Chang}\ \emph {et~al.}(2015)\citenamefont {Chang},
-  \citenamefont {Cohen}, \citenamefont {Phong}, \citenamefont {Myers},
-  \citenamefont {Kim}, \citenamefont {Tseng}, \citenamefont {Lin},
-  \citenamefont {Zhang}, \citenamefont {Boyd}, \citenamefont {Lee},
-  \citenamefont {Kang}, \citenamefont {Lee}, \citenamefont {Li}, \citenamefont
-  {Britt}, \citenamefont {Rust}, \citenamefont {Golden},\ and\ \citenamefont
-  {LiWang}}]{LiWang2015}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.~G.}\ \bibnamefont
-  {Chang}}, \bibinfo {author} {\bibfnamefont {S.~E.}\ \bibnamefont {Cohen}},
-  \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Phong}}, \bibinfo
-  {author} {\bibfnamefont {W.~K.}\ \bibnamefont {Myers}}, \bibinfo {author}
-  {\bibfnamefont {Y.~I.}\ \bibnamefont {Kim}}, \bibinfo {author} {\bibfnamefont
-  {R.}~\bibnamefont {Tseng}}, \bibinfo {author} {\bibfnamefont
-  {J.}~\bibnamefont {Lin}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
-  {Zhang}}, \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont {Boyd}},
-  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Lee}}, \bibinfo {author}
-  {\bibfnamefont {S.}~\bibnamefont {Kang}}, \bibinfo {author} {\bibfnamefont
-  {D.}~\bibnamefont {Lee}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Li}}, \bibinfo {author} {\bibfnamefont {R.~D.}\ \bibnamefont {Britt}},
-  \bibinfo {author} {\bibfnamefont {M.~J.}\ \bibnamefont {Rust}}, \bibinfo
-  {author} {\bibfnamefont {S.~S.}\ \bibnamefont {Golden}}, \ and\ \bibinfo
-  {author} {\bibfnamefont {A.}~\bibnamefont {LiWang}},\ }\href@noop {}
-  {\bibfield  {journal} {\bibinfo  {journal} {Science}\ }\textbf {\bibinfo
-  {volume} {349}},\ \bibinfo {pages} {324} (\bibinfo {year}
-  {2015})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Meier}\ \emph {et~al.}(2007)\citenamefont {Meier},
-  \citenamefont {Jensen}, \citenamefont {David}, \citenamefont {Chapman},
-  \citenamefont {Holstein}, \citenamefont {Grzesiek},\ and\ \citenamefont
-  {Ozbek}}]{Meier2007}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Meier}}, \bibinfo {author} {\bibfnamefont {P.~R.}\ \bibnamefont {Jensen}},
-  \bibinfo {author} {\bibfnamefont {C.~N.}\ \bibnamefont {David}}, \bibinfo
-  {author} {\bibfnamefont {J.}~\bibnamefont {Chapman}}, \bibinfo {author}
-  {\bibfnamefont {T.~W.}\ \bibnamefont {Holstein}}, \bibinfo {author}
-  {\bibfnamefont {S.}~\bibnamefont {Grzesiek}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {S.}~\bibnamefont {Ozbek}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Curr Biol}\ }\textbf {\bibinfo {volume}
-  {17}},\ \bibinfo {pages} {173} (\bibinfo {year} {2007})}\BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {Kuloglu}\ \emph {et~al.}(2002)\citenamefont
-  {Kuloglu}, \citenamefont {McCaslin}, \citenamefont {Markley},\ and\
-  \citenamefont {Volkman}}]{Kuloglu2002}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.~S.}\ \bibnamefont
-  {Kuloglu}}, \bibinfo {author} {\bibfnamefont {D.~R.}\ \bibnamefont
-  {McCaslin}}, \bibinfo {author} {\bibfnamefont {J.~L.}\ \bibnamefont
-  {Markley}}, \ and\ \bibinfo {author} {\bibfnamefont {B.~F.}\ \bibnamefont
-  {Volkman}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
-  Biol Chem}\ }\textbf {\bibinfo {volume} {277}},\ \bibinfo {pages} {17863}
-  (\bibinfo {year} {2002})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Burmann}\ \emph {et~al.}(2012)\citenamefont
-  {Burmann}, \citenamefont {Knauer}, \citenamefont {Sevostyanova},
-  \citenamefont {Schweimer}, \citenamefont {Mooney}, \citenamefont {Landick},
-  \citenamefont {Artsimovitch},\ and\ \citenamefont {R{\"o}sch}}]{Burmann2012}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.~M.}\ \bibnamefont
-  {Burmann}}, \bibinfo {author} {\bibfnamefont {S.~H.}\ \bibnamefont {Knauer}},
-  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Sevostyanova}}, \bibinfo
-  {author} {\bibfnamefont {K.}~\bibnamefont {Schweimer}}, \bibinfo {author}
-  {\bibfnamefont {R.~A.}\ \bibnamefont {Mooney}}, \bibinfo {author}
-  {\bibfnamefont {R.}~\bibnamefont {Landick}}, \bibinfo {author} {\bibfnamefont
-  {I.}~\bibnamefont {Artsimovitch}}, \ and\ \bibinfo {author} {\bibfnamefont
-  {P.}~\bibnamefont {R{\"o}sch}},\ }\href@noop {} {\bibfield  {journal}
-  {\bibinfo  {journal} {Cell}\ }\textbf {\bibinfo {volume} {150}},\ \bibinfo
-  {pages} {291} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Luo}\ \emph {et~al.}(2004)\citenamefont {Luo},
-  \citenamefont {Tang}, \citenamefont {Xia}, \citenamefont {Wassmann},
-  \citenamefont {Matsumoto}, \citenamefont {Rizo},\ and\ \citenamefont
-  {Yu}}]{Luo2004}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
-  {Luo}}, \bibinfo {author} {\bibfnamefont {Z.}~\bibnamefont {Tang}}, \bibinfo
-  {author} {\bibfnamefont {G.}~\bibnamefont {Xia}}, \bibinfo {author}
-  {\bibfnamefont {K.}~\bibnamefont {Wassmann}}, \bibinfo {author}
-  {\bibfnamefont {T.}~\bibnamefont {Matsumoto}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Rizo}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {H.}~\bibnamefont {Yu}},\ }\href@noop {} {\bibfield  {journal}
-  {\bibinfo  {journal} {Nat Struct Mol Biol}\ }\textbf {\bibinfo {volume}
-  {11}},\ \bibinfo {pages} {338} (\bibinfo {year} {2004})}\BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {Alexander}\ \emph {et~al.}(2009)\citenamefont
-  {Alexander}, \citenamefont {He}, \citenamefont {Chen}, \citenamefont
-  {Orban},\ and\ \citenamefont {Bryan}}]{Alexander2009}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.~A.}\ \bibnamefont
-  {Alexander}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {He}},
-  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Chen}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Orban}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {P.~N.}\ \bibnamefont {Bryan}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Proc Natl Acad Sci USA}\ }\textbf {\bibinfo
-  {volume} {106}},\ \bibinfo {pages} {21149} (\bibinfo {year}
-  {2009})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Holzgr{\"a}fe}\ and\ \citenamefont
-  {Wallin}(2014)}]{Holzgrafe2014}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
-  {Holzgr{\"a}fe}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Biophys J}\ }\textbf {\bibinfo {volume} {107}},\ \bibinfo {pages} {1217}
-  (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Holzgr{\"a}fe}\ and\ \citenamefont
-  {Wallin}(2015)}]{Holzgrafe2015}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
-  {Holzgr{\"a}fe}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
-  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys
-  Biol}\ }\textbf {\bibinfo {volume} {12}},\ \bibinfo {pages} {026002}
-  (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Kouza}\ and\ \citenamefont
-  {Hansmann}(2012)}]{Kouza2012}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Kouza}}\ and\ \bibinfo {author} {\bibfnamefont {U.~H.}\ \bibnamefont
-  {Hansmann}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
-  Phys Chem B}\ }\textbf {\bibinfo {volume} {116}},\ \bibinfo {pages} {6645}
-  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Sutto}\ and\ \citenamefont
-  {Camilloni}(2012)}]{Sutto2012}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
-  {Sutto}}\ and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
-  {Camilloni}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
-  Comp. Phys.}\ }\textbf {\bibinfo {volume} {136}},\ \bibinfo {pages} {185101}
-  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Ramirez-Sarmiento}\ \emph {et~al.}(2015)\citenamefont
-  {Ramirez-Sarmiento}, \citenamefont {Noel}, \citenamefont {Valenzuela},\ and\
-  \citenamefont {Artsimovitch}}]{Ramirez-Sarmiento2015}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.~A.}\ \bibnamefont
-  {Ramirez-Sarmiento}}, \bibinfo {author} {\bibfnamefont {J.~K.}\ \bibnamefont
-  {Noel}}, \bibinfo {author} {\bibfnamefont {S.~L.}\ \bibnamefont
-  {Valenzuela}}, \ and\ \bibinfo {author} {\bibfnamefont {I.}~\bibnamefont
-  {Artsimovitch}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {PLOS Comput Biol}\ }\textbf {\bibinfo {volume} {11}},\ \bibinfo {pages}
-  {e1004379} (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Sikosek}, \citenamefont {Krobath},\ and\
-  \citenamefont {Chan}(2016)}]{Sikosek2016}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
-  {Sikosek}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Krobath}}, \
-  and\ \bibinfo {author} {\bibfnamefont {H.~S.}\ \bibnamefont {Chan}},\
-  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {PLOS Comput Biol}\
-  }\textbf {\bibinfo {volume} {12}},\ \bibinfo {pages} {e1004960} (\bibinfo
-  {year} {2016})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Hansen}\ \emph {et~al.}(2013)\citenamefont {Hansen},
-  \citenamefont {Allison}, \citenamefont {Hodel},\ and\ \citenamefont {van
-  Gunsteren}}]{Hansen2013}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
-  {Hansen}}, \bibinfo {author} {\bibfnamefont {J.~R.}\ \bibnamefont {Allison}},
-  \bibinfo {author} {\bibfnamefont {F.~H.}\ \bibnamefont {Hodel}}, \ and\
-  \bibinfo {author} {\bibfnamefont {W.~F.}\ \bibnamefont {van Gunsteren}},\
-  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Biochemistry}\
-  }\textbf {\bibinfo {volume} {52}},\ \bibinfo {pages} {4962} (\bibinfo {year}
-  {2013})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Sikosek}\ and\ \citenamefont
-  {Chan}(2014)}]{Sikosek2014}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
-  {Sikosek}}\ and\ \bibinfo {author} {\bibfnamefont {H.~S.}\ \bibnamefont
-  {Chan}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J R Soc
-  Interface}\ }\textbf {\bibinfo {volume} {11}},\ \bibinfo {pages} {20140419}
-  (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {He}\ \emph {et~al.}(2012)\citenamefont {He},
-  \citenamefont {Chen}, \citenamefont {Alexander}, \citenamefont {Bryan},\ and\
-  \citenamefont {Orban}}]{He2012}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
-  {He}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Chen}}, \bibinfo
-  {author} {\bibfnamefont {P.~A.}\ \bibnamefont {Alexander}}, \bibinfo {author}
-  {\bibfnamefont {P.~N.}\ \bibnamefont {Bryan}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Orban}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Structure}\ }\textbf {\bibinfo {volume}
-  {20}},\ \bibinfo {pages} {283} (\bibinfo {year} {2012})}\BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {Porter}\ \emph {et~al.}(2015)\citenamefont {Porter},
-  \citenamefont {He}, \citenamefont {Chen}, \citenamefont {Orban},\ and\
-  \citenamefont {Bryan}}]{Porter2015}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.~L.}\ \bibnamefont
-  {Porter}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {He}}, \bibinfo
-  {author} {\bibfnamefont {Y.}~\bibnamefont {Chen}}, \bibinfo {author}
-  {\bibfnamefont {J.}~\bibnamefont {Orban}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {P.~N.}\ \bibnamefont {Bryan}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Biophys J}\ }\textbf {\bibinfo {volume}
-  {108}},\ \bibinfo {pages} {154} (\bibinfo {year} {2015})}\BibitemShut
-  {NoStop}%
 \bibitem [{\citenamefont {Marinari}\ and\ \citenamefont
@@
   {2001})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Berg}(2017)}]{Berg2017}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {B.~A.}\ \bibnamefont
-  {Berg}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Eur Phys
-  J Spec Top}\ }\textbf {\bibinfo {volume} {226}},\ \bibinfo {pages} {,
-  551} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Swendsen}\ and\ \citenamefont
-  {Wang}(1986)}]{Swendsen1986}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~H.}\ \bibnamefont
-  {Swendsen}}\ and\ \bibinfo {author} {\bibfnamefont {J.-S.}\ \bibnamefont
-  {Wang}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
-  Rev. Lett.}\ }\textbf {\bibinfo {volume} {57}},\ \bibinfo {pages} {2607}
-  (\bibinfo {year} {1986})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Hansmann}(1997)}]{Hansmann1997}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {U.~H.~E.}\
-  \bibnamefont {Hansmann}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo
-  {journal} {Chem Phys Lett}\ }\textbf {\bibinfo {volume} {281}},\ \bibinfo
-  {pages} {140} (\bibinfo {year} {1997})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Laio}\ and\ \citenamefont
-  {Parrinello}(2002)}]{Laio2002}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
-  {Laio}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
-  {Parrinello}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Proc Natl Acad Sci USA}\ }\textbf {\bibinfo {volume} {99}},\ \bibinfo
-  {pages} {12562} (\bibinfo {year} {2002})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Favrin}, \citenamefont {Irb{\"a}ck},\ and\
@@
   {2001})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Holzgr{\"a}fe}\ and\ \citenamefont
+  {Wallin}(2015)}]{Holzgrafe2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Holzgr{\"a}fe}}\ and\ \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
+  {Wallin}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys
+  Biol}\ }\textbf {\bibinfo {volume} {12}},\ \bibinfo {pages} {026002}
+  (\bibinfo {year} {2015})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Gelman}, \citenamefont {Roberts},\ and\ \citenamefont
@@
   \bibinfo {pages} {1203} (\bibinfo {year} {2000})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Alexander}\ \emph {et~al.}(2009)\citenamefont
+  {Alexander}, \citenamefont {He}, \citenamefont {Chen}, \citenamefont
+  {Orban},\ and\ \citenamefont {Bryan}}]{Alexander2009}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.~A.}\ \bibnamefont
+  {Alexander}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {He}},
+  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Chen}}, \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Orban}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {P.~N.}\ \bibnamefont {Bryan}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Proc Natl Acad Sci USA}\ }\textbf {\bibinfo
+  {volume} {106}},\ \bibinfo {pages} {21149} (\bibinfo {year}
+  {2009})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {He}\ \emph {et~al.}(2012)\citenamefont {He},
+  \citenamefont {Chen}, \citenamefont {Alexander}, \citenamefont {Bryan},\ and\
+  \citenamefont {Orban}}]{He2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {He}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Chen}}, \bibinfo
+  {author} {\bibfnamefont {P.~A.}\ \bibnamefont {Alexander}}, \bibinfo {author}
+  {\bibfnamefont {P.~N.}\ \bibnamefont {Bryan}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Orban}},\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Structure}\ }\textbf {\bibinfo {volume}
+  {20}},\ \bibinfo {pages} {283} (\bibinfo {year} {2012})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Kouza}\ and\ \citenamefont
+  {Hansmann}(2012)}]{Kouza2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Kouza}}\ and\ \bibinfo {author} {\bibfnamefont {U.~H.}\ \bibnamefont
+  {Hansmann}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
+  Phys Chem B}\ }\textbf {\bibinfo {volume} {116}},\ \bibinfo {pages} {6645}
+  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Sikosek}, \citenamefont {Krobath},\ and\
+  \citenamefont {Chan}(2016)}]{Sikosek2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Sikosek}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Krobath}}, \
+  and\ \bibinfo {author} {\bibfnamefont {H.~S.}\ \bibnamefont {Chan}},\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {PLOS Comput Biol}\
+  }\textbf {\bibinfo {volume} {12}},\ \bibinfo {pages} {e1004960} (\bibinfo
+  {year} {2016})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Sutto}\ and\ \citenamefont
+  {Camilloni}(2012)}]{Sutto2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Sutto}}\ and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {Camilloni}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J.
+  Comp. Phys.}\ }\textbf {\bibinfo {volume} {136}},\ \bibinfo {pages} {185101}
+  (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Sikosek}, \citenamefont {Bornberg-Bauer},\ and\
@@
   {pages} {e1002659} (\bibinfo {year} {2012})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Meinke}\ and\ \citenamefont
-  {Hansmann}(2007)}]{Meinke2007}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~H.}\ \bibnamefont
-  {Meinke}}\ and\ \bibinfo {author} {\bibfnamefont {U.~H.~E.}\ \bibnamefont
-  {Hansmann}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J
-  Phys Condens Matter}\ }\textbf {\bibinfo {volume} {19}},\ \bibinfo {pages}
-  {285215} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Zhang}\ and\ \citenamefont {Chen}(2014)}]{Chen2014}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
-  {Zhang}}\ and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Chen}},\
-  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {J Chem Theory
-  Comput}\ }\textbf {\bibinfo {volume} {10}},\ \bibinfo {pages} {918} (\bibinfo
-  {year} {2014})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Bernhardt}\ \emph {et~al.}(2016)\citenamefont
-  {Bernhardt}, \citenamefont {Xi}, \citenamefont {Wang},\ and\ \citenamefont
-  {Hansmann}}]{Bernhardt2016}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~A.}\ \bibnamefont
-  {Bernhardt}}, \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont {Xi}},
-  \bibinfo {author} {\bibfnamefont {W.}~\bibnamefont {Wang}}, \ and\ \bibinfo
-  {author} {\bibfnamefont {U.~H.}\ \bibnamefont {Hansmann}},\ }\href@noop {}
-  {\bibfield  {journal} {\bibinfo  {journal} {J Chem Theory Comput}\ }\textbf
-  {\bibinfo {volume} {12}},\ \bibinfo {pages} {5656} (\bibinfo {year}
-  {2016})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Das}\ and\ \citenamefont {Pappu}(2013)}]{Das2013}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~K.}\ \bibnamefont
-  {Das}}\ and\ \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont
-  {Pappu}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Proc
-  Natl Acad Sci USA}\ }\textbf {\bibinfo {volume} {110}},\ \bibinfo {pages}
-  {13392} (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
-\bibitem [{\citenamefont {Meng}\ \emph {et~al.}(2013)\citenamefont {Meng},
-  \citenamefont {Luan}, \citenamefont {Lyle}, \citenamefont {Pappu},\ and\
-  \citenamefont {Raleigh}}]{Meng2013}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {W.}~\bibnamefont
-  {Meng}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Luan}}, \bibinfo
-  {author} {\bibfnamefont {N.}~\bibnamefont {Lyle}}, \bibinfo {author}
-  {\bibfnamefont {R.~V.}\ \bibnamefont {Pappu}}, \ and\ \bibinfo {author}
-  {\bibfnamefont {D.~P.}\ \bibnamefont {Raleigh}},\ }\href@noop {} {\bibfield
-  {journal} {\bibinfo  {journal} {Biochemistry}\ }\textbf {\bibinfo {volume}
-  {52}},\ \bibinfo {pages} {2662} (\bibinfo {year} {2013})}\BibitemShut
-  {NoStop}%
-\bibitem [{\citenamefont {Serohijos}\ and\ \citenamefont
-  {Shakhnovich}(2014)}]{Serohijos2014}%
-  \BibitemOpen
-  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~W.}\ \bibnamefont
-  {Serohijos}}\ and\ \bibinfo {author} {\bibfnamefont {E.~I.}\ \bibnamefont
-  {Shakhnovich}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
-  {Curr Opin Struct Biol}\ }\textbf {\bibinfo {volume} {26}},\ \bibinfo {pages}
-  {84} (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Swendsen}\ and\ \citenamefont
+  {Wang}(1986)}]{Swendsen1986}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~H.}\ \bibnamefont
+  {Swendsen}}\ and\ \bibinfo {author} {\bibfnamefont {J.-S.}\ \bibnamefont
+  {Wang}},\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. Lett.}\ }\textbf {\bibinfo {volume} {57}},\ \bibinfo {pages} {2607}
+  (\bibinfo {year} {1986})}\BibitemShut {NoStop}%
 \bibitem [{\citenamefont {Bereau}\ and\ \citenamefont
@@
 \end{thebibliography}%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2017-10-05 Manuscript.tex
Final submission and revision

@@
 where $\Delta E = E(\sbar',\rbar)-E(\sbar,\rbar)$ and $\Delta h = h({\sbar'})-h(\sbar)$.
+
+Picking a new sequence $\sbar'$ in a sequence update $\sbar\rightarrow\sbar'$ can be done in several ways.
+One possibility is to draw $\sbar'$ randomly from the set S, such than $\sbar'\ne\sbar$.
+Alternatively, a type of ``mutational" move can be used where an amino acid position is first picked and then assigned a new amino acid type.
+The selection of position and type would have to be chosen such that $\sbar'$ does not end up outside S.
+In this work, we use the former update which is general and guarantees that ergodicity is fulfilled for any S.
+Importantly, both updates fulfill detailed balance and therefore lead to the same estimates of equilibrium quantities, such as native state stabilities, for the different sequences in S.
 
